### 2016-04-15 Notes/Notes.tex
misc notes

\documentclass[12pt]{article}

% Emulate MS Word
\usepackage{wordlike}

% One inch margins
\PassOptionsToPackage{margin=1in}{geometry}

% Remove footnote indentation
\usepackage[hang,flushmargin]{footmisc} 
\usepackage{amsmath}

% Add extra blank line between paragraphs, and remove paragraph indentation
%\usepackage[parfill]{parskip}
\parskip = 0\baselineskip

% Header
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\setlength{\headheight}{57pt}
\lhead{Hannah Director\\HW 5,  Problem 5}

% Add space between text and footnotes section
\setlength{\skip\footins}{0.5in}

% Add a space before the footnote mark
\let\myfootnote\footnote
\renewcommand{\footnote}[1]{\myfootnote{~#1}}

% Don't justify along the right margin
%\raggedright

% Generate placeholder text
\usepackage{lipsum}

\begin{document}

Misc Notes
\begin{itemize}
\item Figure 3: Generate multivariate draws using standard trick (Higdon slide 4, find formal citation)
\item What units are the velocities given in? How did they calculate the complex-valued velocity (observed by instrument directly or calculated based on reversing distances
\item Possible typo: Eq 13  doesn't match Eq 4 (published code matches Eq 4)
\end{itemize} 
\end{document}

### 2016-04-15 Code/figures.R
updated

@@ -59,10 +59,6 @@ mtext("Replication of Figure 2", outer = T, line = -3)
 #abline(v = f0, col = "red")
 
 ###Figure 3
-rm(list = ls())
-source("/users/hdirector/Dropbox/Prelim/ReplicationCode/functions.R")
-set.seed(103)
-
 ###TO DO: How did they come up with these variance (autocorrelation at 0)?
 
 #sample matern
@@ -90,5 +86,9 @@ plot(DELTA*fbm$omega, fbm$sZ, col = "blue", ylim = c(-20, 80), xlim = c(-pi, pi)
      type = "l", xlab = expression(paste(omega, Delta)), ylab = "dB")
 points(DELTA*mat$omega, mat$sZ, col = "red", type = "l")
 
+
+###Figure 4
+optim(par = c(5, 2, 2, 4, 2, 2), ll, Z = bZ, delta = 2,  method = "L-BFGS-B", control=list(fnscale = -1),
+      lower = c(0, -Inf, 0, 0, 0.5, 0), upper = c(Inf, Inf, Inf, Inf, Inf, Inf))
 
 

### 2016-04-15 Code/functions.R

@@ -73,16 +73,14 @@ simFBM <- function(B, alpha, H, N) {
 
 #Blurred Whittle likelihood
 ll <- function(theta, Z, delta) {
+  A <- theta[1]; w0 <- theta[2]; c <- theta[3];
+  B <- theta[4]; alpha <- theta[5]; h <- theta[6]
   N <- length(Z)
   tau <-  seq(0, N - 1, 1)
   curr <- getPerio(Z, delta)
-  sTau <- ouAc(A = 5, w0 = 2, c = 3, tau, N) + maternAc(B = 10, alpha = 2, h = 2, N)  
+  sTau <- ouAc(A, w0, c, tau, N) + maternAc(B, alpha, h, N)  
   sBar <- Re(fftshift(2*fft(sTau*(1 - tau/N)) - sTau[1])) 
+  print(-sum(curr$sZ/sBar + log(sBar)))
   return(-sum(curr$sZ/sBar + log(sBar)))
 }
 
-
-#TESTING 
-theta <- c(10, .9, 1, 2, 2, 2)
-Z <- readMat("/users/hdirector/Dropbox/Prelim/ReplicationCode/samp.mat")$X
-  

### 2016-04-15 Code/functions.R
likelihood almost working

@@ -1,12 +1,20 @@
 #Calculate periodogram
-getPerio <- function(Z, delta, dB = TRUE) {
+getPerio <- function(Z, delta, dB = TRUE, noZero = TRUE) {
   library("waved")
   #fourier frquencies 
   N <- length(Z)
-  if (N%%2 == 1) {
-    omega <- (2*pi/(N*delta))*c(sort(-seq(1, (N - 1)/2, 1)), seq(1, (N - 1)/2, 1))
+  if(noZero == TRUE) {
+    if (N%%2 == 1) {
+      omega <- (2*pi/(N*delta))*c(sort(-seq(1, (N - 1)/2, 1)), seq(1, (N - 1)/2, 1))
+    } else {
+      omega <- (2*pi/(N*delta))*c(sort(-seq(1, N/2, 1)), seq(1, N/2, 1))
+    }
   } else {
-    omega <- (2*pi/(N*delta))*c(sort(-seq(1, N/2, 1)), seq(1, N/2, 1))
+    if (N%%2 == 1) {
+      omega <- (2*pi/(N*delta))*c(sort(-seq(1, (N - 1)/2, 1)), 0,  seq(1, (N - 1)/2, 1))
+    } else {
+      omega <- (2*pi/(N*delta))*c(sort(-seq(1, N/2, 1)), 0, seq(1, N/2, 1))
+    }
   }
   #convert to decibals
   sZ <- (delta/N)*Mod(fft(Z))^2; sZ <- fftshift(sZ)
@@ -70,17 +78,71 @@ simFBM <- function(B, alpha, H, N) {
   return(Z)
 }
 
-
 #Blurred Whittle likelihood
-ll <- function(theta, Z, delta) {
-  A <- theta[1]; w0 <- theta[2]; c <- theta[3];
-  B <- theta[4]; alpha <- theta[5]; h <- theta[6]
+ll <- function(theta, Z, delta, curr) {
+  #theta = (A, B, w0,c, h, alpha) 
+  #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
+  #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
+  A <- theta[1]; B <- theta[2]; w0 <- theta[3];
+  c <- theta[4]; h <- theta[5]; alpha <- theta[6]
   N <- length(Z)
   tau <-  seq(0, N - 1, 1)
[…]

### 2016-04-15 Code/figures.R
likelihood almost working

@@ -8,7 +8,7 @@ DELTA <- 2
 
 #Convert data from matlab to R
 library("R.matlab")
-drifterbetty2 <- readMat("/users/hdirector/Dropbox/Prelim/ReplicationCode/drifterbetty2.mat")
+drifterbetty2 <- readMat("/users/hdirector/Documents/prelim/prelim/Code/drifterbetty2.mat")
 drift <- list("num" = as.vector(drifterbetty2$drifterbetty[,,1]$num),
               "lat" = as.vector(drifterbetty2$drifterbetty[,,1]$lat),
               "lon" = as.vector(drifterbetty2$drifterbetty[,,1]$lon),

### 2016-04-16 Code/figures.R
likelihood works, optim needs transform

@@ -1,7 +1,7 @@
 ###Hannah Director###
 #Research paper prelim 
 rm(list = ls())
-source("/users/hdirector/Dropbox/Prelim/ReplicationCode/functions.R")
+source("/users/hdirector/Documents/prelim/prelim/Code/functions.R")
 
 #Constants
 DELTA <- 2

### 2016-04-16 Code/functions.R
likelihood works, optim needs transform

@@ -79,7 +79,7 @@ simFBM <- function(B, alpha, H, N) {
 }
 
 #Blurred Whittle likelihood
-ll <- function(theta, Z, delta, curr) {
+ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero = FALSE) {
   #theta = (A, B, w0,c, h, alpha) 
   #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
@@ -88,9 +88,15 @@ ll <- function(theta, Z, delta, curr) {
   N <- length(Z)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, w0, c, tau, N) + maternAc(B, alpha, h, N)  
-  sBar <- abs(Re(fftshift(2*fft(sTau*(1 - tau/N)) - sTau[1]))) #Fix this, why no negs for sBar?
-  print(-sum(curr$sZ/sBar + log(sBar)))
-  return(-sum(curr$sZ/sBar + log(sBar)))
+  sBar <- abs(Re(fftshift(2*fft(sTau*(1 - tau/N)) - sTau[1]))) #Interpet this, why no negs for sBar?
+  if (incZero == FALSE) {
+    useIndex <- c(firstIndex:(medIndex -1))#MIGHT NEED TO FIX THIS LINE
+    llVal <- -sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
+  } else {
+    llVal <- -sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
+  }
+  print(llVal)
+  return(llVal)
 }
 
 
@@ -108,8 +114,8 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   
   #set window of frequencies to consider in evalution (assuming frequencies sorted min to max)
   medIndex <- floor(N/2) + 1 #middle value (freq = 0)
-  minIndex <- round((medIndex - 1)*(1 - fracNeg) + 1) #index of minimum frequency considered
-  maxIndex <- round(medIndex + fracPos*(N - medIndex)) #index of maximum frequency considered
+  firstIndex <- round((medIndex - 1)*(1 - fracNeg) + 1) #index of minimum frequency considered
[…]

### 2016-04-16 Code/functions.R
transform added, now rescaling

@@ -79,12 +79,19 @@ simFBM <- function(B, alpha, H, N) {
 }
 
 #Blurred Whittle likelihood
-ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero = FALSE) {
+ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero = FALSE, trans = TRUE) {
   #theta = (A, B, w0,c, h, alpha) 
   #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
-  A <- theta[1]; B <- theta[2]; w0 <- theta[3];
-  c <- theta[4]; h <- theta[5]; alpha <- theta[6]
+  if (trans == TRUE) {
+    A <- expm1(theta[1]); B <- expm1(theta[2]); w0 <- theta[3]
+    c <- expm1(theta[4]) + pi*sqrt(3)/N 
+    h <- expm1(theta[5]) + pi*sqrt(3)/N
+    alpha <- expm1(theta[6]) + 0.5
+  } else {
+    A <- theta[1]; B <- theta[2]; w0 <- theta[3];
+    c <- theta[4]; h <- theta[5]; alpha <- theta[6]
+  }
   N <- length(Z)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, w0, c, tau, N) + maternAc(B, alpha, h, N)  
@@ -147,16 +154,25 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   parInit[5] <- sqrt(median((curr$sZ[turbIndex]*(delta*curr$omega[turbIndex])^2)/(max(curr$sZ) - curr$sZ[turbIndex])))
   parInit[2] <- sqrt(max(curr$sZ))*parInit[5]
 
+  #Transform parameters to an unconstrained space for optimization
+  transParInit <- rep(NA, 6)
+  transParInit[1] <- log1p(parInit[1]) #0 < A < inf ===> -inf < log(A) < inf
+  transParInit[2] <- (log1p(parInit[2])) #0 < B < inf ===> -inf < log(B) < inf 
+  transParInit[3] <- parInit[3] #-inf < w0 < inf ===> -inf <- w0 < inf
[…]

### 2016-04-17 Code/functions.R
still need higher tolerance

@@ -98,11 +98,11 @@ ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero =
   sBar <- abs(Re(fftshift(2*fft(sTau*(1 - tau/N)) - sTau[1]))) #Interpet this, why no negs for sBar?
   if (incZero == FALSE) {
     useIndex <- c(firstIndex:(medIndex -1))#MIGHT NEED TO FIX THIS LINE
-    llVal <- -sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
+    llVal <- sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
   } else {
-    llVal <- -sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
+    llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
   }
-  print(llVal)
+  print(log(llVal))
   return(llVal)
 }
 
@@ -153,7 +153,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   turbIndex <- c((medIndex - numTest):(medIndex - 1), (medIndex + 1):(medIndex + numTest))
   parInit[5] <- sqrt(median((curr$sZ[turbIndex]*(delta*curr$omega[turbIndex])^2)/(max(curr$sZ) - curr$sZ[turbIndex])))
   parInit[2] <- sqrt(max(curr$sZ))*parInit[5]
-
+  
   #Transform parameters to an unconstrained space for optimization
   transParInit <- rep(NA, 6)
   transParInit[1] <- log1p(parInit[1]) #0 < A < inf ===> -inf < log(A) < inf
@@ -168,8 +168,8 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
   opt <- optim(transParInit, ll, Z = Z, delta = delta, curr = curr, 
-                firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex,
-                control = list(fnscale = -1, maxiter = 10000000))
[…]

### 2016-04-17 Code/functions.R
ll: -969.0971 need -969.0898

@@ -84,10 +84,10 @@ ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero =
   #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
   if (trans == TRUE) {
-    A <- expm1(theta[1]); B <- expm1(theta[2]); w0 <- theta[3]
-    c <- expm1(theta[4]) + pi*sqrt(3)/N 
-    h <- expm1(theta[5]) + pi*sqrt(3)/N
-    alpha <- expm1(theta[6]) + 0.5
+    A <- 10^(theta[1]); B <- 10^(theta[2]); w0 <- theta[3]
+    c <- 10^(theta[4]) + pi*sqrt(3)/N 
+    h <- 10^(theta[5]) + pi*sqrt(3)/N
+    alpha <- 10^(theta[6]) + 0.5
   } else {
     A <- theta[1]; B <- theta[2]; w0 <- theta[3];
     c <- theta[4]; h <- theta[5]; alpha <- theta[6]
@@ -98,11 +98,11 @@ ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero =
   sBar <- abs(Re(fftshift(2*fft(sTau*(1 - tau/N)) - sTau[1]))) #Interpet this, why no negs for sBar?
   if (incZero == FALSE) {
     useIndex <- c(firstIndex:(medIndex -1))#MIGHT NEED TO FIX THIS LINE
-    llVal <- sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
+    llVal <- -sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
   } else {
-    llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
+    llVal <- -sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
   }
-  print(log(llVal))
+  print(llVal)
   return(llVal)
 }
 
@@ -156,12 +156,12 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   
[…]

### 2016-04-17 Code/functions.R
almost converges enough

@@ -157,7 +157,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   #Transform parameters to an unconstrained space for optimization
   transParInit <- rep(NA, 6)
   transParInit[1] <- log10(parInit[1]) #0 < A < inf ===> -inf < log(A) < inf
-  transParInit[2] <- logb(parInit[2]) #0 < B < inf ===> -inf < log(B) < inf 
+  transParInit[2] <- log10(parInit[2]) #0 < B < inf ===> -inf < log(B) < inf 
   transParInit[3] <- parInit[3] #-inf < w0 < inf ===> -inf <- w0 < inf
   transParInit[4] <- log10(parInit[4] - pi*sqrt(3)/N) #pi*sqrt(3)/N < c < inf ===> -inf < log(c - pi*sqrt(3)/N) < inf
   transParInit[5] <- log10(parInit[5] - pi*sqrt(3)/N) #pi*sqrt(3)/N < c < inf ===> -inf < log(h - pi*sqrt(3)/N) < inf

### 2016-04-17 Code/functions.R
qualitatively working, slight diff in optimized par values:

@@ -25,30 +25,30 @@ getPerio <- function(Z, delta, dB = TRUE, noZero = TRUE) {
 }
 
 #Function to get autocorrelation for the matern
-maternAc <- function(B, alpha, h, N) {
+maternAc <- function(B, alpha, h, N, delta) {
   ac <- rep(NA, N)
   ac[1] <- ((B^2)*beta(0.5, alpha - 0.5))/(2*pi*abs(h)^(2*alpha  - 1)) #variance (cov at lag 0) 
   tau <- 1:(N - 1) #lags 
-  ac[2:N] <- ((B^2*(h*abs(tau))^(alpha - 0.5)*besselK(h*abs(tau), alpha - 0.5))/
-                (2^(alpha - 1/2)*pi^(1/2)*gamma(alpha)*h^(2*alpha  - 1)))
+  ac[2:N] <- ((B^2*(abs(h)*delta*tau)^(alpha - 0.5)*besselK(abs(h)*delta*tau, alpha - 0.5))/
+                (2^(alpha - 1/2)*pi^(1/2)*exp(lgamma(alpha))*abs(h)^(2*alpha  - 1))) 
   return(ac)
 }
 
 #Function to get autocorrelation for the OU process 
-#(TO DO: DETERMINE IF TYPO IS IN THERE CODE OR PAPER 
+#(TO DO: DETERMINE IF TYPO IS IN THEIR CODE OR PAPER 
 #(should there be a delta in this function)
-ouAc <- function(A, w0, c, tau, N) {
+ouAc <- function(A, w0, C, N, delta) {
   ac <- rep(NA, N)
   tau <- 0:(N - 1) #lags 
-  ac <- (0i + A^2/abs(2*c))*exp(1i*w0*tau)*exp(0i + -c*abs(tau))
+  ac <- (0i + A^2/abs(2*C))*exp(1i*w0*delta*tau)*exp(0i + -abs(C)*delta*tau)
   return(ac)
 }
 
 #Function to simulate a time series with complex Matern covariance
-simMatern <- function(B, alpha, h, N) {
+simMatern <- function(B, alpha, h, N, delta) {
   
   #calculate autocorrelation
-  ac <- maternAc(B, alpha, h, N)
+  ac <- maternAc(B, alpha, h, N, delta)
   
   #Create covariance and generate realizations using standard gaussian tricks
   library(mvtnorm)
[…]

### 2016-04-17 Code/figures.R
qualitatively working, slight diff in optimized par values:

@@ -64,7 +64,7 @@ mtext("Replication of Figure 2", outer = T, line = -3)
 #sample matern
 N <- 1000
 DELTA <- 2
-matSamp <- simMatern(B = 10, alpha  = 0.9, h = 0.1, N = N)
+matSamp <- simMatern(B = 10, alpha  = 0.9, h = 0.1, N = N, delta = DELTA)
 mat <- getPerio(Z = matSamp, delta = DELTA)
 
 #sample FBM
@@ -86,9 +86,39 @@ plot(DELTA*fbm$omega, fbm$sZ, col = "blue", ylim = c(-20, 80), xlim = c(-pi, pi)
      type = "l", xlab = expression(paste(omega, Delta)), ylab = "dB")
 points(DELTA*mat$omega, mat$sZ, col = "red", type = "l")
 
-
 ###Figure 4
-optim(par = c(5, 2, 2, 4, 2, 2), ll, Z = bZ, delta = 2,  method = "L-BFGS-B", control=list(fnscale = -1),
-      lower = c(0, -Inf, 0, 0, 0.5, 0), upper = c(Inf, Inf, Inf, Inf, Inf, Inf))
+CF <- mean(-4*pi*(drift$f[2971:3570]))
+#Fit and plot blue time series
+bFit <- fitModel(bZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0)
+par(mfrow = c(1,2))
+plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
+     ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
+sTau <- ouAc(bFit$A, bFit$w0, bFit$C, N, delta = 1) + maternAc(bFit$B, bFit$alpha, bFit$h, N, delta = 1)  
+tau <- seq(0, N - 1)
+sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
+points(DELTA*blue$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd = 3)
+points(DELTA*blue$omega[bFit$firstIndex:bFit$lastIndex], 10*log10(sBar)[bFit$firstIndex:bFit$lastIndex], 
+       col = "green", type = "l", lwd = 3)
+abline(v = CF, lwd = 3) #inertial frequency
+abline(v = bFit$w0, lwd = 3, lty = 3) #estimated w0
[…]

### 2016-04-17 Code/dataConversion.m
qualitatively working, slight diff in optimized par values:

@@ -1,5 +1,9 @@
 %Hannah Director
 %Prelim Exam, file conversion steps
-addpath('/users/hdirector/Dropbox/Prelim/MatlabCode/Figs124')
-load drifterbetty2.mat
-save('drifterbetty2.mat', '-v6')  
+%addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Figs124')
+%load drifterbetty2.mat
+%save('/users/hdirector/Documents/prelim/prelim/Code/drifterbetty2.mat', '-v6') 
+
+addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig5')
+load newinertial.mat
+save('/users/hdirector/Documents/prelim/prelim/Code/newinertial.mat', '-v6')

### 2016-04-19 Code/dataConversion.m
adding outline and paper

@@ -1,8 +1,8 @@
 %Hannah Director
 %Prelim Exam, file conversion steps
-%addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Figs124')
-%load drifterbetty2.mat
-%save('/users/hdirector/Documents/prelim/prelim/Code/drifterbetty2.mat', '-v6') 
+addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Figs124')
+load drifterbetty2.mat
+save('/users/hdirector/Documents/prelim/prelim/Code/drifterbetty2.mat', '-v6') 
 
 addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig5')
 load newinertial.mat

### 2016-04-19 Code/figures.R

@@ -88,11 +88,13 @@ points(DELTA*mat$omega, mat$sZ, col = "red", type = "l")
 
 ###Figure 4
 CF <- mean(-4*pi*(drift$f[2971:3570]))
+N <- length(bZ)
 #Fit and plot blue time series
 bFit <- fitModel(bZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0)
 par(mfrow = c(1,2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
+mtext("Replication of Figure 4", outer = T, line = -3)
 sTau <- ouAc(bFit$A, bFit$w0, bFit$C, N, delta = 1) + maternAc(bFit$B, bFit$alpha, bFit$h, N, delta = 1)  
 tau <- seq(0, N - 1)
 sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
@@ -118,7 +120,53 @@ abline(v = CF, lwd = 3) #inertial frequency
 
 ###Figure 5
 newinertial <- readMat("/users/hdirector/Documents/prelim/prelim/Code/newInertial.mat")
+DELTA <- 1
 num = newinertial$newinertial[,,1]$drifters
+vel <- num[[6]] + 1i*num[[7]] #complex velocities, measured in cm/s
+vel1 <- vel[,1]
 
+#inertial frequency
+lat <- num[[10]] #latitudes
+psi <- mean(lat[,1])*pi/180 #average lat in radians
+f0  <- (8*pi/23.9345)*sin(psi) #still need to figure out where these
+                          #constant come from (physics, units issue)
 
+#left figure (first of 200 simulated drifters)
+sim1Fit <- fitModel(vel1, f0, DELTA, 1.5*f0/pi, 1.5*f0/pi) #periodogram
+sim1Per <- getPerio(vel1, delta = DELTA, dB = TRUE, noZero = FALSE) #fit model
+N <- length(vel1)
+par(mfrow = c(1, 2))
+plot(sim1Per$omega, sim1Per$sZ, type = "l", col = "blue",
+     ylim = c(-40, 60), xlim = c(-pi, pi), ylab = "dB",
+     xlab = expression(paste(omega, Delta)))
[…]

### 2016-04-19 Paper/outline.tex
adding outline and paper

\documentclass{stat572Style}

%%\setlength{\oddsidemargin}{0.25in}
%%\setlength{\textwidth}{6in}
%%\setlength{\topmargin}{0.5in}
%%\setlength{\textheight}{9in}

\renewcommand{\baselinestretch}{1.5}

\bibliographystyle{plainnat}

\begin{document}
%%\maketitle

\begin{center}
  {\LARGE A Review of Lagrangian Time Series Models for Ocean Surface Drifter Trajectories (Sykulski et al. (2016)}\\\ \\
  {Hannah Director \\ 
    Department of Statistics, University of Washington Seattle, WA, 98195, USA
  }
\end{center}

\begin{abstract}
  Put your project summary here.
\end{abstract}

\section{Introduction}
	\subsection{Application}
		\subsection{Data Description}
			Figure 1 included here 
		\subsection{Oceanography background}
			\begin{itemize}
				\item inertial oscillations
				\item turbulent background
			\end{itemize}

	\subsection{Spectral analyses of time series}
			\begin{itemize}
				\item  wave equations
				\item Euler's formula
				\item Fourier transformation
				\item periodogram
				\item relationship between autocovariance and power spectral density
				\item complex-valued velocities
				\item include Figure 2
			\end{itemize}

\section{Methods}

	\subsection{Model}
		\subsubsection{Inertial oscillations}
			\begin{itemize}
				\item Ornstein-Uhlenbeck process 
				\item frequency as a free parameter
				\item include figure 3 
			\end{itemize}

		\subsubsection{Turbulent background}
				\begin{itemize}
					\item Mat\'{e}rn model 
					\item comparison to other integer order processes (e.g. fractional brownian motion)
				\end{itemize}

		\subsubsection{Aggregate model}
			State that you can add two component models together

	\subsection{Model fitting}
		\subsubsection{Whittle likelihood}
			\begin{itemize}
				\item explanation of original Whittle likelihood and its problems (aliasing, leakage)
				\item description of tapering `solution' to Whittle and discussion of its imperfections
				\item blurred whittle likelihood 
				\item allows for uncertainty estimates via asymptotics (Fisher information)
			\end{itemize}

		\subsubsection{Model misspecification}
			\begin{itemize}
				\item semi-parametric approach in both time and frequency
			\end{itemize}
		\subsubsection{Time-varying parameters for non-stationarity}

		\subsubsection{Model selection/likelihood ratio tests}

\section{Results}

	\subsection{Simulated results}
		Include Figure 5

	\subsection{Real drifter data (with time-varying parameters)}
		Include Figures 6-10

\section{Discussion}
	\begin{itemize}
		\item powerful technique overall
		\item more work needed on selecting windows
	\end{itemize}

\section{Appendix}

\subsection{Errata}
	\begin{itemize}
		\item Typo in equation 13
	\end{itemize}

\subsection{Optimization technique}
	\begin{itemize}
		\item My approach transforming parameters to an unconstrained space gives slightly better (higher maximum likelihood) estimates than their use of Matlab's built-in box constraint approach
			\end{itemize}

\bibliography{stat572}

\end{document}

### 2016-04-19 Code/functions.R

@@ -150,7 +150,6 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
   turbIndex <- c((medIndex - numTest):(medIndex - 1), (medIndex + 1):(medIndex + numTest))
   parInit[5] <- sqrt(median((curr$sZ[turbIndex]*(delta*curr$omega[turbIndex])^2)/(max(curr$sZ) - curr$sZ[turbIndex])))
   parInit[2] <- sqrt(max(curr$sZ))*parInit[5]
-
   
   #Transform parameters to an unconstrained space for optimization
   transParInit <- rep(NA, 6)

### 2016-04-19 Paper/paper.tex
spectral backgound info started

\documentclass{stat572Style}
\usepackage{natbib}
\usepackage{amssymb}
%%\setlength{\oddsidemargin}{0.25in}
%%\setlength{\textwidth}{6in}
%%\setlength{\topmargin}{0.5in}
%%\setlength{\textheight}{9in}

\renewcommand{\baselinestretch}{1.5}

\bibliographystyle{plainnat}

\begin{document}
%%\maketitle

\begin{center}
  {\LARGE A Review of Lagrangian Time Series Models for Ocean Surface Drifter Trajectories (Sykulski et al. (2016)}\\\ \\
  {Hannah Director \\ 
    Department of Statistics, University of Washington Seattle, WA, 98195, USA
  }
\end{center}

\begin{abstract}
  Put your project summary here.
\end{abstract}

\section{Introduction}
	\subsection{Application}
		\subsection{Data Description}
			Figure 1 included here 
		\subsection{Oceanography background}
			\begin{itemize}
				\item inertial oscillations
				\item turbulent background
			\end{itemize}

	\subsection{Spectral analyses of time series}
	TO DO: vary who I cite here, unify notation across sources\\
	\indent \citet{Sykulski2016} exclusively employs spectral analysis to model the  drifters.  To provide context for the  methods they propose, this section summarize the main concepts of this time series analysis method. More thorough coverage of this material can be found in  \citep{Percival1993}, for example. From a statistical perspective, a time series, $X_{t} = X(t\Delta)$, where $\Delta > 0 $ is the sampling interval and $t = \{1,2,...,n\}$,  can be understood as a realization of a stochastic process over time. Spectral analysis models time series, not in their original time domain, but instead in a frequency domain. This is possible because time series can be represented as infinite sums of exponential functions, known as a spectral representation. Following the presentation of \citet{Percival1993}, the idea of spectral representation can be understood by initially considering the harmonic process,
    	\begin{equation}
		\label{eq: harmonic}
		X_{t} = \sum_{l=1}^{L}D_{l} cos(2\pi f_{l}t + \phi_{l}),  t = 0, \pm 1, \pm 2...,
	\end{equation}
where $L$ and  $D_{l}$, and $f_{l}$ are real-valued constants, $\phi_{l}$ are independent Uniform($-\pi, \pi$) random variables, and the frequencies, $0 <f_{l} < 1/2$, are taken to be increasing order. As the number of  sinusoidal functions increases,  more complex behavior can be represented by the sum in Equation ~\ref{eq: harmonic}.  Applying Euler's formula,  Equation \ref{eq: harmonic}, can be written as 
	\begin{equation}
		\label{eq: harmonicSpec}
		X_{t} = \sum_{l = -L}^{L} C_{l}e^{i2\pi f_{l} t}.
	\end{equation}

\noindent which as $L \rightarrow \infty$  gives
	\begin{equation}
		\label{eq: specRep}
		X_{t} = \int_{-1/2}^{1/2} e^{i2\pi ft}dZ(f),
	\end{equation}

\noindent Equation \ref{eq: specRep} can be used to define any real or complex-valued discrete parameter stationary process, $X_{t}$, with mean zero. Specifically, there exists a process $\{Z(f)\}$ on $[-1/2, 1/2]$,  where the intervals $dZ(f)$ and $dZ(f')$  are uncorrelated for all $f,f'$, that satisfies Equation \ref{eq: specRep} for all $t$\citep{Percival1993}. This result is known as the spectral representation theorem for discrete parameter stationary processes. An analogous result holds for continuous parameter stationary processes \citep{Percival1993}.

The relationship between spectral densities and autocovariances also contributes to the utility of analysis in this domain. In particular, the spectral densities of a time series and the autocovariance sequence are the Fourier transform of one another.  In this paper, the focus is on second order stationary processes,  or cases where the autocovariance, $s_{xx}$, is only a function of the distance between the increments,  meaning $c_{xx}(t_{1}, t_{2}) = \mathbb{E}(X_{t1}X_{t2}) = s_{xx}(|t_{1} - t_{2}|)$.  (add derivation here?) In such cases, this relationship is written as 
	\begin{equation}
	S_{xx}(\omega) = \Delta \sum_{\tau = -\infty}^{\infty} s_{xx}(\tau)e^{i\omega \tau \Delta} 
	\end{equation}
and 
	\begin{equation}
		s_{xx}(\tau) = \frac{1}{2\pi} \int_{-\pi / \Delta}^{\pi / \Delta}S_{xx}(\omega)e^{i \omega \Delta}d \omega.
	\end{equation}
\noindent where $\omega \in [-\pi/\Delta, \pi/ \Delta]$ is the angular frequency measured in radians \citep{Sykulski2013}
.  In the complex-valued case, $s_{xx}$ and $S_{xx}$ are replaced with $s_{xy}$ and $S_{xy}$ to represent the cross-covariance terms \citep{Sykulski2013}.

Given an observed time series $X_{t}$, $t = 1,2,..,N$, the common estimator for the spectra is the periodogram, $\hat{S}_{xx}(\omega) = |J_{x}(\omega)|^{2}$ where 
\begin{equation}
J_{x}(\omega) = \sqrt{\frac{\Delta}{N}} \sum_{t=1}^{N} X_{t}e^{-i \omega t \Delta}. 
\end{equation}
Although simple, this estimator has two known issues: aliasing and leakage.  Aliasing is ... Leakage is... Tapering has been proposed to fix, but has problems..

\section{Methods}

	\subsection{Model}
		\subsubsection{Inertial oscillations}
			\begin{itemize}
				\item Ornstein-Uhlenbeck process 
				\item frequency as a free parameter
				\item include figure 3 
			\end{itemize}

		\subsubsection{Turbulent background}
				\begin{itemize}
					\item Mat\'{e}rn model 
					\item comparison to other integer order processes (e.g. fractional brownian motion)
				\end{itemize}

		\subsubsection{Aggregate model}
			State that you can add two component models together

	\subsection{Model fitting}
		\subsubsection{Whittle likelihood}

			\begin{itemize}
				\item explanation of original Whittle likelihood and its problems (aliasing, leakage)
				\item description of tapering `solution' to Whittle and discussion of its imperfections
				\item blurred whittle likelihood 
				\item allows for uncertainty estimates via asymptotics (Fisher information)
			\end{itemize}

		\subsubsection{Model misspecification}
			\begin{itemize}
				\item semi-parametric approach in both time and frequency
			\end{itemize}
		\subsubsection{Time-varying parameters for non-stationarity}

		\subsubsection{Model selection/likelihood ratio tests}

\section{Results}

	\subsection{Simulated results}
		Include Figure 5

	\subsection{Real drifter data (with time-varying parameters)}
		Include Figures 6-10

\section{Discussion}
	\begin{itemize}
		\item powerful technique overall
		\item more work needed on selecting windows
	\end{itemize}

\section{Appendix}

\subsection{Errata}
	\begin{itemize}
		\item Typo in equation 13
	\end{itemize}

\subsection{Optimization technique}
	\begin{itemize}
		\item My approach transforming parameters to an unconstrained space gives slightly better (higher maximum likelihood) estimates than their use of Matlab's built-in box constraint approach
			\end{itemize}

\bibliography{prelim}

\end{document}

### 2016-04-23 Paper/paper.tex
ODE to OU derivation added

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 %%\setlength{\oddsidemargin}{0.25in}
@@
 			Figure 1 included here
-		\subsection{Oceanography background}
-			\begin{itemize}
-				\item inertial oscillations
-				\item turbulent background
-			\end{itemize}
 
@@
 \subsubsection{Inertial oscillations}
-			\begin{itemize}
-				\item Ornstein-Uhlenbeck process 
-				\item frequency as a free parameter
-				\item include figure 3 
-			\end{itemize}
+
+Inertial oscillations are often modeled with the following set of deterministic differential equations (read and add physics citations)
+\begin{align}
+\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
+\frac{\partial v}{\partial t} - f_{0}u &= G - cv
+\end{align}
+where $u$ and $v$ represent the vertical and horizontal velocity, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
+\citet{Sykulski2016} propose instead using a a complex-valued representation of the velocity, $z(t) = u(t) + i v(t)$,  from which we obtain
+\begin{align*}
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
+dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
+&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
+&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
+dz(t) &= (if_{0} - c)z dt + (F + iG)dt. 
+\end{align*}
+We further replace the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, which is justified by blah (look up and cite the Mandelrot and Van Ness justification).
+Thus, we have a stochastic model which retains the desired relationship between the horizontal and vertical velocity components.
+\citet{Sykulski2016} further let $f_{0}$ be a free parameter, $\omega_{0}$, giving the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
+\begin{equation}
+dz(t) = (i \omega -c) z(t) dt + A d Q(t), 
+\end{equation}
+Letting $\omega_{0}$ be estimated by data, rather than fixed at the inertial frequency allows for identification and estimation of the shifts in the angular frequency due to eddies.
+If the observed angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
+
+The complex OU process has
 

### 2016-04-29 Code/dataConversion.m
background and motivation done

@@ -7,3 +7,7 @@ save('/users/hdirector/Documents/prelim/prelim/Code/drifterbetty2.mat', '-v6')
 addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig5')
 load newinertial.mat
 save('/users/hdirector/Documents/prelim/prelim/Code/newinertial.mat', '-v6')
+
+addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig6-10')
+load drifterulysses.mat
+save('/users/hdirector/Documents/prelim/prelim/Code/drifterulysses.mat', '-v6')

### 2016-04-29 Code/functions.R

@@ -79,7 +79,7 @@ simFBM <- function(B, alpha, H, N) {
 }
 
 #Blurred Whittle likelihood
-ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero = FALSE, trans = TRUE) {
+ll <- function(theta,delta, curr, firstIndex, lastIndex, medIndex, incZero = FALSE, trans = TRUE) {
   #theta = (A, B, w0,c, h, alpha) 
   #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
@@ -97,30 +97,31 @@ ll <- function(theta, Z, delta, curr, firstIndex, lastIndex, medIndex, incZero =
   if (alpha > 171) {
     print("alpha")
   }
-  N <- length(Z)
+  N <- length(curr$sZ)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, w0, C, N, delta = 1) + maternAc(B, alpha, h, N, delta = 1)  
   sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) #Interpet this, why no negs for sBar?
-  if (incZero == FALSE) {
-    useIndex <- c(firstIndex:(medIndex - 1)) #check this line
-    llVal <- sum(curr$sZ[useIndex]/sBar[useIndex] + log(sBar[useIndex]))
-  } else {
-    llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
-  }
+  llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
   print(llVal)
   return(100*log(llVal))
 }
 
 #Function to fit likelihood
-fitModel <- function(Z, CF, delta, fracNeg, fracPos) {
-  curr <- getPerio(Z, delta, dB = FALSE, noZero = FALSE)
-  N <- length(Z)
+fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet,
+                     ensem = FALSE, curr = NULL) {
+  
+  #calc. periodogram in function, unless it's an ensemble mean
+  if (ensem == FALSE ) {
+    curr <- getPerio(Z, delta, dB = FALSE, noZero = TRUE)
[…]

### 2016-04-29 Code/figures.R
background and motivation done

@@ -29,8 +29,6 @@ points(drift$lon[gStartTs:gEndTs], drift$lat[gStartTs:gEndTs], type = "l", col =
 points(drift$lon[bStartTs:bEndTs], drift$lat[bStartTs:bEndTs], type = "l", col = "blue")
 
 ###Figure 2###
-#To do: figure out how to get complex valued velocities (e.g. units)
-#Calculate velocities u(t) (horizontal) and v(t) (vertical)
 
 #pull out complex valued velocities
 bZ <- drift$cv[bStartTs:bEndTs]
@@ -90,7 +88,7 @@ points(DELTA*mat$omega, mat$sZ, col = "red", type = "l")
 CF <- mean(-4*pi*(drift$f[2971:3570]))
 N <- length(bZ)
 #Fit and plot blue time series
-bFit <- fitModel(bZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0)
+bFit <- fitModel(bZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0, quantSet = .5)
 par(mfrow = c(1,2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
@@ -107,7 +105,7 @@ abline(v = CF - bFit$w0, lwd = 3, lty = 3) #shifted inertial freq (f = f0 - w_ed
 
 #Fit and plot green time series
 #abline(v = f0, col = "red")
-gFit <- fitModel(gZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0)
+gFit <- fitModel(gZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0,  quantSet = 0.5)
 plot(DELTA*green$omega, green$sZ, type = "l", col = "green", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
 sTau <- ouAc(gFit$A, gFit$w0, gFit$C, N, delta = 1) + maternAc(gFit$B, gFit$alpha, gFit$h, N, delta = 1)  
@@ -128,12 +126,12 @@ vel1 <- vel[,1]
 #inertial frequency
 lat <- num[[10]] #latitudes
 psi <- mean(lat[,1])*pi/180 #average lat in radians
[…]

### 2016-04-29 Paper/paper.tex
background and motivation done

--- 2016-04-19 Paper/paper.tex
@@
 \begin{center}
-  {\LARGE A Review of Lagrangian Time Series Models for Ocean Surface Drifter Trajectories (Sykulski et al. (2016)}\\\ \\
+  {\LARGE A Review of Lagrangian Time Series Models for Ocean Surface Drifter Trajectories (Sykulski et al. (2016))}\\\ \\
   {Hannah Director \\ 
@@
 
-\begin{abstract}
-  Put your project summary here.
-\end{abstract}
+%\begin{abstract}
+%  Put your project summary here.
+%\end{abstract}
 
 \section{Introduction}
-	\subsection{Application}
-		\subsection{Data Description}
-			Figure 1 included here 
-		\subsection{Oceanography background}
-			\begin{itemize}
-				\item inertial oscillations
-				\item turbulent background
-			\end{itemize}
-
-	\subsection{Spectral analyses of time series}
-	TO DO: vary who I cite here, unify notation across sources\\
-	\indent \citet{Sykulski2016} exclusively employs spectral analysis to model the  drifters.
-To provide context for the  methods they propose, this section summarize the main concepts of this time series analysis method.
-More thorough coverage of this material can be found in  \citep{Percival1993}, for example.
-From a statistical perspective, a time series, $X_{t} = X(t\Delta)$, where $\Delta > 0 $ is the sampling interval and $t = \{1,2,...,n\}$,  can be understood as a realization of a stochastic process over time.
-Spectral analysis models time series, not in their original time domain, but instead in a frequency domain.
-This is possible because time series can be represented as infinite sums of exponential functions, known as a spectral representation.
-Following the presentation of \citet{Percival1993}, the idea of spectral representation can be understood by initially considering the harmonic process,
-    	\begin{equation}
-		\label{eq: harmonic}
-		X_{t} = \sum_{l=1}^{L}D_{l} cos(2\pi f_{l}t + \phi_{l}),  t = 0, \pm 1, \pm 2...,
-	\end{equation}
-where $L$ and  $D_{l}$, and $f_{l}$ are real-valued constants, $\phi_{l}$ are independent Uniform($-\pi, \pi$) random variables, and the frequencies, $0 <f_{l} < 1/2$, are taken to be increasing order.
-As the number of  sinusoidal functions increases,  more complex behavior can be represented by the sum in Equation ~\ref{eq: harmonic}.
-Applying Euler's formula,  Equation \ref{eq: harmonic}, can be written as
-	\begin{equation}
-		\label{eq: harmonicSpec}
-		X_{t} = \sum_{l = -L}^{L} C_{l}e^{i2\pi f_{l} t}.
-	\end{equation}
-
-\noindent which as $L \rightarrow \infty$  gives
-	\begin{equation}
-		\label{eq: specRep}
-		X_{t} = \int_{-1/2}^{1/2} e^{i2\pi ft}dZ(f),
-	\end{equation}
-
-\noindent Equation \ref{eq: specRep} can be used to define any real or complex-valued discrete parameter stationary process, $X_{t}$, with mean zero.
-Specifically, there exists a process $\{Z(f)\}$ on $[-1/2, 1/2]$,  where the intervals $dZ(f)$ and $dZ(f')$  are uncorrelated for all $f,f'$, that satisfies Equation \ref{eq: specRep} for all $t$\citep{Percival1993}.
-This result is known as the spectral representation theorem for discrete parameter stationary processes.
-An analogous result holds for continuous parameter stationary processes \citep{Percival1993}.
-
-The relationship between spectral densities and autocovariances also contributes to the utility of analysis in this domain.
-In particular, the spectral densities of a time series and the autocovariance sequence are the Fourier transform of one another.
-In this paper, the focus is on second order stationary processes,  or cases where the autocovariance, $s_{xx}$, is only a function of the distance between the increments,  meaning $c_{xx}(t_{1}, t_{2}) = \mathbb{E}(X_{t1}X_{t2}) = s_{xx}(|t_{1} - t_{2}|)$.
-(add derivation here?) In such cases, this relationship is written as
-	\begin{equation}
-	S_{xx}(\omega) = \Delta \sum_{\tau = -\infty}^{\infty} s_{xx}(\tau)e^{i\omega \tau \Delta} 
-	\end{equation}
-and 
-	\begin{equation}
-		s_{xx}(\tau) = \frac{1}{2\pi} \int_{-\pi / \Delta}^{\pi / \Delta}S_{xx}(\omega)e^{i \omega \Delta}d \omega.
-	\end{equation}
-\noindent where $\omega \in [-\pi/\Delta, \pi/ \Delta]$ is the angular frequency measured in radians \citep{Sykulski2013}
-.  In the complex-valued case, $s_{xx}$ and $S_{xx}$ are replaced with $s_{xy}$ and $S_{xy}$ to represent the cross-covariance terms \citep{Sykulski2013}.
-
-Given an observed time series $X_{t}$, $t = 1,2,..,N$, the common estimator for the spectra is the periodogram, $\hat{S}_{xx}(\omega) = |J_{x}(\omega)|^{2}$ where 
-\begin{equation}
-J_{x}(\omega) = \sqrt{\frac{\Delta}{N}} \sum_{t=1}^{N} X_{t}e^{-i \omega t \Delta}. 
-\end{equation}
-Although simple, this estimator has two known issues: aliasing and leakage.  Aliasing is ... Leakage is... Tapering has been proposed to fix, but has problems..
-
+\indent ``Lagrangian Time Series Models for Ocean Surface Drifter Trajectories," by Adam M.
+Sykulski, Sofia C.
+Olhede, Jonathan M.
+Lilly, and Eric Danioux presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
+Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time.
+Oceanographers use these time series to increase understanding of  ocean circulation patterns.
+Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.\\
+\indent The motion of parcels of water moving over changing latitudes is known to have a rotational component.
+Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.
+This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects being observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude.
+Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations.
+These motions have known frequencies, referred to as inertial frequencies.
+Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
+However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
+The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
+\\
+\indent In achieving this result, several statistical advances are made.
+The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations.
+For the turbulent background model, the Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is extended to the time-series context.
+This allows for a more flexible model for background behavior than previously obtained.
+Inertial oscillations are modeled with a complex Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
+Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
+\\
+\indent The remainder of this review proceeds as follows.
+After a succint review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
+Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model.
+The paper concludes with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
 \section{Methods}
 
-	\subsection{Model}
-		\subsubsection{Inertial oscillations}
-			\begin{itemize}
-				\item Ornstein-Uhlenbeck process 
-				\item frequency as a free parameter
-				\item include figure 3 
-			\end{itemize}
+	\subsection{Spectral Analysis}
+	\indent We briefly review spectral analysis as these methods are central to understanding \citet{Sykulski2016}.
+More thorough coverage of this material can be found in  \citep{Percival1993}, for example.
+From a statistical perspective, a time series, $x(t)$ can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
+Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series.
+Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations 	\begin{align}
+x(t) \overset{ms}{=} \int_{-\infty}^{\infty} f_{x}(\omega)e^{i\omega t}d\omega && f_{x}(\omega) = \int_{-\infty}^{\infty} x(t) e^{-i \omega t }dt.
+\end{align}
+\citep{Percival1993}. To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
 
-		\subsubsection{Turbulent background}
-				\begin{itemize}
-					\item Mat\'{e}rn model 
-					\item comparison to other integer order processes (e.g. fractional brownian motion)
-				\end{itemize}
+\begin{equation}
+S_{x}(\omega) = \underset{T \rightarrow \infty}{lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} x(t) e^{-i \omega t}dt \right|^{2} \right),
+\end{equation}
+is often employed.
+This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}.
+Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.
+Defining the autocovariance as $s_{x}(\tau) = \mathbb{E}[x(t) x(t - \tau)] $ where $\tau$ is the time lag, we have
 
-		\subsubsection{Aggregate model}
-			State that you can add two component models together
+\begin{align*}
+S_{x}(\omega) = \int_{-\infty}^{\infty}s_{x}(\tau) e^{-\omega t}dt  \Longleftrightarrow \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{x}(\omega) e^{i \omega t} d\omega 
+\end{align*}
 
-	\subsection{Model fitting}
-		\subsubsection{Whittle likelihood}
+\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, $\hat{S}_{x}(\omega) = |J_{x}(\omega)|^{2}$ where 
+\begin{equation}
+J_{x}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} x(t) e^{-i \omega t}
+\end{equation}
+\citep{Sykulski2013}.
+Although intuitively appealing, this estimator has known problems that stem from the sampling of a periodic function at discrete intervals.
+In particular,  high frequencies can  capture periodic behavior of other lower frequencies if that happen to align in period.
+Additionally, frequencies above the highest observable frequency cannot be captured, and, are instead, recorded as other frequencies in the spectrum.
+These effects, referred to as $\textbf{\it{aliasing}}$ and $\textbf{\it{leakage}}$ respectively,  are typically reduced by $\textbf{\it{tapering}}$.
+This method uses  a window of frequencies for estimation rather than a single value.
+However, depending on how tapering is used, it can introduce its own  biases in estimation.
 
-			\begin{itemize}
-				\item explanation of original Whittle likelihood and its problems (aliasing, leakage)
-				\item description of tapering `solution' to Whittle and discussion of its imperfections
-				\item blurred whittle likelihood 
-				\item allows for uncertainty estimates via asymptotics (Fisher information)
-			\end{itemize}
+%\subsection{Model}
+%\subsubsection{Inertial oscillations}
 
-		\subsubsection{Model misspecification}
-			\begin{itemize}
-				\item semi-parametric approach in both time and frequency
-			\end{itemize}
-		\subsubsection{Time-varying parameters for non-stationarity}
+%Inertial oscillations are often modeled with the following set of deterministic differential equations \citep{Pollard1970}
+%\begin{align}
+%\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
+%\frac{\partial v}{\partial t} - f_{0}u &= G - cv
+%\end{align}
+%where $u$ and $v$ represent the vertical and horizontal velocity, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
+\citet{Sykulski2016} propose instead using a a complex-valued representation of the velocity, $z(t) = u(t) + i v(t)$,  from which we obtain
+%\begin{align*}
+%\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
+%dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
+%&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
+%&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
+%dz(t) &= (if_{0} - c)z dt + (F + iG)dt. 
+%\end{align*}
+%We further replace the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, which is justified by blah (look up and cite the Mandelrot and Van Ness justification).
+Thus, we have a stochastic model which retains the desired relationship between the horizontal and vertical velocity components.
+\citet{Sykulski2016} further let $f_{0}$ be a free parameter, $\omega_{0}$, giving the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
+%\begin{equation}
+%dz(t) = (i \omega -c) z(t) dt + A d Q(t), 
+%\end{equation}
+%Letting $\omega_{0}$ be estimated by data, rather than fixed at the inertial frequency allows for identification and estimation of the shifts in the angular frequency due to eddies.
+If the observed angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
 
-		\subsubsection{Model selection/likelihood ratio tests}
-
-\section{Results}
-
-	\subsection{Simulated results}
-		Include Figure 5
-
-	\subsection{Real drifter data (with time-varying parameters)}
-		Include Figures 6-10
-
-\section{Discussion}
-	\begin{itemize}
-		\item powerful technique overall
-		\item more work needed on selecting windows
-	\end{itemize}
-
-\section{Appendix}
-
-\subsection{Errata}
-	\begin{itemize}
-		\item Typo in equation 13
-	\end{itemize}
-
-\subsection{Optimization technique}
-	\begin{itemize}
-		\item My approach transforming parameters to an unconstrained space gives slightly better (higher maximum likelihood) estimates than their use of Matlab's built-in box constraint approach
-			\end{itemize}
+%The complex OU process has
 

### 2016-05-01 Code/functions.R
rolling windows and CI added

@@ -29,18 +29,16 @@ maternAc <- function(B, alpha, h, N, delta) {
   ac <- rep(NA, N)
   ac[1] <- ((B^2)*beta(0.5, alpha - 0.5))/(2*pi*abs(h)^(2*alpha  - 1)) #variance (cov at lag 0) 
   tau <- 1:(N - 1) #lags 
-  ac[2:N] <- ((B^2*(abs(h)*delta*tau)^(alpha - 0.5)*besselK(abs(h)*delta*tau, alpha - 0.5))/
+  ac[2:N] <- ((B^2*(abs(h)*tau)^(alpha - 0.5)*besselK(abs(h)*tau, alpha - 0.5))/
                 (2^(alpha - 1/2)*pi^(1/2)*exp(lgamma(alpha))*abs(h)^(2*alpha  - 1))) 
   return(ac)
 }
 
 #Function to get autocorrelation for the OU process 
-#(TO DO: DETERMINE IF TYPO IS IN THEIR CODE OR PAPER 
-#(should there be a delta in this function)
 ouAc <- function(A, w0, C, N, delta) {
   ac <- rep(NA, N)
   tau <- 0:(N - 1) #lags 
-  ac <- (0i + A^2/abs(2*C))*exp(1i*w0*delta*tau)*exp(0i + -abs(C)*delta*tau)
+  ac <- (0i + A^2/abs(2*C))*exp(1i*w0*tau)*exp(0i + -abs(C)*tau)
   return(ac)
 }
 
@@ -93,28 +91,58 @@ ll <- function(theta,delta, curr, firstIndex, lastIndex, medIndex, incZero = FAL
     C <- theta[4]; h <- theta[5]; alpha <- theta[6]
   }
   
-  #value is too high and will crash
-  if (alpha > 171) {
-    print("alpha")
-  }
   N <- length(curr$sZ)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, w0, C, N, delta = 1) + maternAc(B, alpha, h, N, delta = 1)  
   sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) #Interpet this, why no negs for sBar?
-  llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex] + log(sBar[firstIndex:lastIndex]))
-  print(llVal)
+  if (incZero == TRUE) {
[…]

### 2016-05-01 Code/figures.R
rolling windows and CI added

@@ -93,7 +93,7 @@ par(mfrow = c(1,2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
 mtext("Replication of Figure 4", outer = T, line = -3)
-sTau <- ouAc(bFit$A, bFit$w0, bFit$C, N, delta = 1) + maternAc(bFit$B, bFit$alpha, bFit$h, N, delta = 1)  
+sTau <- ouAc(bFit$A, bFit$w0, bFit$C, N) + maternAc(bFit$B, bFit$alpha, bFit$h, N)  
 tau <- seq(0, N - 1)
 sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
 points(DELTA*blue$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd = 3)
@@ -108,7 +108,7 @@ abline(v = CF - bFit$w0, lwd = 3, lty = 3) #shifted inertial freq (f = f0 - w_ed
 gFit <- fitModel(gZ, CF, delta = 2, fracNeg = 0.4, fracPos = 0,  quantSet = 0.5)
 plot(DELTA*green$omega, green$sZ, type = "l", col = "green", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
-sTau <- ouAc(gFit$A, gFit$w0, gFit$C, N, delta = 1) + maternAc(gFit$B, gFit$alpha, gFit$h, N, delta = 1)  
+sTau <- ouAc(gFit$A, gFit$w0, gFit$C, N) + maternAc(gFit$B, gFit$alpha, gFit$h, N)  
 tau <- seq(0, N - 1)
 sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
 points(DELTA*green$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd = 3)
@@ -139,59 +139,68 @@ plot(sim1Per$omega, sim1Per$sZ, type = "l", col = "blue",
      xlab = expression(paste(omega, Delta)))
 mtext("Replication of Figure 5", outer = T, line = -3)
[…]

### 2016-05-08 Code/figures.R
Paper/stat572Style.cls

@@ -144,7 +144,7 @@ f0  <- -(8*pi/23.9345)*sin(psi) #still need to figure out where these
 sim1Fit <- fitModel(vel1, CF = f0, DELTA, 1.5*f0/pi, 1.5*f0/pi, quantSet = .9)  #fit model
 sim1Per <- getPerio(vel1, delta = DELTA, dB = TRUE, noZero = FALSE) #periodogram
 N <- length(vel1)
-png('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig5.png')
+#png('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig5.png')
 par(mfrow = c(1, 2))
 plot(sim1Per$omega, sim1Per$sZ, type = "l", col = "blue",
      ylim = c(-40, 60), xlim = c(-pi, pi), ylab = "dB",
@@ -208,7 +208,7 @@ points(DELTA*sim1Per$omega, meanEnsemSz, col = "blue", type = "l")
 points(DELTA*sim1Per$omega, meanEnsemSBar, col = "red", lty = 3, type = "l", lwd = 3)
 points(DELTA*sim1Per$omega[sim1Fit$firstIndex:sim1Fit$lastIndex],
        meanEnsemSBar[sim1Fit$firstIndex:sim1Fit$lastIndex], col = "green", type = "l", lwd = 3)
-dev.off()
+#dev.off()
 
 ###Figure 6
 #12 measurements per day
@@ -220,7 +220,7 @@ driftUlys <- list("num" = as.vector(drifterulysses$drifterulysses[,,1]$num),
                   "f" = as.vector(drifterulysses$drifterulysses[,,1]$f))
 DELTA <- 2
 
-pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig6.pdf')
+#pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig6.pdf')
 par(mfrow = c(1, 2))
 plot(driftUlys$lon, driftUlys$lat, type = "l", col = "blue", xlim = c(-120, -80),
      ylim = c(-40, -20), xlab = "Longitude", ylab = "Latitude") 
@@ -244,7 +244,7 @@ points(DELTA*samp$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd =
 points(DELTA*samp$omega[sampFit$firstIndex:sampFit$lastIndex], 10*log10(sBar)[sampFit$firstIndex:sampFit$lastIndex], 
[…]

### 2016-05-08 Code/functions.R
Paper/stat572Style.cls

@@ -105,15 +105,14 @@ ll <- function(theta,delta, curr, firstIndex, lastIndex, medIndex,
   return(100*log(llVal))
 }
 
-
-#Simplified Blurred Whittle likelihood (5 parameter model)
+#Simplified Blurred Whittle likelihood (fixed CF)
 llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
-               incZero = FALSE, trans = TRUE, CF) {
-  #theta = (A, B, w0,c, h, alpha) 
-  #A > 0: ou amplitude, B > 0: matern amplitude; 
+                   CF, incZero = FALSE, trans = TRUE) {
+  #theta = (A, B, c, h, alpha) 
+  #A > 0: ou amplitude, B > 0: matern amplitude;
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
   if (trans == TRUE) {
-    A <- expm1(theta[1]) + 1; B <- expm1(theta[2]) + 1
+    A <- expm1(theta[1]) + 1; B <- expm1(theta[2]) + 1; 
     C <- expm1(theta[3]) + pi*sqrt(3)/N  + 1
     h <- expm1(theta[4]) + pi*sqrt(3)/N + 1
     alpha <- expm1(theta[5]) + 0.5 + 1
@@ -135,6 +134,7 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
   return(100*log(llVal))
 }
 
+
 #Function to fit likelihood
 fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
                      needInits = TRUE, parInit = NULL, hess = FALSE, simpModel = FALSE) {
@@ -153,7 +153,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
   if (needInits == TRUE) {
     ####initial parameter estimates (using the well-reasoned parameters suggested in the paper's published code)
     #theta = (A, B, w0,c, h, alpha) 
-    #A > 0: ou amplitude, B > 0: matern amplitude; w0: ou frequency, 
+    #A > 0: ou amplitude, B > 0: matern amplitude; 
     #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37) 
     parInit <- rep(NA, 6) 
[…]

### 2016-05-08 Paper/paper.tex

@@
 \begin{equation}
-S_{x}(\omega) = \underset{T \rightarrow \infty}{lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} x(t) e^{-i \omega t}dt \right|^{2} \right),
+S_{x}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} x(t) e^{-i \omega t}dt \right|^{2} \right),
 \end{equation}
@@
 
-%\subsection{Model}
-%\subsubsection{Inertial oscillations}
+\subsection{Model}
+ To model the drifter's motion over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
+Stochastic models are then developed for both components of the the drifter's behavior, inertial oscillations and movement due to the ocean's turbulent background,  in the spectral domain.
+Finally, the two components of the model are added together, creating a six-parameter model, which is fit via maximizing a blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
 
-%Inertial oscillations are often modeled with the following set of deterministic differential equations \citep{Pollard1970}
-%\begin{align}
-%\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
-%\frac{\partial v}{\partial t} - f_{0}u &= G - cv
-%\end{align}
-%where $u$ and $v$ represent the vertical and horizontal velocity, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
-\citet{Sykulski2016} propose instead using a a complex-valued representation of the velocity, $z(t) = u(t) + i v(t)$,  from which we obtain
-%\begin{align*}
-%\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
-%dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
-%&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
-%&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
-%dz(t) &= (if_{0} - c)z dt + (F + iG)dt. 
-%\end{align*}
-%We further replace the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, which is justified by blah (look up and cite the Mandelrot and Van Ness justification).
-Thus, we have a stochastic model which retains the desired relationship between the horizontal and vertical velocity components.
-\citet{Sykulski2016} further let $f_{0}$ be a free parameter, $\omega_{0}$, giving the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
-%\begin{equation}
-%dz(t) = (i \omega -c) z(t) dt + A d Q(t), 
-%\end{equation}
-%Letting $\omega_{0}$ be estimated by data, rather than fixed at the inertial frequency allows for identification and estimation of the shifts in the angular frequency due to eddies.
-If the observed angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
+\subsubsection{Inertial oscillations}
+We discuss modeling just the inertial oscillations first.
+In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations \citep{Pollard1970}
+\begin{align}
+\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
+\frac{\partial v}{\partial t} - f_{0}u &= G - cv
+\end{align}
+where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
+Suppressing the dependence on $t$ and using the complex representation of the velocity, these relationships can be equivalently expressed as follows
+\begin{align*}
+\label{eq: diffEqDeriv}
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
+dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
+&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
+&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
+&= (if_{0} - c)z dt + (F + iG)dt. 
+\end{align*}
+By further replacing the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, \citet{Sykulski2016} obtain the stochastic model 
+\begin{equation}
+\label{eq: ouEq}
+dz(t) = (i f_{0} -c) z(t) dt + A d Q(t), 
+\end{equation}
+which defines equivalent behavior to the accepted deterministic formulation.
+Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
+It developed from brownian motion to OU to complex OU, mean-reversion, dampening and is a valid in this context by  blah (look up and cite the Mandelrot and Van Ness justification) giving the This model is always Markovian.
+The OU model is stationary assuming the initial conditions of the stochastic differential equation are set as $z(t_{0}) \sim N(0, A^{2}/2|c|)$, and, if not, will revert to the stationary model quickly as the influence of the initial conditions decays exponentially (WHY?
+look this up).
 
-%The complex OU process has
+\citet{Sykulski2016} further let $f_{0}$ be a free parameter, denoted $\omega_{0}$, to be estimated by data.
+In so doing, they enable  identification and estimation of the shifts in the angular frequency due to eddies.
+If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
+
+Letting $z^{*}(t)$ represent the complex conjugate of z(t) and using the previously-specified initial conditions, the complex OU process has the following  autocovariance  and power spectral density
+\begin{align}
+\label{eq: ouAC}
+s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)\\
+\label{eq: ouPSD}
+S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}
+\end{align}
+where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the model.
+In addition to the inertial frequency, $\omega_{0}$, the parameter, $A > 0$ that defines the variability, and the parameter, $c >0$ which controls the dampening,  need to be fit from data.
+\subsubsection{Turbulent Background}
+The drifters' time series are also marked by large-scale turbulence (read and cite Rhines 1979).
+Many potential models have been considered for this background....Lit review here ...
+However, these models have been \textbf{\it{integer order}}, meaning the velocity, acceleration, and hyperacceleration must be Markovian, a constraint previous studies (READ AND CITE) indicate is not appropriate in all cases.
+Thus, the authors propose using the Mat\'{e}rn covariance structure \citep{Gneiting2012}, a model which has gained prominence in the spatial statistics literature due to its flexibility.
+In this context, the Mat\'{e}rn model usefully  allows for markovianity when the data supports it while not universally enforcing it.
+The corresponding autocovariance and power spectral density are (read and cite stein 1991 here)
+
+\begin{align}
+\label{eq: maternAC}
+s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
+\label{eq: maternPSD}
+S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
+\end{align}
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$.
+and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model.
+The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
+
+\subsubsection{Overall Model}
+Combining the inertial oscillation and turbulent background parts of the model is straightforward.
+For both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background.
+Using Equations ~\ref{eq: ouAC} and ~\ref{eq: maternAC} and Equations ~\ref{eq: ouPSD} and ~\ref{eq: maternPSD}, we obtain,
+\begin{align}
+S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
+s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|),
+\end{align}
+ the autocovariance and power spectral density respectively.
 

### 2016-05-08 Code/figures.R
added figure 9

@@ -367,7 +367,7 @@ plot((500:(N - 499))/12, par6Val[, "alpha"], ylim = c(0.58, 1.45), type = "l",
 #dev.off()
 
 
-###Figure 9
+###Figure 9 ####
 #Fraction positive
 fracPos <- 1.75*max(4*pi*driftUlys$f)/pi 
 
@@ -400,8 +400,7 @@ par5Val[1, ] <- c(toStartFit5$A, toStartFit5$B, toStartFit5$C,
 
 #save(llVal5, file = '/users/hdirector/Documents/prelim/prelim/Code/llVal5.rda')
 
-
-#plot LRT
+plot LRT 
 pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig9.pdf')
 par(mfrow = c(1, 1))
 #Note that we have stored the negative of the log likelihoods

### 2016-05-08 Code/functions.R
added figure 9

@@ -176,7 +176,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
     parInit[1] <- curr$sZ[maxIndex]*parInit[4]
     parInit[1] <- sqrt(parInit[1]); parInit[4] <- sqrt(parInit[4])
     
-    #Find initial guess for Matern amplitude and slope; zero in on area around turbulent background peak
+    #Find initial guess for Matern amplitude and slope. Zero in on area around turbulent background peak
     turbIndex <- c((medIndex - numTest):(medIndex - 1), (medIndex + 1):(medIndex + numTest))
     parInit[5] <- quantile(((curr$sZ[turbIndex]*(delta*curr$omega[turbIndex])^2)/(max(curr$sZ) - curr$sZ[turbIndex])), quantSet)
     parInit[2] <- sqrt(max(curr$sZ)*parInit[5]); parInit[5] <- sqrt(parInit[5])

### 2016-05-10 Code/functions.R
hessian still not working

@@ -136,8 +136,8 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
 
 
 #Function to fit likelihood
-fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
-                     needInits = TRUE, parInit = NULL, hess = FALSE, simpModel = FALSE) {
+fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, hess = NULL,
+                     needInits = TRUE, parInit = NULL, getHess = FALSE, simpModel = FALSE) {
   
   #calc. periodogram in function
   curr <- getPerio(Z, delta, dB = FALSE, noZero = FALSE)
@@ -189,7 +189,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
     transParInit[2] <- log1p(parInit[2] - 1) #0 < B < inf ===> -inf < log(B) < inf 
     transParInit[3] <- parInit[3] #-inf < w0 < inf ===> -inf <- w0 < inf
     transParInit[4] <- log1p(parInit[4] - pi*sqrt(3)/N - 1) #pi*sqrt(3)/N < c < inf ===> -inf < log(c - pi*sqrt(3)/N) < inf
-    transParInit[5] <- log1p(parInit[5] - pi*sqrt(3)/N - 1) #pi*sqrt(3)/N < c < inf ===> -inf < log(h - pi*sqrt(3)/N) < inf
+    transParInit[5] <- log1p(parInit[5] - pi*sqrt(3)/N - 1) #pi*sqrt(3)/N < h < inf ===> -inf < log(h - pi*sqrt(3)/N) < inf
     transParInit[6] <- log1p(parInit[6] - 0.5 - 1) #0.5 < alpha < inf ===> -inf < log(alpha - 0.5) < inf
   } else {
     if (needInits == TRUE) {
@@ -199,7 +199,7 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE,
     transParInit[1] <- log1p(parInit[1] - 1) #0 < A < inf ===> -inf < log(A) < inf
     transParInit[2] <- log1p(parInit[2] - 1) #0 < B < inf ===> -inf < log(B) < inf 
[…]

### 2016-05-10 Code/figures.R
hessian still not working

@@ -248,18 +248,19 @@ abline(v = CF, lwd = 3) #inertial frequency
 
 #Figures 7, 8, 10
 #constants
-CF <- max(4*pi*(driftUlys$f)) #positive because of southern hemisphere, why at max not mean
+CF <- mean(4*pi*(driftUlys$f)) #positive because of southern hemisphere
 N <- length(driftUlys$num)
 DELTA <- 2
 nWin <- 999
 
 #find region to fit model in
-fracPos <- 1.75*max(4*pi*driftUlys$f)/pi #figure out exactly the logic of this cut-off
+fracPos <- 1.75*max(4*pi*driftUlys$f)/pi 
 
 #initial periodogram and par fit (use for set up)
 toStartPer <- getPerio(driftUlys$cv[1:nWin], DELTA, dB = TRUE, noZero = TRUE) 
 toStartFit <- fitModel(driftUlys$cv[1:nWin], CF,  DELTA, fracNeg = 0, fracPos = fracPos,
-                       quantSet = 0.8, incZero = TRUE, hess = TRUE)
+                       quantSet = 0.8, incZero = TRUE, getHess = TRUE)
+step <- qnorm(.975)*sqrt(diag(solve(toStartFit$hess)))
 
 #frequencies of interest
 firstIndex <- toStartFit$firstIndex
@@ -285,40 +286,46 @@ colnames(par6Val) <- c("A", "B", "w0", "C", "h", "alpha")
 par6LB <- par6UB <- par6Val
 par6Val[1, ] <- c(toStartFit$A, toStartFit$B, toStartFit$w0,
                   toStartFit$C, toStartFit$h, toStartFit$alpha)
+par6LB[1, ] <- par6Val[1, ] - step
+par6UB[1, ] <- par6Val[1, ] + step
 llVal6 <- rep(NA, N - 499*2)
 llVal6 <- toStartFit$llVal
+hessArray <- array(dim = c(N - 499*2, 6, 6), data = NA)
+hessArray[1,,] <- toStartFit$hess
 
 #Confidence interval for first time point
 step <- qnorm(.975)*sqrt(diag(solve(toStartFit$hess))/nWin)
 par6LB[1, ] <- par6Val[1, ] - step; par6UB[1, ] <- par6Val[1, ] + step
 
-
 #Loop through all time points and calc par estimate and CI
-#for (i in 501:(N - 499)) {
-#  currCv <- driftUlys$cv[(i - 499):(i + 499)] 
[…]

### 2016-05-11 Paper/paper.tex

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
 \usepackage{amsmath}
@@
 \begin{align*}
-S_{x}(\omega) = \int_{-\infty}^{\infty}s_{x}(\tau) e^{-\omega t}dt  \Longleftrightarrow \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{x}(\omega) e^{i \omega t} d\omega 
+S_{x}(\omega) = \int_{-\infty}^{\infty}s_{x}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{x}(\omega) e^{i \omega t} d\omega 
 \end{align*}
@@
 \begin{align*}
-\label{eq: diffEqDeriv}
+\label{eq:diffEqDeriv}
 \frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
@@
 \begin{equation}
-\label{eq: ouEq}
+\label{eq:ouEq}
 dz(t) = (i f_{0} -c) z(t) dt + A d Q(t), 
@@
 which defines equivalent behavior to the accepted deterministic formulation.
-Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
+Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
 It developed from brownian motion to OU to complex OU, mean-reversion, dampening and is a valid in this context by  blah (look up and cite the Mandelrot and Van Ness justification) giving the This model is always Markovian.
@@
 \begin{align}
-\label{eq: ouAC}
+\label{eq:ouAC}
 s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)\\
-\label{eq: ouPSD}
+\label{eq:ouPSD}
 S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}
@@
 \begin{align}
-\label{eq: maternAC}
+\label{eq:maternAC}
 s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
-\label{eq: maternPSD}
+\label{eq:maternPSD}
 S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
@@
 Combining the inertial oscillation and turbulent background parts of the model is straightforward.
-For both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background.
-Using Equations ~\ref{eq: ouAC} and ~\ref{eq: maternAC} and Equations ~\ref{eq: ouPSD} and ~\ref{eq: maternPSD}, we obtain,
+To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background.
+Using Equations ~\ref{eq:ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
 \begin{align}
 S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
-s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|),
+s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
 \end{align}
- the autocovariance and power spectral density respectively.
+
+\section{Results}
+We show the potential utility of this method by analyzing two non-trivial data sets.
+First, we consider the modeled trajectories of 200 near surface particles obtained from a computer simulation of ocean circulation (Set to be `very similar' to \citet{Danioux2008} SPEM 5.2 set-up).
+In Figure ~\ref{fig:numSim}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories.
+The resulting spectra has peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations, which are set not to have shifts.
+We also display the individual periodograms and model fits.
+\citet{Sykulski2016}  note that this highlights the method's robustness, since the individual periodograms show considerable variability, but the model fits are all quite consistent.
+\begin{figure}[hb]
+	\label{fig:numSim}
+  \centering
+    \includegraphics[width=.5\textwidth]{fig5.png}
+        \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
+Right: The ensemble mean periodogram (blue) with the ensemble fit for the frequencies used to fit it (green) and extended to the entire spectrum (red) and individual trajectories' periodograms (light grey) and model fits (dark fit) }
+\end{figure}
+
+We also consider the trajectory of a drifter in the South Pacific Ocean observed 12 times a day over a 1642 day period.
+Using a rolling window of 1000 observations and the semi-parameteric set-up described in Section X, we obtain parameter estimates for the spectral model at each time point.
+In Figure ~\ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density.
+The fitted density appears to be a smoothed version of the observed density, indicating a reasonable model fit.
+We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure 8 (add fig and renumber) and report the correlation among variables in Figure 10 (and fig and renumber).
+
+We also fit a 5-parameter where the frequency is set to the inertial frequency rather than being a free-parameter.
+We then compare the fit to the full model using a likelihood ratio test as described in Section X.
+In Figure 9, we display the test statistic over time where a red line is used to indicate the level of statistical significance.
+We conclude for time periods where the test statistic is greater than this cut-off that there is a shift in the inertial frequency.
+
+\begin{figure}[hb]
+	\label{fig:timeVarying}
+  \centering
+    \includegraphics[width=.5\textwidth]{fig7.png}
+        \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
+On both figures, the white line indicates the average inertial frequency within the rolling window at the current time period.
+Only frequencies used in the estimation are included.
+}
+\end{figure}
+
+\section{Discussion}
+identifiability
+
+\clearpage
 
@@
 
+\section{Appendix}
+
+\subsection{Errata}
+
+\subsection{Additional Figures}
+
 \end{document}

### 2016-05-11 Paper/paper.tex
adding figures

@@
 
-%\begin{abstract}
-%  Put your project summary here.
-%\end{abstract}
+\begin{abstract}
+  This report reviews the spectral analysis method for modeling ocean surface drifters proposed by \citet{Sykulski2016}.
+Previous methods to model drifters are reviewed along with the authors' model.
+Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are considered.
+We conclude by evaluating the strengths and weaknesses of this method.
+  \end{abstract}
 
@@
 However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
-The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
+Previous methods to model drifters are BLAH The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
 \\
@@
  To model the drifter's motion over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
-Stochastic models are then developed for both components of the the drifter's behavior, inertial oscillations and movement due to the ocean's turbulent background,  in the spectral domain.
-Finally, the two components of the model are added together, creating a six-parameter model, which is fit via maximizing a blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
+Stochastic models are then developed for both components of the drifter's behavior, inertial oscillations and movement due to the ocean's turbulent background,  in the spectral domain.
+Finally, the two components of the model are added together, creating a six-parameter model.
+Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood.
+This is a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
 
 \subsubsection{Inertial oscillations}
-We discuss modeling just the inertial oscillations first.
+We begin by discussing the model for the inertial oscillations.
 In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations \citep{Pollard1970}
@@
 where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
-Suppressing the dependence on $t$ and using the complex representation of the velocity, these relationships can be equivalently expressed as follows
+Suppressing the dependence on $t$ and using the complex representation of the velocity, we show how these relationships can be equivalently expressed as follows
 \begin{align*}
@@
 
+\subsection{Fitting the Model}
+A relatively new technique referred to as blurred Whittle likelihood is employed to fit these models.
+Blurred Whittle likelihood builds on Whittle likelihood, a standard approximation to the true likelihood in the time-domain which improves computation from order O to O.
+Whittle likelihood is a technique developed to maximize.
+Unfortunately, blurred whittle likelihood shows.
+
+To allow for non-stationarity, the authors propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
+The length of the window needs to be large enough to allow for a full sampling of the spectrum, but short enough to be treated as locally stationary.
+In this analysis, the authors select  the window length empirically, but acknowledge that further work on this subject would be of value.
+
+\citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred.
+To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$.
+Then this simpler model can be compared to the full six parameter model via a likelihood ratio test with test statistic
+\begin{equation}
+\label{eq: LRT}
+R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ]
+\end{equation}
+where $\hat{\theta}_{0}(t)$ is the fitted parameter vector for the simpler, or null model, and $\hat{\theta}_{1}(t)$ is the fitted parameter vector for the full model.
+The authors claim $R(t)$ then follow a Chi-Squared distribution with 1 degree of freedom.
+
 \section{Results}
@@
 
+\subsection{Computation}
+
 \end{document}

### 2016-05-11 Code/dataConversion.m

@@ -11,3 +11,7 @@ save('/users/hdirector/Documents/prelim/prelim/Code/newinertial.mat', '-v6')
 addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig6-10')
 load drifterulysses.mat
 save('/users/hdirector/Documents/prelim/prelim/Code/drifterulysses.mat', '-v6')
+
+addpath('/users/hdirector/Documents/prelim/prelim/MatlabCode/Fig6-10')
+load Ulysses6p.mat
+save('/users/hdirector/Documents/prelim/prelim/Code/Ulysses6p.mat', '-v6')

### 2016-05-11 Code/figures.R

@@ -169,11 +169,11 @@ sZMat[,1] <- sim1Per$sZ
 #Loop through and fit other 199 simulations 
 #run only once (this is slow)
 #for (i in 2:nSim) {
-  #fit each of the 200 simulations
+#fit each of the 200 simulations
 #  tempFit <- fitModel(vel[,i], CF = f0, DELTA, 1.5*f0/pi, 1.5*f0/pi, quantSet = .9)  #fit model
 #  tempPer <- getPerio(vel[,i], delta = DELTA, dB = TRUE, noZero = FALSE) #periodogram
 #  N <- length(vel[,i])
-  #plot spectrum and model fit 
+#plot spectrum and model fit 
 #  sTau <- ouAc(tempFit$A, tempFit$w0, tempFit$C, N) + maternAc(tempFit$B, tempFit$alpha, tempFit$h, N)  
 #  tau <- seq(0, N - 1)
 #  sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
@@ -388,7 +388,7 @@ colnames(par5Val) <- c("A", "B", "C", "h", "alpha")
 currCv <- driftUlys$cv[(500 - 499):(500 + 499)] 
 #fit parameters, initialize estimates at value used in last run
 toStartFit5 <- fitModel(currCv, CFVec[1], DELTA, fracNeg = 0, fracPos = fracPos,
-                    quantSet = .8, simpModel = TRUE)
+                        quantSet = .8, simpModel = TRUE)
 llVal5 <- rep(NA, N - 499*2)
 llVal5[1] <- toStartFit5$llVal
 par5Val[1, ] <- c(toStartFit5$A, toStartFit5$B, toStartFit5$C, 
@@ -399,11 +399,11 @@ par5Val[1, ] <- c(toStartFit5$A, toStartFit5$B, toStartFit5$C,
 #for (i in 501:(N - 499*2)) {
 #  currCv <- driftUlys$cv[(i - 499):(i + 499)] 
 #  CFCurr <- CFVec[i - 499]
-  #fit parameters, initialize estimates at value used in last run
+#fit parameters, initialize estimates at value used in last run
 #  tempFit <- fitModel(currCv, CFCurr, DELTA, fracNeg = 0, fracPos = fracPos,
[…]

### 2016-05-11 Code/functions.R

@@ -29,7 +29,7 @@ maternAc <- function(B, alpha, h, N, delta) {
   ac <- rep(NA, N)
   ac[1] <- ((B^2)*beta(0.5, alpha - 0.5))/(2*pi*abs(h)^(2*alpha  - 1)) #variance (cov at lag 0) 
   tau <- 1:(N - 1) #lags 
-  ac[2:N] <- ((B^2*(abs(h)*tau)^(alpha - 0.5)*besselK(abs(h)*tau, alpha - 0.5))/
+  ac[2:N] <- ((B^2*(abs(h)*delta*tau)^(alpha - 0.5)*besselK(abs(h)*delta*tau, alpha - 0.5))/
                 (2^(alpha - 1/2)*pi^(1/2)*exp(lgamma(alpha))*abs(h)^(2*alpha  - 1))) 
   return(ac)
 }
@@ -38,7 +38,7 @@ maternAc <- function(B, alpha, h, N, delta) {
 ouAc <- function(A, w0, C, N, delta) {
   ac <- rep(NA, N)
   tau <- 0:(N - 1) #lags 
-  ac <- (0i + A^2/abs(2*C))*exp(1i*w0*tau)*exp(0i + -abs(C)*tau)
+  ac <- (0i + A^2/abs(2*C))*exp(1i*w0*delta*tau)*exp(0i + -abs(C*delta*tau)
   return(ac)
 }
 
@@ -209,20 +209,20 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   #c > 0: ou dampening, h: matern slope, alpha: matern smoothness  (pg. 37)
   
   if (simpModel == FALSE) {
-  opt <- optim(transParInit, ll, delta = delta, curr = curr, 
-               firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex,
-               control = list(maxit = 1000000000, reltol=1e-1000))
+    opt <- optim(transParInit, ll, delta = delta, curr = curr, 
+                 firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex,
+                 control = list(maxit = 1000000000, reltol=1e-1000))
   } else {
-  opt <- optim(transParInit, llSimp, delta = delta, curr = curr, 
-          firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
-          medIndex = medIndex, 
-          control = list(maxit = 1000000000, reltol=1e-1000))
[…]

### 2016-05-16 Paper/paper.tex

@@
 Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are considered.
-We conclude by evaluating the strengths and weaknesses of this method.
+An evaluation of the strengths and weaknesses of this method concludes the review.
   \end{abstract}
@@
 Inertial oscillations are modeled with a complex Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
-Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
+Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that reduces bias.
+Time-varying parameter estimates are also introduced to allow for non-stationarity.
 \\
 \indent The remainder of this review proceeds as follows.
-After a succint review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
+After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model.
-The paper concludes with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+Section 4 concludes the review with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
 \section{Methods}
@@
 	\subsection{Spectral Analysis}
+	\label{sec: specAnalysis}
 	\indent We briefly review spectral analysis as these methods are central to understanding \citet{Sykulski2016}.
 More thorough coverage of this material can be found in  \citep{Percival1993}, for example.
-From a statistical perspective, a time series, $x(t)$ can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
+From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
 Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series.
 Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations 	\begin{align}
-x(t) \overset{ms}{=} \int_{-\infty}^{\infty} f_{x}(\omega)e^{i\omega t}d\omega && f_{x}(\omega) = \int_{-\infty}^{\infty} x(t) e^{-i \omega t }dt.
+z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
 \end{align}
 \citep{Percival1993}. To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
-
 \begin{equation}
-S_{x}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} x(t) e^{-i \omega t}dt \right|^{2} \right),
+S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
 \end{equation}
@@
 Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.
-Defining the autocovariance as $s_{x}(\tau) = \mathbb{E}[x(t) x(t - \tau)] $ where $\tau$ is the time lag, we have
-
-\begin{align*}
-S_{x}(\omega) = \int_{-\infty}^{\infty}s_{x}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{x}(\omega) e^{i \omega t} d\omega 
-\end{align*}
+For real $z(t)$, we define the autocovariance as $s_{x}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{x}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t).
+Letting $\tau$ is the time lag, we have
+\begin{align}
+\label{eq: fourierPair}
+S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
+\end{align}
 
@@
 \begin{equation}
+\label{eq: perio}
 J_{x}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} x(t) e^{-i \omega t}
@@
 \citep{Sykulski2013}.
-Although intuitively appealing, this estimator has known problems that stem from the sampling of a periodic function at discrete intervals.
+Although intuitively appealing, this estimator has known problems that stem from   sampling a periodic function at discrete intervals.
 In particular,  high frequencies can  capture periodic behavior of other lower frequencies if that happen to align in period.
@@
 
-\subsection{Model}
- To model the drifter's motion over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
-Stochastic models are then developed for both components of the drifter's behavior, inertial oscillations and movement due to the ocean's turbulent background,  in the spectral domain.
-Finally, the two components of the model are added together, creating a six-parameter model.
-Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood.
-This is a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
+\subsection{Stochastic Model}
+ To model the drifters' motions over time, the eastward and northward components of the drifters' velocities, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
+Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior.
+Working in the frequency domain, the two components of the model are added together, creating a six-parameter spectral model.
+Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
 
@@
 where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind.
-Suppressing the dependence on $t$ and using the complex representation of the velocity, we show how these relationships can be equivalently expressed as follows
-\begin{align*}
+Suppressing the dependence on $t$ and using the complex representation of the velocity,  these relationships can be equivalently expressed as follows
+\begin{align}
 \label{eq:diffEqDeriv}
-\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
-dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
-&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
-&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\ \nonumber
+dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\ \nonumber
+&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
+&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\ \nonumber
 &= (if_{0} - c)z dt + (F + iG)dt. 
-\end{align*}
-By further replacing the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, \citet{Sykulski2016} obtain the stochastic model 
+\end{align}
+ \citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain  the following stochastic analogue to the accepted deterministic equation for inertial oscillations:
  \begin{equation}
 \label{eq:ouEq}
-dz(t) = (i f_{0} -c) z(t) dt + A d Q(t), 
+dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
 \end{equation}
-which defines equivalent behavior to the accepted deterministic formulation.
-Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
-It developed from brownian motion to OU to complex OU, mean-reversion, dampening and is a valid in this context by  blah (look up and cite the Mandelrot and Van Ness justification) giving the This model is always Markovian.
-The OU model is stationary assuming the initial conditions of the stochastic differential equation are set as $z(t_{0}) \sim N(0, A^{2}/2|c|)$, and, if not, will revert to the stationary model quickly as the influence of the initial conditions decays exponentially (WHY?
-look this up).
 
-\citet{Sykulski2016} further let $f_{0}$ be a free parameter, denoted $\omega_{0}$, to be estimated by data.
-In so doing, they enable  identification and estimation of the shifts in the angular frequency due to eddies.
+Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation and is also sometimes referred to as the complex-valued stationary autoregressive process.
+This stochastic differential equation creates a process  that is both stationary and Markovian.
+The parameter $f_{0}$ represents the frequency of inertial oscillations.
+The dampening parameter $c > 0$  guarantees that the process is mean-reverting, meaning that it will always eventually return to its mean position, $z = 0$.
+The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.
+The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, is a standard complex Wiener process.
+We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle.
+Such a process $\{Q(t, \omega): t \geq 0\}$ on the space $(\omega, Q, P)$ must satisfy
+\begin{enumerate}
+\item $Q(0, \omega) = 0$ almost everywhere
+\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
+\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
+\end{enumerate}
+\citep{Hida1980}.
+
+\citet{Sykulski2016} propose replacing the known Coriolis frequency, $f_{0}$  in Equation ~\ref{eq:ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
+This allows for identification and estimation of the shifts in the angular frequency.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
 
-Letting $z^{*}(t)$ represent the complex conjugate of z(t) and using the previously-specified initial conditions, the complex OU process has the following  autocovariance  and power spectral density
+The covariance of the OU process can be derived by considering the two-dimensional equivalent of $z(t)$, as is done in  \citet{Arato1999}.
+Letting $z^{*}(t)$ represent the complex conjugate of z(t), they obtain the complex OU process autocovariance,
 \begin{align}
 \label{eq:ouAC}
-s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)\\
+s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)
+\end{align}
+where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the full model.
+Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
 \label{eq:ouPSD}
@@
 \end{align}
-where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the model.
-In addition to the inertial frequency, $\omega_{0}$, the parameter, $A > 0$ that defines the variability, and the parameter, $c >0$ which controls the dampening,  need to be fit from data.
+\noindent Setting $\tau = 0$ in Equation ~\ref{eq: ouAc}, gives the variance of the process, $A^{2}/2c$.
+
 \subsubsection{Turbulent Background}
@@
 Thus, the authors propose using the Mat\'{e}rn covariance structure \citep{Gneiting2012}, a model which has gained prominence in the spatial statistics literature due to its flexibility.
-In this context, the Mat\'{e}rn model usefully  allows for markovianity when the data supports it while not universally enforcing it.
-The corresponding autocovariance and power spectral density are (read and cite stein 1991 here)
-
+In this context, the Mat\'{e}rn model proves to be useful, since it allows for markovianity, but does not require it.
+The Mat\'{e}rn  autocovariance and power spectral density are
 \begin{align}
@@
 \end{align}
-where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$.
-and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model.
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
@@
 S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
+\label{eq: fullSpec}
 s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
@@
 \subsection{Fitting the Model}
-A relatively new technique referred to as blurred Whittle likelihood is employed to fit these models.
-Blurred Whittle likelihood builds on Whittle likelihood, a standard approximation to the true likelihood in the time-domain which improves computation from order O to O.
-Whittle likelihood is a technique developed to maximize.
-Unfortunately, blurred whittle likelihood shows.
+\subsubsection{Blurred Whittle likelihood}
+A new technique introduced in  \citet{Sykulski2013}  is used to fit the model.
+Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost.
+Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $Z$,  in the time domain that has the likelihood
+\begin{align*}
+l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z
+\end{align*}
+where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.
+It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
+\begin{equation}
+|J_{Z}(\omega)| \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
+\end{equation}
+under certain regularity conditions CITE.
+\citet{Whittle1953} combined this result with Fourier approximations to the true covariance structure and features of Toeplitz matrices to obtain the approximate likelihood
+\begin{align}
+l_{\omega}(\theta) &= - \sum_{\omega \in \Omega} \left[ J_{Z}^{H}(\omega) \tilde{S}_{Z}^{-1}(\omega; \theta) J_{Z}(\omega) + \log \{S(\omega; \theta) \} \right]\\
+\label{eq: bwl}
+&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]
+\end{align}
+ where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix.
+However, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
 
-To allow for non-stationarity, the authors propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
-The length of the window needs to be large enough to allow for a full sampling of the spectrum, but short enough to be treated as locally stationary.
-In this analysis, the authors select  the window length empirically, but acknowledge that further work on this subject would be of value.
+To improve on these problems, \citep{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$ with the expected value of the spectrum, $\overline{S}_{Z}(\omega, \theta)$\begin{equation}
+\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)
+\end{equation}
+in Equation ~\ref{eq: bwl}. This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
+\begin{equation}
+\mathcal{F}(\cdot) = \frac{\Delta}{2\pi N} \frac{sin^{2}(N \omega \Delta/2)}{sin^{2}(\omega \Delta /2)} 
+\end{equation} \citep{Sykulski2013}.
+This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram; thereby eliminating effects due to sampling at discrete intervals.
+This leads to the final likelihood
+\begin{align}
+l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]
+\end{align}
+which is maximized by standard optimization techniques.
 
-\citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred.
+\subsubsection{Semi-parametric fitting}
+The authors note that the fitting of Equation ~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters such as how the physical system behaves or artifacts from interpolation.
+They refer to these adaptations to the model as the semi-parametric approach and highlight its use  their analyses.
+Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations.
+To avoid this, the authors only fit the model using data on the positive or negative half of the spectrum corresponding to the sign of the inertial oscillation ( which is known based on the latitude of the drifter).
+Similarly, sampling effects make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
+
+\subsubsection{Non-stationarity}
+Further, to allow for non-stationarity in long time series, \citep{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
+This gives the OU differential equation
+\begin{equation}
+dz(t) = \{i \omega_{0}(t) - c(t) \} z(t) dt + A(t) dQ(t)
+\end{equation}
+and spectral power density
+\begin{equation}
+S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}\}^{2} + c(t)^{2}}. 
+\end{equation}
+with parameters as a function of time. Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
+\begin{equation}
+S^{(m)}(\omega, t) = \frac{B^{2}(t)}{\{\omega^{2} + h^{2}(t)\}^{\alpha(t)}}.
+\end{equation}
+The length of the window to fit these time-varying parameters needs to be large enough to allow for a full sampling of the spectrum, but short enough to be treated as locally stationary.
+In this analysis, the authors select a reasonable window length empirically, but acknowledge that further work on this subject would be of value.
+
+\subsubsection{Significance tests for simpler models}
+Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred.
 To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$.

### 2016-05-16 Paper/paper.tex

@@
 \subsubsection{Turbulent Background}
-The drifters' time series are also marked by large-scale turbulence (read and cite Rhines 1979).
-Many potential models have been considered for this background....Lit review here ...
+The drifters' time series are also marked by large-scale turbulence (see \citet{Rhines1979} for a review of the physics of this process.) Many potential models have been considered for this background, which are reviewed in \citep{Lacasce2008} ...Lit review here ...
 However, these models have been \textbf{\it{integer order}}, meaning the velocity, acceleration, and hyperacceleration must be Markovian, a constraint previous studies (READ AND CITE) indicate is not appropriate in all cases.

### 2016-05-16 Code/figures.R
getting close on code

@@ -1,11 +1,10 @@
 ###Hannah Director###
-#Research paper prelim
+#Research paper prelim 
 rm(list = ls())
 source("/users/hdirector/Documents/prelim/prelim/Code/functions.R")
 
-
 ###Figure 1 (left)###
-#Image from global drifter project (paper author's did not make this artistic figure
+#Image from global drifter project (paper author's did not make this artistic figure 
 #or provide the data for it
 
 ############################################
@@ -25,33 +24,22 @@ drift <- list("num" = as.vector(drifterbetty2$drifterbetty[,,1]$num),
               "cv" = as.vector(drifterbetty2$drifterbetty[,,1]$cv),
               "f" = as.vector(drifterbetty2$drifterbetty[,,1]$f))
 
-
 #Indices of sections of interest (b denotes the "blue" highlighted blue section and g the highlighted "green" section of the time series)
 startTs <- 1901; endTs <- 4900 #index of period of interest is obtained from Sykulski et al. posted Matlab code
 bStartTs <- 2971; bEndTs <-3570 #index of 50 day period in blue is obtained from Sykulski et al. posted Matlab code
 gStartTs <- 4201; gEndTs <- 4800 #index of 50 day period in blue is obtained from Sykulski et al. posted Matlab code
 
-#plot time series, coloring particular sections
+#plot time series, coloring particular sections 
 #pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig1.pdf')
 par(mfrow = c(1, 1))
-
-###Figure 1 (right)###
-#Indices of sections of interest
-startTs <- 1901; endTs <- 4900 #index of period of interest is obtained from Sykulski et al posted Matlab code
-bStartTs <- 2971; bEndTs <-3570 #index of 50 day period in blue is obtained from Sykulski et al posted Matlab code
[…]

### 2016-05-16 Code/functions.R

@@ -1,13 +1,11 @@
 ###Function to calculate periodogram###
-#Inputs: Z: complex-valued velocity, dB: Boolean specifying whether result should be
-#returned in decibels or not, noZero: Boolean specifying whether to use the zero in
-#estimation or not
+#Inputs: Z: complex-valued velocity, dB: Boolean specifying whether result should be 
+#returned in decibels or not, noZero: Boolean specifying whether to use the zero in 
+#estimation or not 
 #Outputs: list of omega (Fourier frequencies) and  sZ(spectral values at omega)
-
-#Calculate periodogram
 getPerio <- function(Z, delta, dB = TRUE, noZero = TRUE) {
   library("waved")
-  #fourier frquencies
+  #fourier frquencies 
   N <- length(Z)
   if(noZero == TRUE) {
     if (N%%2 == 1) {
@@ -30,30 +28,27 @@ getPerio <- function(Z, delta, dB = TRUE, noZero = TRUE) {
   return(list("sZ" = sZ, "omega" = omega))
 }
 
-
 ###Function to get autocorrelation for the Matern###
 #Inputs: B: amplitude, alpha: smoothing, h: dampening, N: length of time series, delta: sampling interval
-#Outputs: ac: autocovariance
-#Function to get autocorrelation for the matern
+#Outputs: ac: autocovariance 
 maternAc <- function(B, alpha, h, N, delta) {
   ac <- rep(NA, N)
-  ac[1] <- ((B^2)*beta(0.5, alpha - 0.5))/(2*pi*abs(h)^(2*alpha  - 1)) #variance (cov at lag 0)
-  tau <- 1:(N - 1) #lags
+  ac[1] <- ((B^2)*beta(0.5, alpha - 0.5))/(2*pi*abs(h)^(2*alpha  - 1)) #variance (cov at lag 0) 
+  tau <- 1:(N - 1) #lags 
   ac[2:N] <- ((B^2*(abs(h)*tau)^(alpha - 0.5)*besselK(abs(h)*tau, alpha - 0.5))/
-                (2^(alpha - 1/2)*pi^(1/2)*exp(lgamma(alpha))*abs(h)^(2*alpha  - 1)))
+                (2^(alpha - 1/2)*pi^(1/2)*exp(lgamma(alpha))*abs(h)^(2*alpha  - 1))) 
   return(ac)
 }
 
 ###Function to get autocorrelation for the OU process###
[…]

### 2016-05-16 Paper/paper.tex
full draft up to methods

--- 2016-05-16 Paper/paper.tex
@@
   This report reviews the spectral analysis method for modeling ocean surface drifters proposed by \citet{Sykulski2016}.
-Previous methods to model drifters are reviewed along with the authors' model.
-Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are considered.
-An evaluation of the strengths and weaknesses of this method concludes the review.
+Previous methods to model drifters are discussed along with the authors' model.
+Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are reviewed.
+To conclude the review, we evaluate the strengths and weaknesses of this method.
   \end{abstract}
@@
 The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.
-The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, is a standard complex Wiener process.
+The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$ is a standard complex Wiener process, defined to be such that $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, .
 We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle.
@@
 \subsubsection{Turbulent Background}
-The drifters' time series are also marked by large-scale turbulence (read and cite Rhines 1979).
-Many potential models have been considered for this background....Lit review here ...
-However, these models have been \textbf{\it{integer order}}, meaning the velocity, acceleration, and hyperacceleration must be Markovian, a constraint previous studies (READ AND CITE) indicate is not appropriate in all cases.
-Thus, the authors propose using the Mat\'{e}rn covariance structure \citep{Gneiting2012}, a model which has gained prominence in the spatial statistics literature due to its flexibility.
-In this context, the Mat\'{e}rn model proves to be useful, since it allows for markovianity, but does not require it.
-The Mat\'{e}rn  autocovariance and power spectral density are
+The drifters' time series are also marked by the ocean's large-scale turbulence.
+The underlying physics describing this red noise process is summarized in \citet{Rhines1979}.
+Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}.
+Within the oceanography literature, the stochastic models are designated  first-order \citep{Griffa1995, Falco2000}, second-order \citep{Sawford1991}, and third-order corresponding to  whether the velocity, acceleration, or hyperacceleration is modeled with a Markov process.
+However, calculation of the spectral slopes in observed data indicate that this class of models, referred to as \textbf{\it{integer order}},  does not always adequately reflect the ocean's behavior \citep{Rupolo1996, Sanderson1991}.
+Rather, \textbf{\it{fractional power models}}, such as fractional Brownian motion (FBM), would better reflect the observed behavior in some cases.
+
+Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background.
+The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility.
+In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers.
+The Mat\'{e}rn model is defined either by its autocovariance or power spectral density
 \begin{align}
@@
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
+
+The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model.
+This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values.
+They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
+Figure ~\ref{fig: fbmMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
+
+\begin{figure}[hb]
+	\label{fig: fmbMat}
+  \centering
+    \includegraphics[width=.5\textwidth]{fig3.pdf}
+        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
+Top Left: The eastward component the velocity; Top Right: The eastward and northward components of the vecocity; Bottom: Periodograms on the decibel scale}
+\end{figure}
 

### 2016-05-16 Paper/paper.tex
full draft of intro and methods

@@
 However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
-Previous methods to model drifters are BLAH The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
+The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
 \\
@@
         \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
-Top Left: The eastward component the velocity; Top Right: The eastward and northward components of the vecocity; Bottom: Periodograms on the decibel scale}
+Top Left: The eastward component the velocities.
+Top Right: The eastward and northward components of the velocities.
+Bottom: Periodograms of both processes on the decibel scale.
+}
 \end{figure}
@@
 \end{equation}
-under certain regularity conditions CITE.
+under certain regularity conditions \citep{Contreras2006}.
 \citet{Whittle1953} combined this result with Fourier approximations to the true covariance structure and features of Toeplitz matrices to obtain the approximate likelihood
@@
  where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix.
-However, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
+This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3}$  required for time-domain likelihood  \citep{Sykulski2016}.
+Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
 
@@
 \end{align}
-which is maximized by standard optimization techniques.
+which is maximized by standard optimization techniques. Using asymptotic properties of the Whittle likelihood, we also have 
+\begin{equation}
+\label{eq: fisher}
+\sqrt{N}(\hat{\theta} - \theta) \overset{d}{\rightarrow} \text{ Normal($0, \mathcal{F}^{-1})$ }.
+\end{equation}
+where $\mathcal{F}$ is the Fisher information. Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculations of the Hessian.
 
@@
 where $\hat{\theta}_{0}(t)$ is the fitted parameter vector for the simpler, or null model, and $\hat{\theta}_{1}(t)$ is the fitted parameter vector for the full model.
-The authors claim $R(t)$ then follow a Chi-Squared distribution with 1 degree of freedom.
+The authors claim $R(t)$ then follow a Chi-squared distribution with 1 degree of freedom.
 

### 2016-05-17 Paper/paper.tex
small edits

--- 2016-05-16 Paper/paper.tex
@@
 
+\usepackage{color}
+\usepackage{ulem}
+\newcommand{\vmdel}[1]{\sout{#1}}
+\newcommand{\vmadd}[1]{\textbf{\color{red}{#1}}}
+\newcommand{\vmcomment}[1]{({\color{blue}{VM's comment:}} \textbf{\color{blue}{#1}})}
+
 \begin{document}
@@
 
-\begin{abstract}
-  This report reviews the spectral analysis method for modeling ocean surface drifters proposed by \citet{Sykulski2016}.
-Previous methods to model drifters are reviewed along with the authors' model.
-Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are considered.
-An evaluation of the strengths and weaknesses of this method concludes the review.
-  \end{abstract}
+%\begin{abstract}
+%  Put your project summary here.
+%\end{abstract}
 
@@
 Olhede, Jonathan M.
-Lilly, and Eric Danioux presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
+Lilly, and Eric Danioux \vmcomment{I'd use "citet" here} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
 Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time.
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
-Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.\\
-\indent The motion of parcels of water moving over changing latitudes is known to have a rotational component.
+Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
+\par
+The motion of parcels of water moving over changing latitudes is known to have a rotational component.
 Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.
@@
 Inertial oscillations are modeled with a complex Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
-Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that reduces bias.
-Time-varying parameter estimates are also introduced to allow for non-stationarity.
+Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
 \\
 \indent The remainder of this review proceeds as follows.
-After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
+After a succint review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model.
-Section 4 concludes the review with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+The paper concludes with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
 \section{Methods}
@@
 	\subsection{Spectral Analysis}
-	\label{sec: specAnalysis}
 	\indent We briefly review spectral analysis as these methods are central to understanding \citet{Sykulski2016}.
 More thorough coverage of this material can be found in  \citep{Percival1993}, for example.
-From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
+From a statistical perspective, a time series, $x(t)$ can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
 Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series.
 Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations 	\begin{align}
-z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
+x(t) \overset{ms}{=} \int_{-\infty}^{\infty} f_{x}(\omega)e^{i\omega t}d\omega && f_{x}(\omega) = \int_{-\infty}^{\infty} x(t) e^{-i \omega t }dt.
 \end{align}
 \citep{Percival1993}. To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
+
 \begin{equation}
-S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
+S_{x}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} x(t) e^{-i \omega t}dt \right|^{2} \right),
 \end{equation}
@@
 Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.
-For real $z(t)$, we define the autocovariance as $s_{x}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{x}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t).
-Letting $\tau$ is the time lag, we have
-\begin{align}
-\label{eq: fourierPair}
-S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
-\end{align}
+Defining the autocovariance as $s_{x}(\tau) = \mathbb{E}[x(t) x(t - \tau)] $ where $\tau$ is the time lag, we have
+
+\begin{align*}
+S_{x}(\omega) = \int_{-\infty}^{\infty}s_{x}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{x}(\omega) e^{i \omega t} d\omega 
+\end{align*}
 
@@
 \begin{equation}
-\label{eq: perio}
 J_{x}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} x(t) e^{-i \omega t}
@@
 \citep{Sykulski2013}.
-Although intuitively appealing, this estimator has known problems that stem from   sampling a periodic function at discrete intervals.
+Although intuitively appealing, this estimator has known problems that stem from the sampling of a periodic function at discrete intervals.
 In particular,  high frequencies can  capture periodic behavior of other lower frequencies if that happen to align in period.
@@
 
-\subsection{Stochastic Model}
- To model the drifters' motions over time, the eastward and northward components of the drifters' velocities, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
-Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior.
-Working in the frequency domain, the two components of the model are added together, creating a six-parameter spectral model.
-Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
+\subsection{Model}
+ To model the drifter's motion over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
+Stochastic models are then developed for both components of the the drifter's behavior, inertial oscillations and movement due to the ocean's turbulent background,  in the spectral domain.
+Finally, the two components of the model are added together, creating a six-parameter model, which is fit via maximizing a blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
 
 \subsubsection{Inertial oscillations}
-We begin by discussing the model for the inertial oscillations.
+We discuss modeling just the inertial oscillations first.
 In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations \citep{Pollard1970}
@@
 Suppressing the dependence on $t$ and using the complex representation of the velocity, these relationships can be equivalently expressed as follows
-\begin{align}
+\begin{align*}
 \label{eq:diffEqDeriv}
-\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\ \nonumber
-dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\ \nonumber
-&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
-&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\ \nonumber
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\
+dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\
+&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\
+&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\
 &= (if_{0} - c)z dt + (F + iG)dt. 
-\end{align}
- \citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain  the following stochastic analogue to the accepted deterministic equation for inertial oscillations:
+\end{align*}
+By further replacing the wind forcing term, $(F + iG)dt$, with complex-valued Brownian increments, \citet{Sykulski2016} obtain the stochastic model 
 \begin{equation}
 \label{eq:ouEq}
-dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
+dz(t) = (i f_{0} -c) z(t) dt + A d Q(t), 
 \end{equation}
+which defines equivalent behavior to the accepted deterministic formulation.
+Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation.
+It developed from brownian motion to OU to complex OU, mean-reversion, dampening and is a valid in this context by  blah (look up and cite the Mandelrot and Van Ness justification) giving the This model is always Markovian.
+The OU model is stationary assuming the initial conditions of the stochastic differential equation are set as $z(t_{0}) \sim N(0, A^{2}/2|c|)$, and, if not, will revert to the stationary model quickly as the influence of the initial conditions decays exponentially (WHY?
+look this up).
 
-Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation and is also sometimes referred to as the complex-valued stationary autoregressive process.
-This stochastic differential equation creates a process  that is both stationary and Markovian.
-The parameter $f_{0}$ represents the frequency of inertial oscillations.
-The dampening parameter $c > 0$  guarantees that the process is mean-reverting, meaning that it will always eventually return to its mean position, $z = 0$.
-The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.
-The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, is a standard complex Wiener process.
-We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle.
-Such a process $\{Q(t, \omega): t \geq 0\}$ on the space $(\omega, Q, P)$ must satisfy
-\begin{enumerate}
-\item $Q(0, \omega) = 0$ almost everywhere
-\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
-\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
-\end{enumerate}
-\citep{Hida1980}.
-
-\citet{Sykulski2016} propose replacing the known Coriolis frequency, $f_{0}$  in Equation ~\ref{eq:ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
-This allows for identification and estimation of the shifts in the angular frequency.
+\citet{Sykulski2016} further let $f_{0}$ be a free parameter, denoted $\omega_{0}$, to be estimated by data.
+In so doing, they enable  identification and estimation of the shifts in the angular frequency due to eddies.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
 
-The covariance of the OU process can be derived by considering the two-dimensional equivalent of $z(t)$, as is done in  \citet{Arato1999}.
-Letting $z^{*}(t)$ represent the complex conjugate of z(t), they obtain the complex OU process autocovariance,
+Letting $z^{*}(t)$ represent the complex conjugate of z(t) and using the previously-specified initial conditions, the complex OU process has the following  autocovariance  and power spectral density
 \begin{align}
 \label{eq:ouAC}
-s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)
-\end{align}
-where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the full model.
-Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
+s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)\\
 \label{eq:ouPSD}
@@
 \end{align}
-\noindent Setting $\tau = 0$ in Equation ~\ref{eq: ouAc}, gives the variance of the process, $A^{2}/2c$.
-
+where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the model.
+In addition to the inertial frequency, $\omega_{0}$, the parameter, $A > 0$ that defines the variability, and the parameter, $c >0$ which controls the dampening,  need to be fit from data.
 \subsubsection{Turbulent Background}
-The drifters' time series are also marked by large-scale turbulence (see \citet{Rhines1979} for a review of the physics of this process.) Many potential models have been considered for this background, which are reviewed in \citep{Lacasce2008} ...Lit review here ...
+The drifters' time series are also marked by large-scale turbulence (read and cite Rhines 1979).
+Many potential models have been considered for this background....Lit review here ...
 However, these models have been \textbf{\it{integer order}}, meaning the velocity, acceleration, and hyperacceleration must be Markovian, a constraint previous studies (READ AND CITE) indicate is not appropriate in all cases.
 Thus, the authors propose using the Mat\'{e}rn covariance structure \citep{Gneiting2012}, a model which has gained prominence in the spatial statistics literature due to its flexibility.
-In this context, the Mat\'{e}rn model proves to be useful, since it allows for markovianity, but does not require it.
-The Mat\'{e}rn  autocovariance and power spectral density are
+In this context, the Mat\'{e}rn model usefully  allows for markovianity when the data supports it while not universally enforcing it.
+The corresponding autocovariance and power spectral density are (read and cite stein 1991 here)
+
 \begin{align}
@@
 \end{align}
-where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$.
+and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model.
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
@@
 S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
-\label{eq: fullSpec}
 s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
 \end{align}
-
-\subsection{Fitting the Model}
-\subsubsection{Blurred Whittle likelihood}
-A new technique introduced in  \citet{Sykulski2013}  is used to fit the model.
-Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost.
-Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $Z$,  in the time domain that has the likelihood
-\begin{align*}
-l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z
-\end{align*}
-where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.
-It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
-\begin{equation}
-|J_{Z}(\omega)| \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
-\end{equation}
-under certain regularity conditions CITE.
-\citet{Whittle1953} combined this result with Fourier approximations to the true covariance structure and features of Toeplitz matrices to obtain the approximate likelihood
-\begin{align}
-l_{\omega}(\theta) &= - \sum_{\omega \in \Omega} \left[ J_{Z}^{H}(\omega) \tilde{S}_{Z}^{-1}(\omega; \theta) J_{Z}(\omega) + \log \{S(\omega; \theta) \} \right]\\
-\label{eq: bwl}
-&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]
-\end{align}
- where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix.
-However, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
-
-To improve on these problems, \citep{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$ with the expected value of the spectrum, $\overline{S}_{Z}(\omega, \theta)$\begin{equation}
-\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)
-\end{equation}
-in Equation ~\ref{eq: bwl}. This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
-\begin{equation}
-\mathcal{F}(\cdot) = \frac{\Delta}{2\pi N} \frac{sin^{2}(N \omega \Delta/2)}{sin^{2}(\omega \Delta /2)} 
-\end{equation} \citep{Sykulski2013}.
-This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram; thereby eliminating effects due to sampling at discrete intervals.
-This leads to the final likelihood
-\begin{align}
-l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]
-\end{align}
-which is maximized by standard optimization techniques.
-
-\subsubsection{Semi-parametric fitting}
-The authors note that the fitting of Equation ~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters such as how the physical system behaves or artifacts from interpolation.
-They refer to these adaptations to the model as the semi-parametric approach and highlight its use  their analyses.
-Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations.
-To avoid this, the authors only fit the model using data on the positive or negative half of the spectrum corresponding to the sign of the inertial oscillation ( which is known based on the latitude of the drifter).
-Similarly, sampling effects make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
-
-\subsubsection{Non-stationarity}
-Further, to allow for non-stationarity in long time series, \citep{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
-This gives the OU differential equation
-\begin{equation}
-dz(t) = \{i \omega_{0}(t) - c(t) \} z(t) dt + A(t) dQ(t)
-\end{equation}
-and spectral power density
-\begin{equation}
-S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}\}^{2} + c(t)^{2}}. 
-\end{equation}
-with parameters as a function of time. Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
-\begin{equation}
-S^{(m)}(\omega, t) = \frac{B^{2}(t)}{\{\omega^{2} + h^{2}(t)\}^{\alpha(t)}}.
-\end{equation}
-The length of the window to fit these time-varying parameters needs to be large enough to allow for a full sampling of the spectrum, but short enough to be treated as locally stationary.
-In this analysis, the authors select a reasonable window length empirically, but acknowledge that further work on this subject would be of value.
-
-\subsubsection{Significance tests for simpler models}
-Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred.
-To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$.
-Then this simpler model can be compared to the full six parameter model via a likelihood ratio test with test statistic
-\begin{equation}
-\label{eq: LRT}
-R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ]
-\end{equation}
-where $\hat{\theta}_{0}(t)$ is the fitted parameter vector for the simpler, or null model, and $\hat{\theta}_{1}(t)$ is the fitted parameter vector for the full model.
-The authors claim $R(t)$ then follow a Chi-Squared distribution with 1 degree of freedom.
 
@@
 
-\subsection{Computation}
-
 \end{document}

### 2016-05-17 Code/figures.R
minor changes

@@ -54,6 +54,10 @@ green <- getPerio(Z = gZ, delta = DELTA)
 gCf <- mean(-4*pi*drift$f[gStartTs:gEndTs])
 bCf <- mean(-4*pi*drift$f[bStartTs:gStartTs])
 
+#time series lengths
+bN <- length(bZ)
+gN <- length(gZ)
+
 #plot figure
 #pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig2.pdf')
 par(mfrow = c(1, 2))
@@ -71,6 +75,7 @@ abline(v = gCf, col = "red")
 bCF <- mean(-4*pi*(drift$f[bStartTs:bEndTs]))
 gCF <- mean(-4*pi*(drift$f[gStartTs:gEndTs]))
 
+
 #Fit and plot blue time series
 bFit <- fitModel(bZ, bCF, delta = DELTA, fracNeg = 0.4, fracPos = 0, quantSet = .5)
 #pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig4.pdf')
@@ -78,36 +83,36 @@ par(mfrow = c(1,2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
 mtext("Replication of Figure 4", outer = T, line = -3)
-sTau <- ouAc(bFit$A, bFit$w0, bFit$C, N, delta = DELTA) + maternAc(bFit$B, bFit$alpha, bFit$h, N, delta = DELTA)  
-tau <- seq(0, N - 1)
-sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
+sTau <- ouAc(bFit$A, bFit$w0, bFit$C, bN, delta = DELTA) + maternAc(bFit$B, bFit$alpha, bFit$h, bN, delta = DELTA)  
+tau <- seq(0, bN - 1)
+sBar <- 2*fft(sTau*(1 - (tau/bN))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
 points(DELTA*blue$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd = 3)
[…]

### 2016-05-17 Code/functions.R

@@ -99,6 +99,7 @@ simFBM <- function(B, alpha, H, N) {
 #output: negative of the log likelihood value 
 ll <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
                incZero = FALSE, trans = TRUE) {
+  N <- length(curr$sZ)
   #parameters to evaluate
   if (trans == TRUE) {
     A <- expm1(theta[1]) + 1; B <- expm1(theta[2]) + 1; w0 <- theta[3]
@@ -110,7 +111,6 @@ ll <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
     C <- theta[4]; h <- theta[5]; alpha <- theta[6]
   }
   #calculate likelihood
-  N <- length(curr$sZ)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, w0, C, N, delta = 1) + maternAc(B, alpha, h, N, delta = 1)  
   sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) #Interpet this, why no negs for sBar?
@@ -185,7 +185,6 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   #calculate periodogram for data
   curr <- getPerio(Z, delta, dB = FALSE, noZero = FALSE)
   
-  
   N <- length(curr$sZ)
   
   #set window of frequencies to consider in evalution (assuming frequencies sorted min to max)
@@ -247,11 +246,11 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   if (simpModel == FALSE) {
     library("pracma")
     opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,
-                      firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
+                      firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex) 
   } else {
     opt <- fminsearch(llSimp, transParInit, delta = delta, curr = curr, 
                       firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
-                      medIndex = medIndex, maxiter = 3000)
+                      medIndex = medIndex)
   }
   llVal <- opt$fval
   

### 2016-05-17 Paper/paper.tex
minor changes
intro edits

--- 2016-05-16 Paper/paper.tex
@@
 \bibliographystyle{plainnat}
+
+\usepackage{color}
+\usepackage{ulem}
+\newcommand{\vmdel}[1]{\sout{#1}}
+\newcommand{\vmadd}[1]{\textbf{\color{red}{#1}}}
+\newcommand{\vmcomment}[1]{({\color{blue}{VM's comment:}} \textbf{\color{blue}{#1}})}
 
@@
 \section{Introduction}
-\indent ``Lagrangian Time Series Models for Ocean Surface Drifter Trajectories," by Adam M.
+``Lagrangian Time Series Models for Ocean Surface Drifter Trajectories," by Adam M.
 Sykulski, Sofia C.
 Olhede, Jonathan M.
-Lilly, and Eric Danioux presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
+Lilly, and Eric Danioux \vmcomment{I'd use "citet" here} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
 Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time.
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
-Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.\\
-\indent The motion of parcels of water moving over changing latitudes is known to have a rotational component.
+Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
+\par
+The motion of parcels of water moving over changing latitudes is known to have a rotational component.
 Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.
@@
 The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
-\\
-\indent In achieving this result, several statistical advances are made.
+\vmcomment{What other approaches have been tried before?
+Need more literature review of previously used statistical methods for identifying eddies.}
+\par
+In achieving this result, several statistical advances are made.
 The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations.
-For the turbulent background model, the Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is extended to the time-series context.
-This allows for a more flexible model for background behavior than previously obtained.
-Inertial oscillations are modeled with a complex Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
-Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that reduces bias.
-Time-varying parameter estimates are also introduced to allow for non-stationarity.
-\\
-\indent The remainder of this review proceeds as follows.
-After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
+For the turbulent background model, the Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is extended to the time-series context \vmcomment{This sounds strange to me.
+I thought the Mat\'{e}rn kernel was general enough to work with any metric space, including the real line, and has long been used for time series.} This allows for a more flexible model for background behavior than previously obtained.
+Inertial oscillations are modeled with a complex\vmadd{-domain} Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
+Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
+\par
+The remainder of this review proceeds as follows.
+After a succint review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it.
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model.
-Section 4 concludes the review with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+The paper concludes with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+
 \section{Methods}

### 2016-05-17 Paper/paper.tex
updated code slightly and added line breaks

@@
 This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects being observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude.
-Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations.
-These motions have known frequencies, referred to as inertial frequencies.
+Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. These motions have known frequencies, referred to as inertial frequencies.
  Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
@@
 For the turbulent background model, the Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is extended to the time-series context \vmcomment{This sounds strange to me.
-I thought the Mat\'{e}rn kernel was general enough to work with any metric space, including the real line, and has long been used for time series.} This allows for a more flexible model for background behavior than previously obtained.
+I thought the Mat\'{e}rn kernel was general enough to work with any metric space, including the real line, and has long been used for time series.}
+This allows for a more flexible model for background behavior than previously obtained.  
 Inertial oscillations are modeled with a complex\vmadd{-domain} Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
@@
 \end{align}
-\citep{Percival1993}. To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
+\citep{Percival1993}. 
+To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
 \begin{equation}
@@
 \end{align}
+
  \citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain  the following stochastic analogue to the accepted deterministic equation for inertial oscillations:
@@
  The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.
-The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$ is a standard complex Wiener process, defined to be such that $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes, .
+The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$ is a standard complex Wiener process, defined to be such that $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
   We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle. 
@@
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
+ The parameter $\alpha$ is of particular importance, since it controls the smoothness, or degree of differentiability, of the process \citep{Fuentes2010}.
 
@@
 where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.
+
  It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
@@
 \end{equation}
-in Equation ~\ref{eq: bwl}. This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
+in Equation ~\ref{eq: bwl}. 
+This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
 \begin{equation}
@@
 \end{align}
-which is maximized by standard optimization techniques. Using asymptotic properties of the Whittle likelihood, we also have 
+which is maximized by standard optimization techniques.
+
+Using asymptotic properties of the Whittle likelihood, we also have 
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{F}$ is the Fisher information. Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculations of the Hessian.
+where $\mathcal{F}$ is the Fisher information. 
+Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculations of the Hessian.
 
@@
 \end{equation}
-with parameters as a function of time. Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
+with parameters as a function of time. 
+Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
 \begin{equation}

### 2016-05-17 Code/figures.R

@@ -305,7 +305,7 @@ for (i in 500:(N - 500)) {
   sZMat[i - 499, ] <- tempPer$sZ[firstIndex:lastIndex]
 }
 
-#Storage vectors for parameters, confidence interval bands, and hessians
+#Storage vectors for parameters, negative log likelihood, confidence interval bands, and hessians
 par6Val <- matrix(nrow = N - 999, ncol = 6)
 colnames(par6Val) <- c("A", "B", "w0", "C", "h", "alpha")
 par6LB <- par6UB <- par6Val
@@ -321,41 +321,43 @@ par6Val[1, ] <- c(toStartFit$A, toStartFit$B, toStartFit$w0,
 par6LB[1, ] <- par6Val[1, ] - step
 par6UB[1, ] <- par6Val[1, ] + step
 
-#Loop through all windows and calculate parameter estimates, confidence intervals, and hessian
-for (i in 501:(N - 500)) {
-  currCv <- driftUlys$cv[(i - 499):(i + 500)] 
-  CFCurr <- CFVec[i - 499]
-  #fit parameters, initialize estimates at value used in last run
-  tempFit <- fitModel(currCv, CFCurr, DELTA, fracNeg = 0, fracPos = fracPos,
-                      quantSet = .5, needInits = FALSE, parInit = par6Val[i - 499 - 1, ],
-                      getHess = TRUE)
-  par6Val[i - 499, ] <- c(tempFit$A, tempFit$B, tempFit$w0,
-                          tempFit$C, tempFit$h, tempFit$alpha)
-  llVal6[i - 499] <-  tempFit$llVal
-  #calculate confidence interval via Fisher's information
-  step <- qnorm(.975)*sqrt(diag(solve(tempFit$hess)))
-  par6LB[i - 499, ] <- par6Val[i - 499, ] - step; par6UB[i - 499, ] <- par6Val[i - 499, ] + step
-  hessArray[i - 499,,] <- tempFit$hess
-  print(i)
-}
-
-#store or load results
-save(par6Val, file = '/users/hdirector/Documents/prelim/prelim/Code/par6Val.rda')
-save(par6LB, file =  '/users/hdirector/Documents/prelim/prelim/Code/par6LB.rda')
-save(par6UB, file = '/users/hdirector/Documents/prelim/prelim/Code/par6UB.rda')
[…]

### 2016-05-17 Code/functions.R

@@ -137,6 +137,7 @@ ll <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
 #output: negative of the log likelihood value 
 llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
                    CF, incZero = FALSE, trans = TRUE) {
+  N <- length(curr$sZ)
   #parameters to evaluate
   if (trans == TRUE) {
     A <- expm1(theta[1]) + 1; B <- expm1(theta[2]) + 1; 
@@ -148,7 +149,6 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
     C <- theta[3]; h <- theta[4]; alpha <- theta[5]
   }
   #calculate likelihood
-  N <- length(curr$sZ)
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, CF, C, N, delta = 1) + maternAc(B, alpha, h, N, delta = 1)  
   sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) #Interpet this, why no negs for sBar?
@@ -259,8 +259,8 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
     fin <-  c(expm1(opt$xval[1]) + 1, expm1(opt$xval[2]) + 1 , opt$xval[3], expm1(opt$xval[4]) + pi*sqrt(3)/N + 1, 
               expm1(opt$xval[5]) + pi*sqrt(3)/N + 1, expm1(opt$xval[6]) + 0.5 + 1)
   } else {
-    fin <-  c(expm1(opt$par[1]) + 1, expm1(opt$par[2]) + 1, expm1(opt$par[3]) + pi*sqrt(3)/N + 1, 
-              expm1(opt$par[4]) + pi*sqrt(3)/N + 1, expm1(opt$par[5]) + 0.5 + 1)
+    fin <-  c(expm1(opt$xval[1]) + 1, expm1(opt$xval[2]) + 1, expm1(opt$xval[3]) + pi*sqrt(3)/N + 1, 
+              expm1(opt$xval[4]) + pi*sqrt(3)/N + 1, expm1(opt$xval[5]) + 0.5 + 1)
   }
   
   #Calculate Hessian where needed

### 2016-05-18 Code/figures.R
adding replicated figures to paper

@@ -30,7 +30,8 @@ bStartTs <- 2971; bEndTs <-3570 #index of 50 day period in blue is obtained from
 gStartTs <- 4201; gEndTs <- 4800 #index of 50 day period in blue is obtained from Sykulski et al. posted Matlab code
 
 #plot time series, coloring particular sections 
-#pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig1.pdf')
+#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig1.pdf')
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.4,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1, 1))
 plot(drift$lon[startTs:endTs], drift$lat[startTs:endTs], type = "l", 
      xlab = "Longitude", ylab = "Latitude", main = "Replication of Figure 1 (Right)")
@@ -59,7 +60,9 @@ bN <- length(bZ)
 gN <- length(gZ)
 
 #plot figure
-#pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig2.pdf')
+#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig2.pdf',
+#    height = 3, width = 8.5)
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1, 2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
@@ -75,10 +78,11 @@ abline(v = gCf, col = "red")
 bCF <- mean(-4*pi*(drift$f[bStartTs:bEndTs]))
 gCF <- mean(-4*pi*(drift$f[gStartTs:gEndTs]))
 
-
 #Fit and plot blue time series
 bFit <- fitModel(bZ, bCF, delta = DELTA, fracNeg = 0.4, fracPos = 0, quantSet = .5)
-#pdf('/users/hdirector/Documents/prelim/prelim/ReplicatedFigures/fig4.pdf')
+#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig4.pdf',
+#    height = 3, width = 8.5)
[…]

### 2016-05-18 Code/functions.R

@@ -151,12 +151,13 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
   #calculate likelihood
   tau <-  seq(0, N - 1, 1)
   sTau <- ouAc(A, CF, C, N, delta = 1) + maternAc(B, alpha, h, N, delta = 1)  
-  sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) #Interpet this, why no negs for sBar?
+  sBar <- 2*fft(sTau*(1 - (tau/N))) - sTau[1]; sBar = abs(Re(fftshift(sBar))) 
   if (incZero == TRUE) {
     llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex]) + sum(log(sBar[firstIndex:lastIndex]))
   } else {
-    tempIndex <- c(firstIndex:(medIndex - 1), (medIndex + 1):lastIndex)
-    llVal <- sum(curr$sZ[firstIndex:lastIndex]/sBar[firstIndex:lastIndex]) + sum(log(sBar[firstIndex:lastIndex]))
+    tempIndex <- firstIndex:lastIndex
+    tempIndex <- tempIndex[-which(tempIndex == medIndex)]
+    llVal <- sum(curr$sZ[tempIndex]/sBar[tempIndex]) + sum(log(sBar[tempIndex]))
   }
   return(llVal)
 }
@@ -243,8 +244,8 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   }
   
   #Maximize likelihood numerically
+  library("pracma")
   if (simpModel == FALSE) {
-    library("pracma")
     opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,
                       firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex) 
   } else {

### 2016-05-21 Paper/paper.tex
adding replicated figures to paper
methods written

@@
 \usepackage{amsmath}
+\usepackage{wrapfig}
+\usepackage{subfigure}
 %%\setlength{\oddsidemargin}{0.25in}
@@
 \newcommand{\vmcomment}[1]{({\color{blue}{VM's comment:}} \textbf{\color{blue}{#1}})}
+\newcommand{\hdcomment}[1]{({\color{red}{HD's comment:}} \textbf{\color{red}{#1}})}
 
@@
   Previous methods to model drifters are discussed along with the authors' model. 
-    Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex Ornstein-Uhlenbeck process, and Whittle likelihood are reviewed. 
-     To conclude the review, we evaluate the strengths and weaknesses of this method. 
+    Where relevant to understanding the model of \citet{Sykulski2016},  spectral analysis, Mat\'{e}rn covariances, the complex-valued Ornstein-Uhlenbeck process, and the Whittle likelihood are reviewed.
+     To conclude the report, we evaluate the strengths and weaknesses of this method. 
   \end{abstract}
@@
 \section{Introduction}
-``Lagrangian Time Series Models for Ocean Surface Drifter Trajectories," by Adam M.
-Sykulski, Sofia C.
-Olhede, Jonathan M.
-Lilly, and Eric Danioux \vmcomment{I'd use "citet" here} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters.
+\citet{Sykulski2016} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
 Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
+Figure ~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
  Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
+ \begin{figure}[h!]
+	\label{fig: fig1}
+  \centering
+    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
+        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue.
+The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
+\end{figure}
+
 \par
@@
 This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects being observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude.
-Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. These motions have known frequencies, referred to as inertial frequencies.
+Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
+These motions have known frequencies, referred to as inertial frequencies.
  Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
   However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
- The methods in this paper overcome this challenge and successfully find shifts in the inertial frequency with a spectral time series model.
-\vmcomment{What other approaches have been tried before?
-Need more literature review of previously used statistical methods for identifying eddies.}
+ The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citet{Isern2003} or \citet{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
+\citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
+ Similarly, \citet{Boebel2003} calculate the curvature of the trajectory  and then classify eddies as periods of consistent rotations in one direction. 
+ \citet{Sykulski2016} assert that they are the first to identify eddies by explicitly modeling inertial oscillations within a stochastic model.  
  \par
@@
 The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations. 
-For the turbulent background model, the Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is extended to the time-series context \vmcomment{This sounds strange to me.
-I thought the Mat\'{e}rn kernel was general enough to work with any metric space, including the real line, and has long been used for time series.}
-This allows for a more flexible model for background behavior than previously obtained.  
-Inertial oscillations are modeled with a complex\vmadd{-domain} Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
+The Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is applied for the first time to model the turbulent background. 
+Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
+Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
 Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
 \par
-The remainder of this review proceeds as follows.
- After a succint review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
+The remainder of this report proceeds as follows.
+ After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
  Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
- The paper concludes with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
 
@@
 	\label{sec: specAnalysis}
-	\indent We briefly review spectral analysis as these methods are central to understanding \citet{Sykulski2016}.  
-	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
+	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
+	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
@@
 Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.  
-For real $z(t)$, we define the autocovariance as $s_{x}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{x}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t).
-Letting $\tau$ is the time lag, we have
+For real $z(t)$, we define the autocovariance as $s_{z}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t).
+Letting $\tau$ be the time lag, we have
 \begin{align}
 \label{eq: fourierPair}
-S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}dt  \Longleftrightarrow s_{x}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
+S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
 \end{align}
 
-\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, $\hat{S}_{x}(\omega) = |J_{x}(\omega)|^{2}$ where 
+\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, $\hat{S}_{z}(\omega) = |J_{z}(\omega)|^{2}$ where 
 \begin{equation}
 \label{eq: perio}
-J_{x}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} x(t) e^{-i \omega t}
+J_{z}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} z(t) e^{-i \omega t}
 \end{equation}
@@
 Although intuitively appealing, this estimator has known problems that stem from  sampling a periodic function at discrete intervals. 
-In particular,  high frequencies can  capture periodic behavior of other lower frequencies if that happen to align in period. 
+In particular,  high frequencies can  capture periodic behavior of other lower frequencies if they happen to align in period. 
 Additionally, frequencies above the highest observable frequency cannot be captured, and, are instead, recorded as other frequencies in the spectrum.  
@@
 \subsection{Stochastic Model}
- To model the drifters' motions over time, the eastward and northward components of the drifters' velocities, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
+ To model a drifter's movement over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, is converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
  Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior. 
@@
 We begin by discussing the model for the inertial oscillations. 
-In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations \citep{Pollard1970}
+In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
 \begin{align}
+\label{eq: deterOsc}
 \frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
@@
 \end{align}
-where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind. 
-Suppressing the dependence on $t$ and using the complex representation of the velocity,  these relationships can be equivalently expressed as follows
+where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}.
+Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
 \begin{align}
@@
 \end{align}
-
- \citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain  the following stochastic analogue to the accepted deterministic equation for inertial oscillations:
+\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation ~\ref{eq: deterOsc},
  \begin{equation}
-\label{eq:ouEq}
+\label{eq: ouEq}
 dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
 \end{equation}  
-
-Equation ~\ref{eq:ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation and is also sometimes referred to as the complex-valued stationary autoregressive process.
-This stochastic differential equation creates a process  that is both stationary and Markovian. 
-The parameter $f_{0}$ represents the frequency of inertial oscillations.
-The dampening parameter $c > 0$  guarantees that the process is mean-reverting, meaning that it will always eventually return to its mean position, $z = 0$.
- The final term of the differential equation is a positive constant $A$ multiplied by $dQ(t)$.
-The term $Q(t) = Q_{1}(t) + i Q_{2}(t), t \geq 0$ is a standard complex Wiener process, defined to be such that $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
-  We remind the reader that Wiener processes are stochastic models originally used to represent the random motion of a particle. 
-  Such a process $\{Q(t, \omega): t \geq 0\}$ on the space $(\omega, Q, P)$ must satisfy
+Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
+In general,  OU processes are considered continuous-time equivalents to the AR(1) process.
+Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate Markov process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
+ The complex-valued OU process is the complex representation of a bivariate OU process.  
+ In the formulation of Equation ~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
+ The final term of Equation ~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
+The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
+ We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
+  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
 \begin{enumerate}
@@
 
-\citet{Sykulski2016} propose replacing the known Coriolis frequency, $f_{0}$  in Equation ~\ref{eq:ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data. 
-This allows for identification and estimation of the shifts in the angular frequency. 
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation ~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
 
-The covariance of the OU process can be derived by considering the two-dimensional equivalent of $z(t)$, as is done in  \citet{Arato1999}. 
-Letting $z^{*}(t)$ represent the complex conjugate of z(t), they obtain the complex OU process autocovariance,
+For time series modeling, we are interested in the autocovariance of this process at different time lags.
+To derive this value given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
+This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
+\begin{equation}
+\label{eq: expec}
+\mathbb{E}\{z(t) z^{*}(t + \tau)\} = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i\{\mathbb{E}[u(t) v(t + \tau)] + \mathbb{E}[u(t)v( t + \tau)]\}
+\end{equation}
+\citep{DeIaco2003} where $z^{*}(t + \tau)$ represents the complex conjugate of $z(t + \tau)$. The differential equations for components $u$ and $v$ can be expressed in matrix form as 
 \begin{align}
-\label{eq:ouAC}
-s^{(0)}(\tau) &= \mathbb{E}\{z(t)z^{*}(t + \tau) \} = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|)
+Z(t) = \left( \begin{array}{c} u(t) \\ v(t) \end{array} \right) &= 
+\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q(t) \\ Ad Q(t) \end{array} \right).
 \end{align}
-where $\tau$ is a time lag and the $(o)$ subscripts denotes that these equations are for the OU component of the full model. 
+For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0) $ where $s(0)$ represents the covariance at zero lag and $B$ is the generator matrix \citep{Schach1971}.
+Using the known variance of a univariate OU process and the independence of the components $u(0)$, we have
+\begin{align}
+s(0) = \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right).
+\end{align}
+Thus, 
+\begin{align}
+\label{eq: covDeriv}
+\mathbb{E} \left[\left( \begin{array}{c} u(t + \tau) \\ v(t + \tau)  \end{array} \right) \left( \begin{array}{cc} u(t) & v(t) \end{array} \right)  \right] = e^{B \tau} s(0) = 
+e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)
+\end{align}
+where the second equality can be seen by expanding the infinite series representations for $\exp(-\lambda \tau)$, $\sin(\tau \omega)$, and $\cos( \tau \omega)$. 
+Using the relevant components of Equation ~\ref{eq: covDeriv} in Equation ~\ref{eq: expec} and applying Euler's formula, we obtain the complex OU process autocovariance as defined in \citet{Sykulski2016}
+\begin{align}
+\label{eq: ouAC}
+s^{(0)}(\tau) & = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
+\end{align}
  Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
 \label{eq:ouPSD}
-S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}
+S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
 \end{align}
-\noindent Setting $\tau = 0$ in Equation ~\ref{eq: ouAc}, gives the variance of the process, $A^{2}/2c$.
+The $(o)$ subscripts denotes that these equations are for the OU component of the full model.
 
@@
 The drifters' time series are also marked by the ocean's large-scale turbulence. 
-The underlying physics describing this red noise process is summarized in \citet{Rhines1979}. 
+The physics of ocean turbulence is summarized in \citet{Rhines1979}. 
 Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. 
@@
 They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
- Figure ~\ref{fig: fbmMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
+ Figure ~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
 
-\begin{figure}[hb]
+\begin{figure}[h!]
 	\label{fig: fmbMat}
   \centering
-    \includegraphics[width=.5\textwidth]{fig3.pdf}
+    \includegraphics[width=.7\textwidth]{ReplicatedFigures/fig3.pdf}
         \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
@@
 Combining the inertial oscillation and turbulent background parts of the model is straightforward.
- To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background. 
- Using Equations ~\ref{eq:ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
+ To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
+ Using Equations ~\ref{eq: ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
 \begin{align}
@@
 
+\begin{figure}[h!]
+	\label{fig: fig4}
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
+        \caption{Periodograms of the time series sections colored blue and green in Figure ~\ref{fig: fig1}.
+The black line denotes the inertial frequency.
+The dotted lines on the left figure indicates the location of the eddy peak and the expected shift in the inertial frequency.
+The semi-parametric model fit on the section of the data used in fitting the model is highlighted green on the left figure and blue on the right figure.
+The red dotted line indicates the model fit extended beyond the frequencies where it was fit.
+}
+\end{figure}
+
 \subsection{Fitting the Model}
 \subsubsection{Blurred Whittle likelihood}
-A new technique introduced in  \citet{Sykulski2013}  is used to fit the model.
+\par A recent technique introduced in  \citet{Sykulski2013}  is used to fit the model.
  Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost.
@@
  where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix. 
-  This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3}$  required for time-domain likelihood  \citep{Sykulski2016}.
+  This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}.
   Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
 
-To improve on these problems, \citep{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$ with the expected value of the spectrum, $\overline{S}_{Z}(\omega, \theta)$\begin{equation}
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$ with the expected value of the spectrum, $\overline{S}_{Z}(\omega, \theta)$\begin{equation}
 \overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)
@@
 \end{equation} \citep{Sykulski2013}. 
-This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram; thereby eliminating effects due to sampling at discrete intervals.
+This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram.
+Thereby, effects due to sampling at discrete intervals are avoided.
 This leads to the final likelihood
@@
 where $\mathcal{F}$ is the Fisher information. 
-Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculations of the Hessian.
+Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculation of the Hessian.
 
 \subsubsection{Semi-parametric fitting}
-The authors note that the fitting of Equation ~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters such as how the physical system behaves or artifacts from interpolation.
-They refer to these adaptations to the model as the semi-parametric approach and highlight its use  their analyses. 
+\label{sec: semi}
+The authors note that the fitting of Equation ~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
+They refer to making these adjustments as applying a semi-parametric approach. 
+In their analysis, they highlight using knowledge of the physical system and adjusting for artifacts from interpolation. 
 Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations. 
-To avoid this, the authors only fit the model using data on the positive or negative half of the spectrum corresponding to the sign of the inertial oscillation ( which is known based on the latitude of the drifter).
+To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.
 Similarly, sampling effects make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
+Figure ~\ref{fig: fig4} illustrates a semi-parametric model fit.
 
@@
 \end{equation}
-with parameters as a function of time. 
+with parameters now a function of time. 
 Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
@@
 \subsubsection{Significance tests for simpler models}
+\label{sec: sig}
 Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred. 
@@
  The authors claim $R(t)$ then follow a Chi-squared distribution with 1 degree of freedom. 
+ However, we note that this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
 
@@
 We show the potential utility of this method by analyzing two non-trivial data sets.  
-First, we consider the modeled trajectories of 200 near surface particles obtained from a computer simulation of ocean circulation (Set to be `very similar' to \citet{Danioux2008} SPEM 5.2 set-up). 
-In Figure ~\ref{fig:numSim}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
+First, we consider the modeled trajectories of 200 near-surface particles obtained from a computer simulation of ocean circulation. 
+The simulation is set to be `very similar' to the SPEM 5.2 set-up of \citet{Danioux2008}). 
+In Figure ~\ref{fig: numSim}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
 The resulting spectra has peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations, which are set not to have shifts.
@@
 \citet{Sykulski2016}  note that this highlights the method's robustness, since the individual periodograms show considerable variability, but the model fits are all quite consistent.
-\begin{figure}[hb]
-	\label{fig:numSim}
+
+\begin{figure}[h!]
+	\label{fig: numSim}
   \centering
-    \includegraphics[width=.5\textwidth]{fig5.png}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig5.png}
         \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
@@
 We also consider the trajectory of a drifter in the South Pacific Ocean observed 12 times a day over a 1642 day period. 
-Using a rolling window of 1000 observations and the semi-parameteric set-up described in Section X, we obtain parameter estimates for the spectral model at each time point.
+Using a rolling window of 1000 observations and the semi-parameteric set-up described in Section ~\ref{sec: semi} , we obtain parameter estimates for the spectral model at each time point.
  In Figure ~\ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density.
   The fitted density appears to be a smoothed version of the observed density, indicating a reasonable model fit.
-   We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure 8 (add fig and renumber) and report the correlation among variables in Figure 10 (and fig and renumber).
+   We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure ~\ref{fig: fig8} report the correlation among variables in Figure 10 (and fig and renumber).
+Figure ~\ref{fig: timeVarying} displays the observed spectra versus the modeled spectra.
 
 We also fit a 5-parameter where the frequency is set to the inertial frequency rather than being a free-parameter. 
-We then compare the fit to the full model using a likelihood ratio test as described in Section X.
+We then compare the fit to the full model using a likelihood ratio test as described in Section ~\ref{sec: sig}.
  In Figure 9, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
@@
   \centering
-    \includegraphics[width=.5\textwidth]{fig7.png}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig7.png}
         \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
@@
 }
+\end{figure}
+
+ \begin{figure}[h!]
+	\label{fig: fig9}
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig9.pdf}
+        \caption{ADD CAPTION}
+\end{figure}
+
+\begin{figure}[h!]
+	\label{fig: fig8}
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.pdf}
+        \caption{ADD CAPTION}
 \end{figure}
@@
 
+\begin{figure}[h!]
+	\label{fig: fig2}
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig2.pdf}
+        \caption{ADD CAPTION}
+\end{figure}
+
 \subsection{Computation}

### 2016-05-21 Code/figures.R

@@ -335,36 +335,36 @@ par6LB[1, ] <- par6Val[1, ] - step
 par6UB[1, ] <- par6Val[1, ] + step
 
 #load results, rather than re-running
-load('/users/hdirector/Documents/prelim/prelim/Code/par6Val.rda')
-load('/users/hdirector/Documents/prelim/prelim/Code/par6LB.rda')
-load('/users/hdirector/Documents/prelim/prelim/Code/par6UB.rda')
-load('/users/hdirector/Documents/prelim/prelim/Code/llVal6.rda')
-load('/users/hdirector/Documents/prelim/prelim/Code/hess.rda')
+#load('/users/hdirector/Documents/prelim/prelim/Code/par6Val.rda')
+#load('/users/hdirector/Documents/prelim/prelim/Code/par6LB.rda')
+#load('/users/hdirector/Documents/prelim/prelim/Code/par6UB.rda')
+#load('/users/hdirector/Documents/prelim/prelim/Code/llVal6.rda')
+#load('/users/hdirector/Documents/prelim/prelim/Code/hess.rda')
 
 #Loop through all windows and calculate parameter estimates, confidence intervals, and hessian
-#for (i in 501:(N - 500)) {
-#  currCv <- driftUlys$cv[(i - 499):(i + 500)] 
-#  CFCurr <- CFVec[i - 499]
-  #fit parameters, initialize estimates at value used in last run
-#  tempFit <- fitModel(currCv, CFCurr, DELTA, fracNeg = 0, fracPos = fracPos,
-#                      quantSet = .5, needInits = FALSE, parInit = par6Val[i - 499 - 1, ],
-#                      getHess = TRUE)
-#  par6Val[i - 499, ] <- c(tempFit$A, tempFit$B, tempFit$w0,
-#                          tempFit$C, tempFit$h, tempFit$alpha)
-#  llVal6[i - 499] <-  tempFit$llVal
-  #calculate confidence interval via Fisher's information
-#  step <- qnorm(.975)*sqrt(diag(solve(tempFit$hess)))
-#  par6LB[i - 499, ] <- par6Val[i - 499, ] - step; par6UB[i - 499, ] <- par6Val[i - 499, ] + step
-#  hessArray[i - 499,,] <- tempFit$hess
-#  print(i)
-#}
+for (i in 501:(N - 500)) {
+  currCv <- driftUlys$cv[(i - 499):(i + 500)] 
[…]

### 2016-05-21 Code/functions.R

@@ -246,12 +246,12 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   #Maximize likelihood numerically
   library("pracma")
   if (simpModel == FALSE) {
-    opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,
-                      firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex) 
+    opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,firstIndex = firstIndex,
+                       lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
   } else {
     opt <- fminsearch(llSimp, transParInit, delta = delta, curr = curr, 
                       firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
-                      medIndex = medIndex)
+                      medIndex = medIndex, maxiter = 3000)
   }
   llVal <- opt$fval
   
@@ -266,9 +266,10 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   
   #Calculate Hessian where needed
   if (simpModel == FALSE & getHess == TRUE) {
-    library("numDeriv")
-    hess <- hessian(ll, fin, delta = delta, curr = curr, trans = FALSE,
-                    firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex)
+    library("optimx")
+    hess <- optimHess(fin, ll,  delta = delta, curr = curr, trans = FALSE,
+                  firstIndex = firstIndex, lastIndex = lastIndex, medIndex = medIndex,
+                  control = list(maxit = 2000))
   } 
   
   #Return results

### 2016-05-21 Paper/paper.tex
methods written

@@
 The black line denotes the inertial frequency.
-The dotted lines on the left figure indicates the location of the eddy peak and the expected shift in the inertial frequency.
-The semi-parametric model fit on the section of the data used in fitting the model is highlighted green on the left figure and blue on the right figure.
+The dotted lines on the left figure indicate the location of the eddy peak and the expected shift in the inertial frequency.
+The semi-parametric model fit for the frequencies used in model fitting is highlighted green on the left figure and blue on the right figure.
 The red dotted line indicates the model fit extended beyond the frequencies where it was fit.

### 2016-05-21 Paper/paper.tex
full draft done

@@
  \begin{figure}[h!]
-	\label{fig: fig1}
   \centering
@@
 The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
+        	\label{fig: fig1}
 \end{figure}
@@
 Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
-These motions have known frequencies, referred to as inertial frequencies.
+These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions of rotation in the two hemispheres.
  Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
@@
 \begin{figure}[h!]
-	\label{fig: fmbMat}
   \centering
@@
 }
+        	\label{fig: fmbMat}
 \end{figure}
@@
 \begin{figure}[h!]
-	\label{fig: fig4}
   \centering
@@
 }
+        	\label{fig: fig4}
 \end{figure}
@@
 Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred. 
-To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$. 
+To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$, creating a five parameter model. 
 Then this simpler model can be compared to the full six parameter model via a likelihood ratio test with test statistic
@@
 \section{Results}
-We show the potential utility of this method by analyzing two non-trivial data sets.  
-First, we consider the modeled trajectories of 200 near-surface particles obtained from a computer simulation of ocean circulation. 
+\subsection{Simulated data}
+\par We show the potential utility of this method by analyzing two non-trivial data sets.  
+First, we consider the modeled trajectories of 200 near-surface particles obtained from a computer simulation of ocean circulation, where each of the trajectories are recorded at two hour intervals. 
 The simulation is set to be `very similar' to the SPEM 5.2 set-up of \citet{Danioux2008}). 
+In the  simulation, the turbulence is designed to be at a steady-state and wind forcing is set to be  constant. 
+The simulated data is useful as a verification tool, since  it contains no eddies and its only  notable asymmetry between the two directions of rotation is due to the inertial oscillations.
+
+\par For all two hundred trajectories, the full six-parameter model, given in Equation ~\ref{eq: fullSpec}, is fit to each trajectory. 
+The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
+The parameters are assumed constant, since this portion of the time series does is fairly short. 
 In Figure ~\ref{fig: numSim}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
-The resulting spectra has peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations, which are set not to have shifts.
 We also display the individual periodograms and model fits. 
+The resulting spectra consistently have peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations accurately. 
 \citet{Sykulski2016}  note that this highlights the method's robustness, since the individual periodograms show considerable variability, but the model fits are all quite consistent.
@@
 \begin{figure}[h!]
-	\label{fig: numSim}
   \centering
@@
         \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
-Right: The ensemble mean periodogram (blue) with the ensemble fit for the frequencies used to fit it (green) and extended to the entire spectrum (red) and individual trajectories' periodograms (light grey) and model fits (dark fit) }
+Right: The ensemble mean periodogram (blue) with the ensemble fit for the frequencies used to fit it (green) and extended to the entire spectrum (red).
+Individual trajectories' periodograms (light grey) and model fits (dark fit) are also displayed.
+}
+        \label{fig: fig5}
 \end{figure}
 
-We also consider the trajectory of a drifter in the South Pacific Ocean observed 12 times a day over a 1642 day period. 
-Using a rolling window of 1000 observations and the semi-parameteric set-up described in Section ~\ref{sec: semi} , we obtain parameter estimates for the spectral model at each time point.
- In Figure ~\ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density.
+\subsection{South Pacific Drifter}
+\label{sec: spDrift}
+We also consider the trajectory of a drifter in the south Pacific ocean observed 12 times a day over a 1642 day period.  
+The trajectory is displayed in the right panel of Figure ~\ref{fig: fig6}. 
+We first considering fitting a 20-day sample of this trajectory. 
+Its periodogram, displayed in ~\ref{fig: fig6}, shows substantial asymmetry between the positive and negative frequencies.
+Thus, the semi-parametric approach is applied and only positive frequencies are used in fitting the model.
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig6.pdf}
+        \caption{Left: The trajectory of the drifter in the south Pacific ocean.
+Right: The periodogram and model fit for the section of the trajectory highlighted in red.
+The green portion of the model fit corresponds to the frequencies used in fitting and the red dotted line is an extension of the fit to other frequencies.
+}
+        \label{fig: fig6}
+\end{figure}
+
+\par
+Moving on to the entire time series, we use rolling windows of 1000 observations and the semi-parametric set-up described in Section ~\ref{sec: semi}  to obtain parameter estimates for the spectral model at each time point.
+ In Figure ~\ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density in a time-frequency plot. 
   The fitted density appears to be a smoothed version of the observed density,  indicating a reasonable model fit.
-   We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure ~\ref{fig: fig8} report the correlation among variables in Figure 10 (and fig and renumber).
-Figure ~\ref{fig: timeVarying} displays the observed spectra versus the modeled spectra.
+   Typically the inertial frequency matches the Coriolis frequency, which would be expected given that shifts from the Coriolis frequency are only rare events.
 
-We also fit a 5-parameter where the frequency is set to the inertial frequency rather than being a free-parameter. 
-We then compare the fit to the full model using a likelihood ratio test as described in Section ~\ref{sec: sig}.
- In Figure 9, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
- We conclude for time periods where the test statistic is greater than this cut-off that there is a shift in the inertial frequency.
-
-\begin{figure}[hb]
-	\label{fig:timeVarying}
+   \begin{figure}[h!]
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig7.png}
+    \includegraphics[width=1\textwidth]{ReplicatedFigures/fig7.png}
         \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
@@
 }
+        	\label{fig:timeVarying}
+\end{figure}
+
+ \par
+ We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure ~\ref{fig: fig8}. 
+The 95$\%$ intervals are obtained using numerical calculation of the Hessian.
+The parameters all change smoothly over time. 
+We note that the inertial frequency differs from the Coriolis Frequency from roughly days 600 to 900. 
+\citet{Sykulski2016} note that this deviation is appropriate and can be explained by a change in tidal energy over shallow periods of water. 
+ \begin{figure}[h!]
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.png}
+        \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands. The red line in the top figure is the inertial frequency.}
+        	\label{fig: fig8}
 \end{figure}
@@
  \begin{figure}[h!]
+  \centering
+    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig9.pdf}
+        \caption{Time series of the likelihood ratio test statistic comparing the 5-parameter to 6-parameter models.
+The dotted red line indicates the value at which test statistics indicate a significant difference between the model.
+}
         	\label{fig: fig9}
-  \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig9.pdf}
-        \caption{ADD CAPTION}
 \end{figure}
 
+\par We also fit the 5-parameter model where the frequency is set to the inertial frequency rather than being a free-parameter. 
+We then compare this fit to the full model using a likelihood ratio test as described in Section ~\ref{sec: sig}.
+ In Figure 9, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
+The authors conclude that when there are statistically significant differences between the 5-parameter and 6-parameter model as indicated by the likelihood ratio test,   the inertial frequency does not equal the Coriolis frequency.
+They note that the period from days 600 to 900 always shows a significant difference between the models, which corresponds to the period of shifts observed in Figure \ref{fig: fig8}.
+
+\par We, however, are skeptical of this use of simple hypothesis testing. 
+The parameters are being estimated repeatedly using a rolling window. 
+This means that there is not independence between the time periods and there is repeated testing. 
+Both of these issues should be accounted for in designing a hypothesis test. 
+Further, the authors acknowledge elsewhere in the paper that the length of the rolling window is currently selected empirically and that further work on selecting windows would be beneficial. 
+This suggests that the p-value for any time point could be easily influenced by the length of the window used in fitting it.
+Thus, robustness checks of this method to the length of the window would be appropriate prior to using it.
+In other words, further thought should go in to designing a hypothesis that accurately reflects the problem set-up and modeling.
+
  \begin{figure}[h!]
-	\label{fig: fig8}
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.pdf}
-        \caption{ADD CAPTION}
+    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig10.pdf}
+        \caption{Table of the average correlation between parameters.
+Note that the parameters are ordered differently than in the paper, due to differences in the structure of the author's and my code.
+There are also minor differences in the correlation values, particularly for values close to zero.
+These differences can be attributed to slight differences in the numerical optimization used by R and Matlab in calculating the Hessian.
+}
+        	\label{fig: fig10}
 \end{figure}
 
+\par Finally, in Figure ~\ref{fig: fig10}, we display a matrix of the average correlation of the parameters over time. \citet{Sykulski2016} highlight several useful results from this table. 
+They note that the inertial frequency is generally uncorrelated with other parameters values, suggesting that the inertial frequency only controls the location of the peak, but does not influence the shape or amplitude of the spectra as desired.
+They also highlight the many correlations between parameters that are part of the turbulent background component and inertial oscillation components of the model. 
+This indicates that these two aspects of the ocean's behavior are linked and should be modeled together as has been done in this paper.
+
 \section{Discussion}
-identifiability
+\par Overall, \citet{Sykulski2016} present a  well-developed model for capturing the behavior of ocean surface drifters.
+ In particular, they develop a two component spectral model that jointly accounts for the turbulent background behavior of the ocean and the influence of inertial oscillations. 
+ Using the complex-valued Ornstein-Uhlenbeck process, the authors develop a stochastic analogue to the accepted deterministic differential equations used to describe this phenomena. 
+ In so doing, they are the first to explicitly model inertial oscillations within the context of a stochastic model. 
+ Further, to model the turbulent background, the authors propose using the  Mat\'{e}rn autocovariance structure.
+  This flexible model cleverly encompasses all models that have previously been proposed for the turbulent background as special cases; thus, allowing the data to determine what is the most appropriate behavior.
+   They also employ the newly-developed Blurred Whittle likelihood for model fitting, which reduces biases compared to other fitting techniques. 
+   All together, this spectral model provides an automated way to detect shifts from the Coriolis frequency in a long time series. 
+This provide an efficient way to detect eddies for large numbers and/or long time series. 
+So, the authors have both satisfied their scientific goal and have introduced new statistical ideas in doing so.
+
+Nevertheless, despite its overall sophistication, some areas in the paper could be improved. 
+As discussed in Section \ref{sec: spDrift}, the simple likelihood ratio test proposed in the paper may not be sufficient for the involved, time-varying model used in the paper.  
+Another area that is insufficiently addressed in the paper is possible identifiability issues between the model parameters. 
+Since primary interest is in the inertial frequency, if parameter estimates are highly correlated, physical changes in one parameter might actually be reflected by changes in other fitted parameter.
+ This would subsequently invalidate conclusions about observed changes in specific parameters, since they could be masked by changes in other parameters. 
+ The authors do allude to a lack of correlation between the inertial frequency and other model components, but this idea is not fully addressed.
+Overall, though \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
+ Their work is a substantial contribution to stochastic modeling, particularly, in terms, of explicitly incorporating known physics into statistical models.
 
@@
 \subsection{Errata}
+We note a few minor errors in the paper.
+\begin{itemize}
+\item Equation 13 has a typo. The second term in the sum should exactly match Equation 6. 
+\item In Figure 6, the periodogram and model fit are for days 350 to 370 not days 358 to 378. (This is apparent from looking at the author's code.)
+\item In Figure 10, the fifth row in the table should be labeled $\alpha$ and not $\delta$.
+\end{itemize}
 
-\subsection{Additional Figures}
+\subsection{Additional Replicated Figures}
 
 \begin{figure}[h!]
-	\label{fig: fig2}
   \centering
     \includegraphics[width=\textwidth]{ReplicatedFigures/fig2.pdf}
-        \caption{ADD CAPTION}
+        \caption{Periodograms of the highlighted green and blue sections of the trajectories in Figure ~\ref{fig: fig1}. The inertial frequencies are marked in red. }
+        	\label{fig: fig2}
 \end{figure}
 
-\subsection{Computation}
-
 \end{document}

### 2016-05-21 Code/figures.R

@@ -335,11 +335,11 @@ par6LB[1, ] <- par6Val[1, ] - step
 par6UB[1, ] <- par6Val[1, ] + step
 
 #load results, rather than re-running
-#load('/users/hdirector/Documents/prelim/prelim/Code/par6Val.rda')
-#load('/users/hdirector/Documents/prelim/prelim/Code/par6LB.rda')
-#load('/users/hdirector/Documents/prelim/prelim/Code/par6UB.rda')
-#load('/users/hdirector/Documents/prelim/prelim/Code/llVal6.rda')
-#load('/users/hdirector/Documents/prelim/prelim/Code/hess.rda')
+load('/users/hdirector/Documents/prelim/prelim/Code/par6Val.rda')
+load('/users/hdirector/Documents/prelim/prelim/Code/par6LB.rda')
+load('/users/hdirector/Documents/prelim/prelim/Code/par6UB.rda')
+load('/users/hdirector/Documents/prelim/prelim/Code/llVal6.rda')
+load('/users/hdirector/Documents/prelim/prelim/Code/hess.rda')
 
 #Loop through all windows and calculate parameter estimates, confidence intervals, and hessian
 for (i in 501:(N - 500)) {
@@ -367,16 +367,16 @@ save(llVal6, file = '/users/hdirector/Documents/prelim/prelim/Code/llVal6.rda')
 save(hessArray, file = '/users/hdirector/Documents/prelim/prelim/Code/hess.rda')
 
 #plot observed periodogram
-png('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig7.png',
-    height = 400, width = 600)
-par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.4,5,4,1.1), mai = c(.5, 1, .5, 1), oma = rep(.5, 4))
+#png('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig7.png',
+#    height = 600, width = 600)
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3,5,4,1.1), mai = c(.9, 1, .5, 1), oma = rep(.5, 4))
 par(mfrow = c(2, 1))
 
 library("fields")
 image.plot(xVal, 2*toStartPer$omega[firstIndex:lastIndex],
            sZMat, useRaster = TRUE, zlim = c(20, 60),
[…]

### 2016-05-22 Paper/paper.tex
methods comments

@@
 \section{Introduction}
-\citet{Sykulski2016} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
+\citet{Sykulski2016} present\vmdel{s} a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
 Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
-Figure ~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
+Figure~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
 Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
+\vmcomment{The last sentence is very low on information content.}
  \begin{figure}[h!]
@@
 	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
-	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
+	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
 	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series. 
-	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations 	\begin{align}
-z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
+	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations\vmadd{:} 	
+	\begin{align}
+z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega\vmadd{,} && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
 \end{align}
@@
 \label{eq: deterOsc}
-\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
-\frac{\partial v}{\partial t} - f_{0}u &= G - cv
+\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu\vmadd{,} \\ \nonumber
+\frac{\partial v}{\partial t} - f_{0}u &= G - cv\vmadd{,}
 \end{align}
@@
  The complex-valued OU process is the complex representation of a bivariate OU process.  
- In the formulation of Equation ~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
- The final term of Equation ~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
+ In the formulation of Equation~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
+ The final term of Equation~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
 The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation ~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
 \end{align}
-For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0) $ where $s(0)$ represents the covariance at zero lag and $B$ is the generator matrix \citep{Schach1971}.
+For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0)$\vmadd{,} where $s(0)$ represents the covariance at zero lag \vmdel{and $B$ is the generator matrix} \citep{Schach1971}.
+\vmcomment{Generator is defined neither in general nor in your particular case.}
 Using the known variance of a univariate OU process and the independence of the components $u(0)$, we have
@@
 \mathbb{E} \left[\left( \begin{array}{c} u(t + \tau) \\ v(t + \tau)  \end{array} \right) \left( \begin{array}{cc} u(t) & v(t) \end{array} \right)  \right] = e^{B \tau} s(0) = 
-e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)
+e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)\vmadd{,}
 \end{align}
@@
 The drifters' time series are also marked by the ocean's large-scale turbulence. 
-The physics of ocean turbulence is summarized in \citet{Rhines1979}. 
+The physics of ocean turbulence is summarized \vmdel{in} \vmadd{by} \citet{Rhines1979}. 
 Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. 
@@
  In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
- The Mat\'{e}rn model is defined either by its autocovariance or power spectral density
+ The Mat\'{e}rn model is defined either by its autocovariance \vmdel{or power spectral density}
 \begin{align}
 \label{eq:maternAC}
-s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
+s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\vmadd{,}
+\end{align}
+\vmadd{or power spectral density,}
+\begin{align}
 \label{eq:maternPSD}
-S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
+S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}
 \end{align}
-where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the s\vmdel{ub}\vmadd{uper}script $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
@@
 They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
- Figure ~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
+ Figure~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
 
@@
 Bottom: Periodograms of both processes on the decibel scale.
-}
+\vmcomment{I would include legends or direct labeling for all your figures.}}
         	\label{fig: fmbMat}
@@
 \begin{align}
-S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
+S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}\\
 \label{eq: fullSpec}
@@
 \subsubsection{Blurred Whittle likelihood}
-\par A recent technique introduced in  \citet{Sykulski2013}  is used to fit the model.
+\par A recent technique introduced by \citet{Sykulski2013}  is used to fit the model.
  Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost.
@@
 \begin{align*}
-l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z
+l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z\vmadd{,}
 \end{align*}
 where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.
+\vmcomment{I should have mentioned this earlier in the test that all vectors and matrices need to be bolded.}
 
@@
 \label{eq: bwl}
-&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]
+&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]\vmadd{,}
 \end{align}
@@
   This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}.
-  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.
+  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section~\ref{sec: specAnalysis}.
 
-To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$ with the expected value of the spectrum, $\overline{S}_{Z}(\omega, \theta)$\begin{equation}
-\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$\vmadd{, in Equation ~\ref{eq: bwl}} with the expected value of the spectrum, \vmdel{$\overline{S}_{Z}(\omega, \theta)$}
+\begin{equation}
+\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)\vmadd{.}
 \end{equation} 
-in Equation ~\ref{eq: bwl}. 
 This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
@@
 \begin{align}
-l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]
+l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]\vmadd{,}
 \end{align}
-which is maximized by standard optimization techniques.
+which is maximized by standard \vmadd{off-the-shelf nonlinear} optimization techniques. 
+\vmcomment{State the exact method(s) that the paper and you used.}
 
@@
 \label{eq: fisher}
-\sqrt{N}(\hat{\theta} - \theta) \overset{d}{\rightarrow} \text{ Normal($0, \mathcal{F}^{-1})$ }.
+\sqrt{N}(\hat{\theta} - \theta) \overset{d}{\rightarrow} \text{ Normal($0, \mathcal{F}^{-1})$ }\vmdel{.}\vmadd{,}
 \end{equation}
@@
 \label{sec: semi}
-The authors note that the fitting of Equation ~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
+The authors note that the fitting of Equation~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
 They refer to making these adjustments as applying a semi-parametric approach. 
@@
 Similarly, sampling effects make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
-Figure ~\ref{fig: fig4} illustrates a semi-parametric model fit.
+Figure~\ref{fig: fig4} illustrates a semi-parametric model fit.
 
 \subsubsection{Non-stationarity}
-Further, to allow for non-stationarity in long time series, \citep{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
-This gives the OU differential equation
+Further, to allow for non-stationarity in long time series, \citet{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
+This \vmdel{gives} \vmadd{makes parameters in} the OU differential equation
 \begin{equation}
@@
 \begin{equation}
-S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}\}^{2} + c(t)^{2}}. 
+S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}\}^{2} + c(t)^{2}}\vmdel{.} 
 \end{equation}
-with parameters now a function of time. 
+\vmdel{with parameters} now \vmadd{be} \vmdel{a} function\vmadd{s} of time. 
 Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
@@
 \label{eq: LRT}
-R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ]
+R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ]\vmadd{,}
 \end{equation}
@@
  However, we note that this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
+\vmcomment{I don't think the lack of independence between time points has an impact on the LRT, but using multiple LRTs without multiple testing correction, which is complicated by non-independence, is the main problem here, I think.}
 

### 2016-05-27 Paper/paper.tex
adding suggested additional figs

@@
 \usepackage{wrapfig}
+\usepackage{url}
 \usepackage{subfigure}
@@
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
-Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
-\vmcomment{The last sentence is very low on information content.}
+Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper makes a notable contribution to the statistical literature by illustrating how to directly embed knowledge of a physical system into a stochastic model.
  \begin{figure}[h!]
@@
 Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.  
-This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects being observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude.
+This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude.
 Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
@@
 Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
-Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood to reduce bias and in allowing for non-stationarity through time-varying parameter estimates.
+Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
 \par
@@
 	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
-	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.
-	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series. 
-	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations\vmadd{:} 	
+	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.
+	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately weighted complex exponentials. 
+	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations:
+	\hdcomment{these are wrong} 	
 	\begin{align}
-z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega\vmadd{,} && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
+z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega,&& f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
 \end{align}
@@
 This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
-Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.  
-For real $z(t)$, we define the autocovariance as $s_{z}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t).
-Letting $\tau$ be the time lag, we have
+Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
+We define the autocovariance  for $z(t)$ as $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
 \begin{align}
 \label{eq: fourierPair}
-S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
+S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
 \end{align}
 
-\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, $\hat{S}_{z}(\omega) = |J_{z}(\omega)|^{2}$ where 
+\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, where 
 \begin{equation}
 \label{eq: perio}
-J_{z}(\omega) = \sqrt{\frac{t}{N}} \sum_{t=1}^{N} z(t) e^{-i \omega t}
+\hat{S}_{z}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
 \end{equation}
 \citep{Sykulski2013}. 
-Although intuitively appealing, this estimator has known problems that stem from  sampling a periodic function at discrete intervals. 
-In particular,  high frequencies can  capture periodic behavior of other lower frequencies if they happen to align in period. 
-Additionally, frequencies above the highest observable frequency cannot be captured, and, are instead, recorded as other frequencies in the spectrum.  
-These effects, referred to as $\textbf{\it{aliasing}}$ and $\textbf{\it{leakage}}$ respectively,  are typically reduced by $\textbf{\it{tapering}}$.
-This method uses  a window of frequencies for estimation rather than a single value.
+Although intuitively appealing, this estimator has known problems that stem from  sampling a time series at discrete intervals. 
+\textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.  
+\hdcomment{Get this right!} Additionally, \textbf{\it{Leakage}} refers to the effects of because Fourier transforms are applied are a smooth, create discontinuities
+$\textbf{\it{Tapering}}$ reduces leaking by \hdcomment{get this right!}
 However, depending on how tapering is used, it can introduce its own  biases in estimation.
@@
  Working in the frequency domain, the two components of the model are added together, creating a six-parameter spectral model. 
- Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true spectral likelihood that balances computational efficiency with minimizing bias \citep{Sykulski2013}.
+ Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true Whittle likelihood that compensates for bias and handles aliasing \citep{Sykulski2013}.
 
@@
 \label{eq: deterOsc}
-\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu\vmadd{,} \\ \nonumber
-\frac{\partial v}{\partial t} - f_{0}u &= G - cv\vmadd{,}
+\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu,\\ \nonumber
+\frac{\partial v}{\partial t} - f_{0}u &= G - cv,
 \end{align}
@@
 \label{eq:diffEqDeriv}
-\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\ \nonumber
-dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\ \nonumber
-&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
-&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\ \nonumber
+\nonumber
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt} \\ \nonumber
+dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
+&= f_{0}(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
+&= if_{0}(u + iv)dt - c(u + iv)dt + (F + iG)dt\\ 
 &= (if_{0} - c)z dt + (F + iG)dt. 
 \end{align}
-\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation ~\ref{eq: deterOsc},
+\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
  \begin{equation}
@@
  The complex-valued OU process is the complex representation of a bivariate OU process.  
- In the formulation of Equation~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
+ In the formulation of Equation~\ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
  The final term of Equation~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
@@
 \item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
-\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
+\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $\tau$ for all $t, \tau > 0$
 \end{enumerate}
@@
 For time series modeling, we are interested in the autocovariance of this process at different time lags.
-To derive this value given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
+To derive the autocovariance as given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
 This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
@@
 \label{eq: expec}
-\mathbb{E}\{z(t) z^{*}(t + \tau)\} = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i\{\mathbb{E}[u(t) v(t + \tau)] + \mathbb{E}[u(t)v( t + \tau)]\}
+s(\tau) = \mathbb{E}[z(t) z^{*}(t + \tau)] = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i(\mathbb{E}[u(t) v(t + \tau)] -\mathbb{E}[u(t + \tau)v( t)])
 \end{equation}
-\citep{DeIaco2003} where $z^{*}(t + \tau)$ represents the complex conjugate of $z(t + \tau)$. The differential equations for components $u$ and $v$ can be expressed in matrix form as 
+with time lag $\tau >0$ \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
 \begin{align}
-Z(t) = \left( \begin{array}{c} u(t) \\ v(t) \end{array} \right) &= 
-\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q(t) \\ Ad Q(t) \end{array} \right).
+\left( \begin{array}{c} du(t) \\ dv(t) \end{array} \right) &= 
+\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q_{1}(t) \\ Ad Q_{2}(t) \end{array} \right).
 \end{align}
-For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0)$\vmadd{,} where $s(0)$ represents the covariance at zero lag \vmdel{and $B$ is the generator matrix} \citep{Schach1971}.
-\vmcomment{Generator is defined neither in general nor in your particular case.}
-Using the known variance of a univariate OU process and the independence of the components $u(0)$, we have
+For a bivariate OU process, the covariance matrix at time lag $\tau > 0$ is 
 \begin{align}
-s(0) = \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right).
+\label{eq: bivOU}
+C(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
+\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}C(0)
 \end{align}
-Thus, 
+\citep{Schach1971}.
+\citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$).
+We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$.
+Thus, Equation \ref{eq: expec} simplifies to
+\begin{equation}
+\label{eq: s0}
+s(0) = \frac{A^{2}}{2c} =  \mathbb{E}[u(t)^{2}] + \mathbb{E}[(v(t)^{2}]. 
+\end{equation}
+Since $u(t)$ and $v(t)$ are assumed to be identical processes, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation ~\ref{eq: bivOU} can be expressed as
 \begin{align}
 \label{eq: covDeriv}
-\mathbb{E} \left[\left( \begin{array}{c} u(t + \tau) \\ v(t + \tau)  \end{array} \right) \left( \begin{array}{cc} u(t) & v(t) \end{array} \right)  \right] = e^{B \tau} s(0) = 
-e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)\vmadd{,}
+ C(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
+\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right) 
 \end{align}
-where the second equality can be seen by expanding the infinite series representations for $\exp(-\lambda \tau)$, $\sin(\tau \omega)$, and $\cos( \tau \omega)$. 
-Using the relevant components of Equation ~\ref{eq: covDeriv} in Equation ~\ref{eq: expec} and applying Euler's formula, we obtain the complex OU process autocovariance as defined in \citet{Sykulski2016}
+where the second equality can be seen by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
+Finally, using the values in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
 \begin{align}
-\label{eq: ouAC}
-s^{(0)}(\tau) & = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
+\label{eq: ouAC} \nonumber
+s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}) \\
+&= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
 \end{align}
+ The $(o)$ superscript denotes that this equation is for the OU component of the full model.
 Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
 \label{eq:ouPSD}
-S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
+S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
 \end{align}
-The $(o)$ subscripts denotes that these equations are for the OU component of the full model.
 
@@
 The drifters' time series are also marked by the ocean's large-scale turbulence. 
-The physics of ocean turbulence is summarized \vmdel{in} \vmadd{by} \citet{Rhines1979}. 
+The physics of ocean turbulence is summarized by \citet{Rhines1979}. 
 Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. 
@@
  In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
- The Mat\'{e}rn model is defined either by its autocovariance \vmdel{or power spectral density}
+ The Mat\'{e}rn model is defined either by its autocovariance 
 \begin{align}
 \label{eq:maternAC}
-s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\vmadd{,}
+s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|),
 \end{align}
-\vmadd{or power spectral density,}
+or power spectral density,
 \begin{align}
 \label{eq:maternPSD}
-S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}
+S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}},
 \end{align}
-where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the s\vmdel{ub}\vmadd{uper}script $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\mathcal{K}_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the superscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
@@
         \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
-Top Left: The eastward component the velocities.
+Top Left: The eastward component of the velocities.
 Top Right: The eastward and northward components of the velocities.
@@
 \begin{align}
-S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}\\
 \label{eq: fullSpec}
-s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
+s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|)
+,\\
+S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}.
 \end{align}
@@
 \begin{align*}
-l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z\vmadd{,}
+l_{t}(\theta) = - \frac{1}{2} \log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z\vmadd{,}
 \end{align*}
@@
 
- It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
+ It is known that  the periodogram satisfies
 \begin{equation}
-|J_{Z}(\omega)| \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
+\hat{S}_{z}(\omega) \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
 \end{equation}
@@
 \begin{align}
-l_{\omega}(\theta) &= - \sum_{\omega \in \Omega} \left[ J_{Z}^{H}(\omega) \tilde{S}_{Z}^{-1}(\omega; \theta) J_{Z}(\omega) + \log \{S(\omega; \theta) \} \right]\\
-\label{eq: bwl}
-&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]\vmadd{,}
+l_{\omega}(\theta) 
+&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right],
 \end{align}
-where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix. 
-  This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}.
+where the subscript $w$ indicates that this is a Whittle likelihood. 
+Leveraging the fast Fourier transform \citep{Cooley1965}, this  likelihood can be evaluated in $\mathcal{O}(N log(N))$, where $N$ is the length of the time series. 
+This is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}. 
   Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section~\ref{sec: specAnalysis}.
 
-To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$\vmadd{, in Equation ~\ref{eq: bwl}} with the expected value of the spectrum, \vmdel{$\overline{S}_{Z}(\omega, \theta)$}
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$\, in Equation ~\ref{eq: bwl} with the expected value of the periodogram, 
 \begin{equation}
-\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)\vmadd{.}
+\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) \exp ( - i \omega \tau \Delta).
 \end{equation} 
-This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
+This form of the expected periodogram can be shown to be equivalent to convolving the theoretical spectrum with the Fej\'{e}r kernel 
 \begin{equation}
-\mathcal{F}(\cdot) = \frac{\Delta}{2\pi N} \frac{sin^{2}(N \omega \Delta/2)}{sin^{2}(\omega \Delta /2)} 
+\mathcal{F}(\cdot) = \frac{\Delta}{2\pi N} \frac{\sin^{2}(N \omega \Delta/2)}{\sin^{2}(\omega \Delta /2)} 
 \end{equation} \citep{Sykulski2013}. 
-This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram.
-Thereby, effects due to sampling at discrete intervals are avoided.
+\hdcomment{sentence wrong, look this up!} This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram.
+Also, aliasing effects due to sampling at discrete intervals are taken into account.
 This leads to the final likelihood
 \begin{align}
-l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]\vmadd{,}
+l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right],
 \end{align}
-which is maximized by standard \vmadd{off-the-shelf nonlinear} optimization techniques. 
-\vmcomment{State the exact method(s) that the paper and you used.}
+which is maximized by standard off-the-shelf nonlinear optimization techniques.
+(Specifically, \citet{Sykulski2016} uses the `fminsearchbnd' function in Matlab \citep{DErrico2006} and we use the `fminsearch' function in the `pracma' R package \citep{Borchers2015}.)
 
@@
 They refer to making these adjustments as applying a semi-parametric approach. 
-In their analysis, they highlight using knowledge of the physical system and adjusting for artifacts from interpolation. 
+In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
 Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations. 
 To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.
-Similarly, sampling effects make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
+Similarly, sampling effects due to interpolation make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
 Figure~\ref{fig: fig4} illustrates a semi-parametric model fit.
@@
 \subsubsection{Non-stationarity}
-Further, to allow for non-stationarity in long time series, \citet{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest.
-This \vmdel{gives} \vmadd{makes parameters in} the OU differential equation
+To allow for non-stationarity in long time series, \citet{Sykulski2016} propose fitting the model independently at each observed time point using a rolling window around the time point of interest. 
+This makes parameters in the OU differential equation
 \begin{equation}
@@
 \begin{equation}
-S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}\}^{2} + c(t)^{2}}\vmdel{.} 
+S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}(t)\}^{2} + c(t)^{2}}
 \end{equation}
-\vmdel{with parameters} now \vmadd{be} \vmdel{a} function\vmadd{s} of time. 
+ now be a functions of time. 
 Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
@@
 \end{equation}
-The length of the window to fit these time-varying parameters needs to be large enough to allow for a full sampling of the spectrum, but short enough to be treated as locally stationary. 
+The length of the window to fit these time-varying parameters needs to be large enough to provide good estimation of the spectrum, but short enough to be treated as locally stationary. 
  In this analysis, the authors select a reasonable window length empirically, but acknowledge that further work on this subject would be of value.
@@
 Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred. 
-To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f$, creating a five parameter model. 
+To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f_{0}$, creating a five parameter model. 
 Then this simpler model can be compared to the full six parameter model via a likelihood ratio test with test statistic
@@
 where $\hat{\theta}_{0}(t)$ is the fitted parameter vector for the simpler, or null model, and $\hat{\theta}_{1}(t)$ is the fitted parameter vector for the full model.
- The authors claim $R(t)$ then follow a Chi-squared distribution with 1 degree of freedom. 
+ The authors claim $R(t)$ then follows a chi-squared distribution with 1 degree of freedom. 
  However, we note that this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
@@
 First, we consider the modeled trajectories of 200 near-surface particles obtained from a computer simulation of ocean circulation, where each of the trajectories are recorded at two hour intervals. 
-The simulation is set to be `very similar' to the SPEM 5.2 set-up of \citet{Danioux2008}). 
+The simulation is set to be `very similar' to the SPEM 5.2 set-up of \citet{Danioux2008}. 
 In the  simulation, the turbulence is designed to be at a steady-state and wind forcing is set to be  constant. 
@@
 The parameters are assumed constant, since this portion of the time series does is fairly short. 
-In Figure ~\ref{fig: numSim}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
+In Figure ~\ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
 We also display the individual periodograms and model fits. 
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig5.png}
+    \includegraphics[width=.8\textwidth]{ReplicatedFigures/fig5.png}
         \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.png}
+    \includegraphics[width=.8\textwidth]{ReplicatedFigures/fig8.png}
         \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands. The red line in the top figure is the inertial frequency.}
@@
 We then compare this fit to the full model using a likelihood ratio test as described in Section ~\ref{sec: sig}.
- In Figure 9, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
+ In Figure ~\ref{fig: fig9}, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
 The authors conclude that when there are statistically significant differences between the 5-parameter and 6-parameter model as indicated by the likelihood ratio test,   the inertial frequency does not equal the Coriolis frequency.
@@
  Further, to model the turbulent background, the authors propose using the  Mat\'{e}rn autocovariance structure.
-  This flexible model cleverly encompasses all models that have previously been proposed for the turbulent background as special cases; thus, allowing the data to determine what is the most appropriate behavior.
-   They also employ the newly-developed Blurred Whittle likelihood for model fitting, which reduces biases compared to other fitting techniques. 
+  This flexible model cleverly encompasses all models that have previously been proposed for the turbulent background as special cases.
+  This allows the data to determine what is the most appropriate behavior. 
+   They also employ the newly-developed Blurred Whittle likelihood for model fitting, which compensates for bias and aliasing, unlike typical Whittle likelihood. 
    All together, this spectral model provides an automated way to detect shifts from the Coriolis frequency in a long time series. 
@@
  The authors do allude to a lack of correlation between the inertial frequency and other model components, but this idea is not fully addressed.
-Overall, though \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
+Overall, \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
  Their work is a substantial contribution to stochastic modeling, particularly, in terms, of explicitly incorporating known physics into statistical models.

### 2016-05-27 Code/figures.R

@@ -30,8 +30,9 @@ bStartTs <- 2971; bEndTs <-3570 #index of 50 day period in blue is obtained from
 gStartTs <- 4201; gEndTs <- 4800 #index of 50 day period in blue is obtained from Sykulski et al. posted Matlab code
 
 #plot time series, coloring particular sections 
-#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig1.pdf')
-#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(1.4,5,4,1.1)) #for pretty exporting
+#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig1.pdf',
+#    height = 3, width = 4)
+par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(5,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1, 1))
 plot(drift$lon[startTs:endTs], drift$lat[startTs:endTs], type = "l", 
      xlab = "Longitude", ylab = "Latitude", main = "Replication of Figure 1 (Right)")
@@ -131,7 +132,7 @@ fbmSamp <- simFBM(B = 10, alpha  = 0.9, H = 0.1, N = N)
 #pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig3.pdf',
 #    height = 6, width = 6)
 #par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4,5,4,1.1), mai = rep(.5, 4), oma = rep(1, 4)) #for pretty exporting
-#layout(matrix(c(1, 2, 3, 3), 2, 2, byrow = TRUE))
+layout(matrix(c(1, 2, 3, 3), 2, 2, byrow = TRUE))
 #u coordinate vs. time
 plot(fbmSamp[,"u"], col = "blue", xlim = c(0, 1000), ylim = c(-200, 200),
      type = "l", xlab = expression(paste(t, Delta)), ylab = "u (cm/s)")
@@ -163,7 +164,6 @@ DELTA <- 1 #constant sampling increment
 vel <- num[[6]] + 1i*num[[7]] #complex velocities, measured in cm/s
 vel1 <- vel[,1] #first simulation, pulled out for left figure
 
-###TO-DO: See if we can just use the 4pi value here
 #inertial frequency
 lat <- num[[10]] 
[…]

### 2016-05-27 Code/functions.R

@@ -246,23 +246,33 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   #Maximize likelihood numerically
   library("pracma")
   if (simpModel == FALSE) {
-    opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,firstIndex = firstIndex,
-                       lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
+    #use transformation if possible (since faster)
+    if (any(is.infinite(transParInit))) {
+      opt <- fminsearch(ll, parInit, delta = delta, curr = curr,firstIndex = firstIndex,
+                        lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000, trans = FALSE)
+      fin <- opt$xval
+    } else {
+      opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,firstIndex = firstIndex,
+                        lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
+      fin <-  c(expm1(opt$xval[1]) + 1, expm1(opt$xval[2]) + 1 , opt$xval[3], expm1(opt$xval[4]) + pi*sqrt(3)/N + 1, 
+                expm1(opt$xval[5]) + pi*sqrt(3)/N + 1, expm1(opt$xval[6]) + 0.5 + 1)
+    }
   } else {
-    opt <- fminsearch(llSimp, transParInit, delta = delta, curr = curr, 
-                      firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
-                      medIndex = medIndex, maxiter = 3000)
+    if (any(is.infinite(transParInit))) {
+      opt <- fminsearch(llSimp, parInit, delta = delta, curr = curr, trans = FALSE,
+                        firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
+                        medIndex = medIndex, maxiter = 3000)
+      fin <- opt$xval
+    } else {
+      opt <- fminsearch(llSimp, transParInit, delta = delta, curr = curr,
+                        firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
+                        medIndex = medIndex, maxiter = 3000)
[…]

### 2016-05-28 Paper/paper.tex
more paper edits

@@
 \section{Introduction}
-\citet{Sykulski2016} present\vmdel{s} a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
+\citet{Sykulski2016} present a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
 Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
-Figure~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
+Figure \ref{fig: fig1} shows a sample 250-day drifter trajectory. 
 Oceanographers use these time series to increase understanding of  ocean circulation patterns.
@@
 Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
-These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions of rotation in the two hemispheres.
+These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions objects rotate in the two hemispheres.
  Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
@@
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.
-	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately weighted complex exponentials. 
+	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately-weighted complex exponentials. 
 	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations:
@@
 \begin{equation}
-S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
+S(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
 \end{equation}
@@
 Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
-We define the autocovariance  for $z(t)$ as $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
+We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
 \begin{align}
 \label{eq: fourierPair}
-S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
+S(\omega) = \int_{-\infty}^{\infty}s(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S(\omega) e^{i \omega t} d\omega 
 \end{align}
@@
 \label{eq: perio}
-\hat{S}_{z}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
+\hat{S}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
 \end{equation}
@@
 \subsection{Stochastic Model}
- To model a drifter's movement over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, is converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
+ To model a drifter's movement over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$.
  Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior. 
@@
 \end{equation}  
-Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
+Equation \ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
 In general,  OU processes are considered continuous-time equivalents to the AR(1) process.
-Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate Markov process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
+Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
  The complex-valued OU process is the complex representation of a bivariate OU process.  
- In the formulation of Equation~\ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
- The final term of Equation~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
+ In the formulation of Equation \ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
+ The final term of Equation \ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
 The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation \ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
 For time series modeling, we are interested in the autocovariance of this process at different time lags.
-To derive the autocovariance as given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
-This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
+To derive the autocovariance as given in Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
+This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ with time lag $\tau >0$ is defined as 
 \begin{equation}
@@
 \end{equation}
-with time lag $\tau >0$ \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
+ \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
 \begin{align}
@@
 \label{eq: bivOU}
-C(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
-\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}C(0)
+\boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
+\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)
 \end{align}
-\citep{Schach1971}.
+where $\boldsymbol{C}(0)$ gives the covariance matrix at the initial time, $t$ \citep{Schach1971}.
 \citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$).
@@
 \end{equation}
-Since $u(t)$ and $v(t)$ are assumed to be identical processes, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation ~\ref{eq: bivOU} can be expressed as
+Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \ref{eq: bivOU} can be expressed as
 \begin{align}
 \label{eq: covDeriv}
- C(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
+ \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &=
 \frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right) 
@@
 where the second equality can be seen by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
-Finally, using the values in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
+Finally, using the values obtained in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
 \begin{align}
@@
  The $(o)$ superscript denotes that this equation is for the OU component of the full model.
-Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
+Using Equation \ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
 \label{eq:ouPSD}
@@
 \end{align}
-where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\mathcal{K}_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the superscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}.
+where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\mathcal{K}_{\eta}$ is a modified Bessel function of the second kind of order $\eta$ \citep{Stein2012}.
+The superscript $(m)$ indicates that this is the Mat\'{e}rn part of the full model.
 The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
@@
 The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. 
-This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values. 
+This means that these models can still be obtained under the current model if model fitting leads to their parameter estimate values. 
 They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
- Figure~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
+ Figure \ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.
 
@@
  To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
- Using Equations ~\ref{eq: ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
+ Using Equations \ref{eq: ouAC} and \ref{eq:maternAC} and Equations \ref{eq:ouPSD} and\ref{eq:maternPSD}, we obtain
 \begin{align}
 \label{eq: fullSpec}
-s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|)
+s(\tau ;\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|)
 ,\\
@@
     \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
-        \caption{Periodograms of the time series sections colored blue and green in Figure ~\ref{fig: fig1}.
+        \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}.
 The black line denotes the inertial frequency.
@@
  Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost.
-  Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $Z$,  in the time domain that has the likelihood
+  Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $\boldsymbol{Z}$,  in the time domain that has the likelihood
 \begin{align*}
-l_{t}(\theta) = - \frac{1}{2} \log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z\vmadd{,}
+l_{t}(\boldsymbol{\theta}) = - \frac{1}{2} \log \left|\boldsymbol{\Sigma}_{Z}(\boldsymbol{\theta}) \right| - \frac{1}{2} Z^{T} \boldsymbol{\Sigma}_{\boldsymbol{Z}}^{-1}(\boldsymbol{\theta})\boldsymbol{Z},
 \end{align*}
-where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.
-\vmcomment{I should have mentioned this earlier in the test that all vectors and matrices need to be bolded.}
+where $\boldsymbol{\theta}$ are the model parameters for a particular choice of covariance structure, $\boldsymbol{\Sigma}_{z}$,  and $t$ denotes the time domain  \citep{Sykulski2013}.
 
@@
 \begin{equation}
-\hat{S}_{z}(\omega) \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
+\hat{S}(\omega) \overset{d}{\rightarrow} S(\omega) \chi^{2}_{2}/2
 \end{equation}
@@
 \begin{align}
-l_{\omega}(\theta) 
-&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right],
+\label{eq: bwl}
+l_{\omega}(\boldsymbol{\theta}) 
+&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}(\omega)}{S(\omega;\boldsymbol{\theta})}  + \log  \{ S(\omega; \boldsymbol{\theta}) \}\right],
 \end{align}
@@
 This is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}. 
-  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section~\ref{sec: specAnalysis}.
+  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section \ref{sec: specAnalysis}.
 
-To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$\, in Equation ~\ref{eq: bwl} with the expected value of the periodogram, 
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S(\omega; \theta)$\, in Equation \ref{eq: bwl} with the expected value of the periodogram, 
 \begin{equation}
-\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) \exp ( - i \omega \tau \Delta).
+\overline{S}(\omega; \boldsymbol{\theta}) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s(\tau ; \boldsymbol{\theta}) \exp ( - i \omega \tau \Delta).
 \end{equation} 
@@
 \end{equation} \citep{Sykulski2013}. 
-\hdcomment{sentence wrong, look this up!} This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram.
+This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram \hdcomment{sentence wrong, look this up!}.
 Also, aliasing effects due to sampling at discrete intervals are taken into account.
@@
 \begin{align}
-l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right],
+l_{b}(\boldsymbol{\theta}) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}(\omega)}{\overline{S}(\omega; \boldsymbol{\theta})} + \log \{ \overline{S} (\omega; \boldsymbol{\theta}) \}\right],
 \end{align}
-which is maximized by standard off-the-shelf nonlinear optimization techniques.
+where $b$ denotes blurred.
+This likelihood is maximized by standard off-the-shelf nonlinear optimization techniques.
 (Specifically, \citet{Sykulski2016} uses the `fminsearchbnd' function in Matlab \citep{DErrico2006} and we use the `fminsearch' function in the `pracma' R package \citep{Borchers2015}.)
@@
 \label{eq: fisher}
-\sqrt{N}(\hat{\theta} - \theta) \overset{d}{\rightarrow} \text{ Normal($0, \mathcal{F}^{-1})$ }\vmdel{.}\vmadd{,}
+\sqrt{N}(\hat{\boldsymbol{\theta}} - \boldsymbol{\theta}) \overset{d}{\rightarrow} \text{ Normal($0, \boldsymbol{\mathcal{F}}^{-1})$ },
 \end{equation}
-where $\mathcal{F}$ is the Fisher information. 
+where $\boldsymbol{\mathcal{F}}$ is the Fisher information. 
 Thus, uncertainty estimates of the parameters can easily be obtained using numerical calculation of the Hessian.
@@
 \label{sec: semi}
-The authors note that the fitting of Equation~\ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
+The authors note that the fitting of Equation \ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
 They refer to making these adjustments as applying a semi-parametric approach. 
@@
 Similarly, sampling effects due to interpolation make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
-Figure~\ref{fig: fig4} illustrates a semi-parametric model fit.
+Figure \ref{fig: fig4} illustrates a semi-parametric model fit.
 
@@
 \end{equation}
-and spectral power density
+and power spectral density
 \begin{equation}
-S^{(0)}(\omega, t) = \frac{A(t)^{2}}{\{\omega - \omega_{0}(t)\}^{2} + c(t)^{2}}
+S^{(0)}(\omega, t) = \frac{A(t)^{2}}{(\omega - \omega_{0}(t))^{2} + c(t)^{2}}
 \end{equation}
  now be a functions of time. 
-Similarly, the time-varying version of Mat\'{e}rn power spectral density is 
+Similarly, the time-varying version of the Mat\'{e}rn power spectral density is 
 \begin{equation}
-S^{(m)}(\omega, t) = \frac{B^{2}(t)}{\{\omega^{2} + h^{2}(t)\}^{\alpha(t)}}.
+S^{(m)}(\omega, t) = \frac{B^{2}(t)}{(\omega^{2} + h^{2}(t))^{\alpha(t)}}.
 \end{equation}
@@
 Finally, \citet{Sykulski2016} propose that a simple likelihood ratio test can be used at each time period to test whether or not a shift in the inertial oscillation has occurred. 
-To do so, the authors suggest fitting a simpler model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f_{0}$, creating a five parameter model. 
+To do so, the authors suggest fitting a simpler five parameter model where the the parameter $\omega_{0}$ is fixed to be the inertial frequency $f_{0}$. 
 Then this simpler model can be compared to the full six parameter model via a likelihood ratio test with test statistic
@@
 \label{eq: LRT}
-R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ]\vmadd{,}
+R(t) = 2[l_{b} \{\hat{\theta}_{1}(t) \} - l_{b}\{\hat{\theta}_{0}(t) \} ],
 \end{equation}
@@
  The authors claim $R(t)$ then follows a chi-squared distribution with 1 degree of freedom. 
- However, we note that this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
-\vmcomment{I don't think the lack of independence between time points has an impact on the LRT, but using multiple LRTs without multiple testing correction, which is complicated by non-independence, is the main problem here, I think.}
+ However, as we will discuss further this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
 
@@
 
-\par For all two hundred trajectories, the full six-parameter model, given in Equation ~\ref{eq: fullSpec}, is fit to each trajectory. 
+\par For all two hundred trajectories, the full six-parameter model, given in Equation \ref{eq: fullSpec}, is fit to each trajectory. 
 The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
 The parameters are assumed constant, since this portion of the time series does is fairly short. 
-In Figure ~\ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
+In Figure \ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
 We also display the individual periodograms and model fits. 
@@
 We also consider the trajectory of a drifter in the south Pacific ocean observed 12 times a day over a 1642 day period.  
-The trajectory is displayed in the right panel of Figure ~\ref{fig: fig6}. 
-We first considering fitting a 20-day sample of this trajectory. 
-Its periodogram, displayed in ~\ref{fig: fig6}, shows substantial asymmetry between the positive and negative frequencies.
+The trajectory is displayed in the left panel of Figure \ref{fig: fig6}. 
+We first consider fitting a 20-day sample of this trajectory. 
+Its periodogram, displayed in Figure \ref{fig: fig6}, shows substantial asymmetry between the positive and negative frequencies.
 Thus, the semi-parametric approach is applied and only positive frequencies are used in fitting the model.
@@
 \par
-Moving on to the entire time series, we use rolling windows of 1000 observations and the semi-parametric set-up described in Section ~\ref{sec: semi}  to obtain parameter estimates for the spectral model at each time point.
- In Figure ~\ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density in a time-frequency plot. 
+Moving on to the entire time series, we use rolling windows of 1000 observations and the semi-parametric set-up described in Section \ref{sec: semi}  to obtain parameter estimates for the spectral model at each time point.
+ In Figure \ref{fig:timeVarying}, we compare the observed spectral density over time and the fitted spectral density in a time-frequency plot. 
   The fitted density appears to be a smoothed version of the observed density,  indicating a reasonable model fit.
@@
   \centering
-    \includegraphics[width=1\textwidth]{ReplicatedFigures/fig7.png}
+    \includegraphics[width=.9\textwidth]{ReplicatedFigures/fig7.png}
         \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
-On both figures, the white line indicates the average inertial frequency within the rolling window at the current time period.
+On both figures, the semi-transparent gray line indicates the average inertial frequency within the rolling window at the current time period.
 Only frequencies used in the estimation are included.
@@
  \par
- We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure ~\ref{fig: fig8}. 
+ We also display the fitted parameter values over time and their 95$\%$ confidence interval bands in Figure \ref{fig: fig8}. 
 The 95$\%$ intervals are obtained using numerical calculation of the Hessian.
@@
 \par We also fit the 5-parameter model where the frequency is set to the inertial frequency rather than being a free-parameter. 
-We then compare this fit to the full model using a likelihood ratio test as described in Section ~\ref{sec: sig}.
- In Figure ~\ref{fig: fig9}, we display the test statistic over time where a red line is used to indicate the level of statistical significance. 
-The authors conclude that when there are statistically significant differences between the 5-parameter and 6-parameter model as indicated by the likelihood ratio test,   the inertial frequency does not equal the Coriolis frequency.
+We then compare this fit to the full model using a likelihood ratio test as described in Section \ref{sec: sig}.
+ In Figure \ref{fig: fig9}, we display the test statistics over time where a red line is used to indicate the level of statistical significance. 
+The authors conclude that when there are statistically significant differences between the 5-parameter and 6-parameter model, as indicated by the likelihood ratio test,   the inertial frequency does not equal the Coriolis frequency.
 They note that the period from days 600 to 900 always shows a significant difference between the models, which corresponds to the period of shifts observed in Figure \ref{fig: fig8}.
@@
 
-\par Finally, in Figure ~\ref{fig: fig10}, we display a matrix of the average correlation of the parameters over time. \citet{Sykulski2016} highlight several useful results from this table. 
+\par Finally, in Figure \ref{fig: fig10}, we display a matrix of the average correlation of the parameters over time. \citet{Sykulski2016} highlight several useful results from this table. 
 They note that the inertial frequency is generally uncorrelated with other parameters values, suggesting that the inertial frequency only controls the location of the peak, but does not influence the shape or amplitude of the spectra as desired.
@@
  In particular, they develop a two component spectral model that jointly accounts for the turbulent background behavior of the ocean and the influence of inertial oscillations. 
- Using the complex-valued Ornstein-Uhlenbeck process, the authors develop a stochastic analogue to the accepted deterministic differential equations used to describe this phenomena. 
+ Using the complex-valued OU process, the authors develop a stochastic analogue to the accepted deterministic differential equations used to describe this phenomena. 
  In so doing, they are the first to explicitly model inertial oscillations within the context of a stochastic model. 
@@
 This provide an efficient way to detect eddies for large numbers and/or long time series. 
-So, the authors have both satisfied their scientific goal and have introduced new statistical ideas in doing so.
+So, the authors have  satisfied their scientific, and, in so doing, introduced new statistical ideas.
 
@@
 Another area that is insufficiently addressed in the paper is possible identifiability issues between the model parameters. 
-Since primary interest is in the inertial frequency, if parameter estimates are highly correlated, physical changes in one parameter might actually be reflected by changes in other fitted parameter.
+ If parameter estimates are highly correlated, physical changes in one parameter might actually be reflected by changes in other fitted parameters.
   This would subsequently invalidate conclusions about observed changes in specific parameters, since they could be masked by changes in other parameters. 
- The authors do allude to a lack of correlation between the inertial frequency and other model components, but this idea is not fully addressed.
+ For example, in Figure \ref{fig: fig10}, we see a high correlation (0.74) between the OU dampening and amplitude parameters, suggesting caution should be taken in interpreting their physical meaning.
+ For the primary variable of interest, inertial frequency, all correlations with other parameters are low, suggesting that the main conclusions of the paper are appropriate. 
+Nevertheless, a fuller discussion of these limitations in interpretation would have been valuable.
 Overall, \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
@@
     \includegraphics[width=\textwidth]{ReplicatedFigures/fig2.pdf}
-        \caption{Periodograms of the highlighted green and blue sections of the trajectories in Figure ~\ref{fig: fig1}. The inertial frequencies are marked in red. }
+        \caption{Periodograms of the highlighted green and blue sections of the trajectories in Figure \ref{fig: fig1}. The inertial frequencies are marked in red. }
         	\label{fig: fig2}

### 2016-05-28 Code/figures.R

@@ -556,7 +556,39 @@ for (i in 1001:(N - 1000)) {
 }
 
 #store  results
-save(llVal6_500, file = '/users/hdirector/Documents/prelim/prelim/Code/llVal6_500.rda')
+#save(llVal6_2000, file = '/users/hdirector/Documents/prelim/prelim/Code/llVal6_2000.rda')
+
+###Simplified model using longer window
+#Periodogram and paremeter fit for first rolling window (used for set up)
+toStartPer <- getPerio(driftUlys$cv[1:2000], DELTA, dB = TRUE, noZero = TRUE) 
+toStartFit <- fitModel(driftUlys$cv[1:2000], mean(CFVec[1]),  DELTA, fracNeg = 0, fracPos = fracPos,
+                       quantSet = 0.8, incZero = FALSE, getHess = TRUE, simpModel = TRUE)
+firstIndex <- toStartFit$firstIndex
+lastIndex <- toStartFit$lastIndex
+
+#Storage vectors for parameters, negative log likelihood, confidence interval bands, and hessians
+par5Val_2000 <- matrix(nrow = N - 1999, ncol = 5)
+colnames(par5Val_2000) <- c("A", "B", "C", "h", "alpha")
+llVal5_2000 <- rep(NA, N - 1999)
+llVal5_2000[1] <- toStartFit$llVal
+
+#store data for first window 
+par5Val_2000[1, ] <- c(toStartFit$A, toStartFit$B, 
+                       toStartFit$C, toStartFit$h, toStartFit$alpha)
+
+#Loop through all windows and calculate parameter estimates, confidence intervals, and hessian
+for (i in 1001:(N - 1000)) {
+  currCv <- driftUlys$cv[(i - 999):(i + 1000)] 
+  CFCurr <- CFVec[i - 999]
+  #fit parameters, initialize estimates at value used in last run
+  tempFit <- fitModel(currCv, CFCurr, DELTA, fracNeg = 0, fracPos = fracPos,
+                      quantSet = .8, needInits = FALSE, parInit = par5Val_2000[i - 999 - 1, ],
+                      getHess = FALSE, simpModel = TRUE)
+  par5Val_2000[i - 999, ] <- c(tempFit$A, tempFit$B,
+                               tempFit$C, tempFit$h, tempFit$alpha)
+  llVal5_2000[i - 999] <-  tempFit$llVal
+  print(i)
+}
 
 
 

### 2016-05-29 Paper/paper.tex
more or less done

@@
         \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue.
-The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
+The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.}
         	\label{fig: fig1}
@@
 	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
-	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
+	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.
 	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately-weighted complex exponentials. 
-	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations:
-	\hdcomment{these are wrong} 	
+	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations,
 	\begin{align}
-z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega,&& f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
+Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
 \end{align}
-\citep{Percival1993}. 
+where $T$ indicates the period \citep{Percival1993}. 
 To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
 \begin{equation}
-S(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
+S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
 \end{equation}
@@
 \label{eq: fourierPair}
-S(\omega) = \int_{-\infty}^{\infty}s(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S(\omega) e^{i \omega t} d\omega 
+S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
 \end{align}
 
-\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, where 
+\noindent \citep{Percival1993}. For an observed time series, the power spectral density is often estimated with the periodogram, where 
 \begin{equation}
 \label{eq: perio}
-\hat{S}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
+\hat{S}_{z}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t \Delta} \right)^{2}
 \end{equation}
@@
 Although intuitively appealing, this estimator has known problems that stem from  sampling a time series at discrete intervals. 
-\textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.  
-\hdcomment{Get this right!} Additionally, \textbf{\it{Leakage}} refers to the effects of because Fourier transforms are applied are a smooth, create discontinuities
-$\textbf{\it{Tapering}}$ reduces leaking by \hdcomment{get this right!}
-However, depending on how tapering is used, it can introduce its own  biases in estimation.
+Specifically, \textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.
+\textbf{\it{Leakage}} describes how as a result of convolution, power from one part of the spectral density function can be transferred to another part of the density.
 
@@
 In general,  OU processes are considered continuous-time equivalents to the AR(1) process.
-Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
+Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
  The complex-valued OU process is the complex representation of a bivariate OU process.  
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation \ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation ~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
   \centering
-    \includegraphics[width=.7\textwidth]{ReplicatedFigures/fig3.pdf}
+    \includegraphics[width=.9\textwidth]{ReplicatedFigures/fig3.pdf}
         \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
@@
 Top Right: The eastward and northward components of the velocities.
-Bottom: Periodograms of both processes on the decibel scale.
-\vmcomment{I would include legends or direct labeling for all your figures.}}
+Bottom: Periodograms of both processes on the decibel scale.}
         	\label{fig: fmbMat}
@@
  To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
- Using Equations \ref{eq: ouAC} and \ref{eq:maternAC} and Equations \ref{eq:ouPSD} and\ref{eq:maternPSD}, we obtain
+ Using Equations \ref{eq: ouAC} and \ref{eq:maternAC} and Equations \ref{eq:ouPSD} and \ref{eq:maternPSD}, we obtain
 \begin{align}
@@
 
-\begin{figure}[h!]
-  \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
-        \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}.
-The black line denotes the inertial frequency.
-The dotted lines on the left figure indicate the location of the eddy peak and the expected shift in the inertial frequency.
-The semi-parametric model fit for the frequencies used in model fitting is highlighted green on the left figure and blue on the right figure.
-The red dotted line indicates the model fit extended beyond the frequencies where it was fit.
-}
-        	\label{fig: fig4}
-\end{figure}
-
 \subsection{Fitting the Model}
@@
 \begin{align*}
-l_{t}(\boldsymbol{\theta}) = - \frac{1}{2} \log \left|\boldsymbol{\Sigma}_{Z}(\boldsymbol{\theta}) \right| - \frac{1}{2} Z^{T} \boldsymbol{\Sigma}_{\boldsymbol{Z}}^{-1}(\boldsymbol{\theta})\boldsymbol{Z},
+l_{t}(\boldsymbol{\theta}) = - \frac{1}{2} \log \left|\boldsymbol{\Sigma}_{z}(\boldsymbol{\theta}) \right| - \frac{1}{2} Z^{T} \boldsymbol{\Sigma}_{z}^{-1}(\boldsymbol{\theta})\boldsymbol{Z},
 \end{align*}
@@
 \begin{equation}
-\hat{S}(\omega) \overset{d}{\rightarrow} S(\omega) \chi^{2}_{2}/2
+\hat{S}_{z}(\omega) \overset{d}{\rightarrow} S(\omega) \chi^{2}_{2}/2
 \end{equation}
@@
 l_{\omega}(\boldsymbol{\theta}) 
-&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}(\omega)}{S(\omega;\boldsymbol{\theta})}  + \log  \{ S(\omega; \boldsymbol{\theta}) \}\right],
+&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{z}(\omega)}{S(\omega;\boldsymbol{\theta})}  + \log  \{ S(\omega; \boldsymbol{\theta}) \}\right],
 \end{align}
@@
 
-To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S(\omega; \theta)$\, in Equation \ref{eq: bwl} with the expected value of the periodogram, 
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S(\omega; \boldsymbol{\theta})$\, in Equation \ref{eq: bwl} with the expected value of the periodogram, 
 \begin{equation}
@@
 \end{equation} \citep{Sykulski2013}. 
-This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram \hdcomment{sentence wrong, look this up!}.
-Also, aliasing effects due to sampling at discrete intervals are taken into account.
+This approach improves the estimators obtained through maximizing the likelihood, by bypassing the problems, namely aliasing and leakage,  associated with comparison against the true spectrum. 
 This leads to the final likelihood
 \begin{align}
-l_{b}(\boldsymbol{\theta}) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}(\omega)}{\overline{S}(\omega; \boldsymbol{\theta})} + \log \{ \overline{S} (\omega; \boldsymbol{\theta}) \}\right],
+l_{b}(\boldsymbol{\theta}) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{z}(\omega)}{\overline{S}(\omega; \boldsymbol{\theta})} + \log \{ \overline{S} (\omega; \boldsymbol{\theta}) \}\right],
 \end{align}
-where $b$ denotes blurred.
+where $b$ denotes that this is the blurred Whittle likelihood.
 This likelihood is maximized by standard off-the-shelf nonlinear optimization techniques.
@@
 Figure \ref{fig: fig4} illustrates a semi-parametric model fit.
+
+\begin{figure}[h!]
+  \centering
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
+        \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}.
+The black line denotes the inertial frequency.
+The dotted lines on the left figure indicate the location of the eddy peak and the expected shift in the inertial frequency.
+The semi-parametric model fit for the frequencies used in model fitting is highlighted green on the left figure and blue on the right figure.
+The red dotted line indicates the model fit extended beyond the frequencies where it was fit.
+}
+        	\label{fig: fig4}
+\end{figure}
 
@@
  The authors claim $R(t)$ then follows a chi-squared distribution with 1 degree of freedom. 
- However, as we will discuss further this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
+ However, as we will discuss further, this may be overly simplistic, because of potential identifiability issues with the parameters and lack of independence between time points stemming from the use of a rolling window.
 
@@
   \centering
-    \includegraphics[width=.8\textwidth]{ReplicatedFigures/fig5.png}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig5.png}
         \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
@@
   \centering
-    \includegraphics[width=.9\textwidth]{ReplicatedFigures/fig7.png}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig7.png}
         \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
@@
   \centering
-    \includegraphics[width=.8\textwidth]{ReplicatedFigures/fig8.png}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.png}
         \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands. The red line in the top figure is the inertial frequency.}
@@
   \centering
-    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig9.pdf}
+    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig9Extend.pdf}
         \caption{Time series of the likelihood ratio test statistic comparing the 5-parameter to 6-parameter models.
 The dotted red line indicates the value at which test statistics indicate a significant difference between the model.
-}
+The first figure uses the window size specified in \citet{Sykulski2016} and the second figure doubles the window size.}
         	\label{fig: fig9}
@@
 This suggests that the p-value for any time point could be easily influenced by the length of the window used in fitting it.
-Thus, robustness checks of this method to the length of the window would be appropriate prior to using it.
-In other words, further thought should go in to designing a hypothesis that accurately reflects the problem set-up and modeling.
+To explore this, in Figure \ref{fig: fig9} we replicate Figure 9  in the paper using both the window length specified in the paper and a doubled window length.
+The corresponding likelihood ratio tests do not reject the null hypothesis at all the same times. 
+This indicates that window size and its influence on hypothesis testing should be further investigated prior to applying this formal testing approach for the presence of a shift from the Coriolis frequency.
 
@@
         \caption{Table of the average correlation between parameters.
-Note that the parameters are ordered differently than in the paper, due to differences in the structure of the author's and my code.
-There are also minor differences in the correlation values, particularly for values close to zero.
-These differences can be attributed to slight differences in the numerical optimization used by R and Matlab in calculating the Hessian.
+Note that there are  minor differences in the correlation values between these results and those in \citet{Sykulski2016} These differences can be attributed to slight differences in the numerical optimization used by R and Matlab in calculating the Hessian.
 }
@@
  For example, in Figure \ref{fig: fig10}, we see a high correlation (0.74) between the OU dampening and amplitude parameters, suggesting caution should be taken in interpreting their physical meaning.
- For the primary variable of interest, inertial frequency, all correlations with other parameters are low, suggesting that the main conclusions of the paper are appropriate. 
+ For the primary variable of interest, inertial frequency, all correlations with other parameters are low, so the main conclusions of the paper are not affected by this issue. 
 Nevertheless, a fuller discussion of these limitations in interpretation would have been valuable.
-Overall, \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
+\par Overall, \citet{Sykulski2016} present a novel method for modeling ocean surface drifters.
  Their work is a substantial contribution to stochastic modeling, particularly, in terms, of explicitly incorporating known physics into statistical models.

### 2016-05-29 Code/figures.R

@@ -32,12 +32,13 @@ gStartTs <- 4201; gEndTs <- 4800 #index of 50 day period in blue is obtained fro
 #plot time series, coloring particular sections 
 #pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig1.pdf',
 #    height = 3, width = 4)
-par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(5,5,4,1.1)) #for pretty exporting
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(5,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1, 1))
 plot(drift$lon[startTs:endTs], drift$lat[startTs:endTs], type = "l", 
      xlab = "Longitude", ylab = "Latitude", main = "Replication of Figure 1 (Right)")
 points(drift$lon[gStartTs:gEndTs], drift$lat[gStartTs:gEndTs], type = "l", col = "green")
 points(drift$lon[bStartTs:bEndTs], drift$lat[bStartTs:bEndTs], type = "l", col = "blue")
+legend("bottomright", legend = c("exhibits eddies", "no major eddies"), fill = c("blue", "green"), cex = .6)
 #dev.off()
 
 ###Figure 2###
@@ -68,10 +69,16 @@ par(mfrow = c(1, 2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
 abline(v = bCf, col = "red")
+legend("topright", legend = c("data", "inertial freq."),
+       lty = c("solid", "solid"), col = c("blue", "red"),
+       cex = .7, bg = "white")
 plot(DELTA*green$omega, green$sZ, type = "l", col = "green", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi), ylim = c(-20, 60))
 mtext("Replication of Figure 2", outer = T, line = -3)
 abline(v = gCf, col = "red")
+legend("topright", legend = c("data", "inertial freq."),
+       lty = c("solid", "solid"), col = c("green", "red"),
+       cex = .7, bg = "white")
 #dev.off()
 
 ###Figure 4###
[…]

### 2016-05-29 Code/functions.R

@@ -159,6 +159,7 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
     tempIndex <- tempIndex[-which(tempIndex == medIndex)]
     llVal <- sum(curr$sZ[tempIndex]/sBar[tempIndex]) + sum(log(sBar[tempIndex]))
   }
+  print(llVal)
   return(llVal)
 }
 
@@ -246,33 +247,23 @@ fitModel <- function(Z, CF, delta, fracNeg, fracPos, quantSet, incZero = FALSE, 
   #Maximize likelihood numerically
   library("pracma")
   if (simpModel == FALSE) {
-    #use transformation if possible (since faster)
-    if (any(is.infinite(transParInit))) {
-      opt <- fminsearch(ll, parInit, delta = delta, curr = curr,firstIndex = firstIndex,
-                        lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000, trans = FALSE)
-      fin <- opt$xval
-    } else {
-      opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,firstIndex = firstIndex,
-                        lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
-      fin <-  c(expm1(opt$xval[1]) + 1, expm1(opt$xval[2]) + 1 , opt$xval[3], expm1(opt$xval[4]) + pi*sqrt(3)/N + 1, 
-                expm1(opt$xval[5]) + pi*sqrt(3)/N + 1, expm1(opt$xval[6]) + 0.5 + 1)
-    }
+    opt <- fminsearch(ll, transParInit, delta = delta, curr = curr,firstIndex = firstIndex,
+                       lastIndex = lastIndex, medIndex = medIndex, maxiter = 3000) 
   } else {
-    if (any(is.infinite(transParInit))) {
-      opt <- fminsearch(llSimp, parInit, delta = delta, curr = curr, trans = FALSE,
-                        firstIndex = firstIndex, lastIndex = lastIndex, CF = CF,
-                        medIndex = medIndex, maxiter = 3000)
-      fin <- opt$xval
-    } else {
-      opt <- fminsearch(llSimp, transParInit, delta = delta, curr = curr,
[…]

### 2016-06-01 Paper/paper.tex
whole manuscript edits

@@
         \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue.
-The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.}
+The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.
+        \vmcomment{You could crop away the white space in the figure.}}
         	\label{fig: fig1}
@@
   However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
- The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citet{Isern2003} or \citet{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
+ The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citep{Isern2003} or \citep{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
 \citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
@@
 Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
-Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
+Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \vmadd{process} \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
 Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
@@
 	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
-	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
+	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.
@@
 	\begin{align}
-Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
+Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt\vmadd{,} && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
 \end{align}
@@
 Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
-We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
+We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $\vmadd{,} where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
 \begin{align}
@@
 We begin by discussing the model for the inertial oscillations. 
-In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
+In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations\vmadd{:}
 \begin{align}
@@
 where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}.
-Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
+Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows\vmadd{:}
 \begin{align}
@@
 \nonumber
-\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt} \\ \nonumber
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\vmadd{,} \\ \nonumber
 dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
@@
 \end{align}
-\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
+\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$\vmadd{,}  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
  \begin{equation}
@@
 \end{equation}  
-Equation \ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
+Equation \eqref{eq: ouEq}  is known as the complex-valued \vmdel{Ornstein-Uhlenbeck (}OU\vmdel{)} stochastic differential equation. 
 In general,  OU processes are considered continuous-time equivalents to the AR(1) process.
@@
  The complex-valued OU process is the complex representation of a bivariate OU process.  
- In the formulation of Equation \ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
- The final term of Equation \ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
+ In the formulation of Equation \eqref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
+ The final term of Equation \eqref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.
 The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation ~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated \vmdel{by} \vmadd{from} the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
 \boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
-\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)
+\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)\vmadd{,}
 \end{align}
@@
 \end{equation}
-Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \ref{eq: bivOU} can be expressed as
+Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \eqref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \eqref{eq: bivOU} can be expressed as
 \begin{align}
@@
  \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &=
-\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right) 
+\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right)\vmadd{,}
 \end{align}
-where the second equality can be seen by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
-Finally, using the values obtained in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
+where the second equality can be \vmdel{seen} \vmadd{obtained} by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
+Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined \vmdel{in} \vmadd{by} \citet{Sykulski2016}\vmadd{:}
 \begin{align}
 \label{eq: ouAC} \nonumber
-s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}) \\
+s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0})\vmadd{,} \\
 &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
@@
  The $(o)$ superscript denotes that this equation is for the OU component of the full model.
-Using Equation \ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
+Using Equation \eqref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
 \label{eq:ouPSD}
@@
 Top Right: The eastward and northward components of the velocities.
-Bottom: Periodograms of both processes on the decibel scale.}
+Bottom: Periodograms of both processes on the decibel scale.
+\vmcomment{Label all axes.}}
         	\label{fig: fmbMat}
@@
  To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
- Using Equations \ref{eq: ouAC} and \ref{eq:maternAC} and Equations \ref{eq:ouPSD} and \ref{eq:maternPSD}, we obtain
+ Using Equations \eqref{eq: ouAC} and \eqref{eq:maternAC} and Equations \eqref{eq:ouPSD} and \eqref{eq:maternPSD}, we obtain
 \begin{align}
@@
 
-To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S(\omega; \boldsymbol{\theta})$\, in Equation \ref{eq: bwl} with the expected value of the periodogram, 
+To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S(\omega; \boldsymbol{\theta})$\, in Equation \eqref{eq: bwl} with the expected value of the periodogram,
 \begin{equation}
@@
 \label{sec: semi}
-The authors note that the fitting of Equation \ref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters. 
+The authors note that the fitting of Equation \eqref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters \vmcomment{I don't like the phrase "fitting of Equation (17).}. 
 They refer to making these adjustments as applying a semi-parametric approach. 
 In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
-Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations. 
+Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations \vmcomment{correlations between correlations?
+just checking.}.
 To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.
@@
 
-\par For all two hundred trajectories, the full six-parameter model, given in Equation \ref{eq: fullSpec}, is fit to each trajectory. 
+\par For all two hundred trajectories, the full six-parameter model, given in Equation \eqref{eq: fullSpec}, is fit to each trajectory. 
 The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
-The parameters are assumed constant, since this portion of the time series does is fairly short. 
+The parameters are assumed constant, since this portion of the time series \vmdel{does} is fairly short. 
 In Figure \ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
@@
     \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.png}
-        \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands. The red line in the top figure is the inertial frequency.}
+        \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands.
+The red line in the top figure is the inertial frequency.
+\vmcomment{Add x-axis label to the bottom plot.}}
         	\label{fig: fig8}
@@
 The dotted red line indicates the value at which test statistics indicate a significant difference between the model.
-The first figure uses the window size specified in \citet{Sykulski2016} and the second figure doubles the window size.}
+The first figure uses the window size specified in \citet{Sykulski2016} and the second figure doubles the window size.
+\vmcomment{Add x-axis label to the bottom plot.}}
         	\label{fig: fig9}
@@
 
+\vmcomment{Remove URLs from References}
+
 \bibliography{prelim}

### 2016-06-03 Paper/paper.tex
minor updates

@@
   \centering
-    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
+    \includegraphics[width=.55\textwidth]{ReplicatedFigures/fig1.pdf}
         \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue.
-The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.
-        \vmcomment{You could crop away the white space in the figure.}}
+The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.}
         	\label{fig: fig1}
@@
 Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
-Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \vmadd{process} \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
+Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) process \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations.
 Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
@@
 	\begin{align}
-Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt\vmadd{,} && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
+Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt,  && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
 \end{align}
@@
 Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
-We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $\vmadd{,} where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
+We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $, where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
 \begin{align}
@@
 We begin by discussing the model for the inertial oscillations. 
-In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations\vmadd{:}
+In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations:
 \begin{align}
@@
 where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}.
-Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows\vmadd{:}
+Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows:
 \begin{align}
@@
 \nonumber
-\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\vmadd{,} \\ \nonumber
+\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}, \\ \nonumber
 dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
@@
 \end{align}
-\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$\vmadd{,}  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
+\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$,  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
  \begin{equation}
@@
 \end{equation}  
-Equation \eqref{eq: ouEq}  is known as the complex-valued \vmdel{Ornstein-Uhlenbeck (}OU\vmdel{)} stochastic differential equation. 
+Equation \eqref{eq: ouEq}  is known as the complex-valued OU stochastic differential equation. 
 In general,  OU processes are considered continuous-time equivalents to the AR(1) process.
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated \vmdel{by} \vmadd{from} the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated  from the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
 \boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
-\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)\vmadd{,}
+\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0),
 \end{align}
@@
  \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &=
-\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right)\vmadd{,}
+\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right),
 \end{align}
-where the second equality can be \vmdel{seen} \vmadd{obtained} by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
-Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined \vmdel{in} \vmadd{by} \citet{Sykulski2016}\vmadd{:}
+where the second equality can be obtained by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
+Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined by \citet{Sykulski2016}:
 \begin{align}
 \label{eq: ouAC} \nonumber
-s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0})\vmadd{,} \\
+s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}), \\
 &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
@@
   \centering
-    \includegraphics[width=.9\textwidth]{ReplicatedFigures/fig3.pdf}
+    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig3.pdf}
         \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes.
@@
 Bottom: Periodograms of both processes on the decibel scale.
-\vmcomment{Label all axes.}}
+}
         	\label{fig: fmbMat}
@@
 \label{sec: semi}
-The authors note that the fitting of Equation \eqref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters \vmcomment{I don't like the phrase "fitting of Equation (17).}. 
-They refer to making these adjustments as applying a semi-parametric approach. 
+The authors note that how the overall model is fit can be adjusted to reflect background knowledge.
+They refer to doing this as applying a semi-parametric approach.  
 In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
-Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations \vmcomment{correlations between correlations?
-just checking.}.
+Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative frequencies. 
 To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
+    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig4.pdf}
         \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}.
@@
 The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
-The parameters are assumed constant, since this portion of the time series \vmdel{does} is fairly short. 
+The parameters are assumed constant, since this portion of the time series  is fairly short. 
 In Figure \ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig5.png}
+    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig5.png}
         \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red).
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig6.pdf}
+    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig6.pdf}
         \caption{Left: The trajectory of the drifter in the south Pacific ocean.
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig7.png}
+    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig7.png}
         \caption{Top: Observed spectra of the drifter over time; Bottom: Modeled spectra over time.
@@
   \centering
-    \includegraphics[width=\textwidth]{ReplicatedFigures/fig8.png}
-        \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands.
-The red line in the top figure is the inertial frequency.
-\vmcomment{Add x-axis label to the bottom plot.}}
+    \includegraphics[width=.8\textwidth]{ReplicatedFigures/fig8.png}
+        \caption{Estimates of the six model parameters overtime with 95$\%$ confidence interval bands. The red line in the top figure is the inertial frequency. }
         	\label{fig: fig8}
@@
   \centering
-    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig9Extend.pdf}
+    \includegraphics[width=\textwidth]{ReplicatedFigures/fig9Extend.pdf}
         \caption{Time series of the likelihood ratio test statistic comparing the 5-parameter to 6-parameter models.
 The dotted red line indicates the value at which test statistics indicate a significant difference between the model.
-The first figure uses the window size specified in \citet{Sykulski2016} and the second figure doubles the window size.
-\vmcomment{Add x-axis label to the bottom plot.}}
+The first figure uses the window size specified in \citet{Sykulski2016} and the second figure doubles the window size.}
         	\label{fig: fig9}
@@
         \caption{Table of the average correlation between parameters.
-Note that there are  minor differences in the correlation values between these results and those in \citet{Sykulski2016} These differences can be attributed to slight differences in the numerical optimization used by R and Matlab in calculating the Hessian.
+Note that there are  minor differences in the correlation values between these results and those in \citet{Sykulski2016} These differences can be attributed to differences in the numerical optimization used by R and Matlab in calculating the Hessian.
 }
@@
 
-\vmcomment{Remove URLs from References}
-
 \bibliography{prelim}

### 2016-06-03 Code/figures.R

@@ -31,9 +31,8 @@ gStartTs <- 4201; gEndTs <- 4800 #index of 50 day period in blue is obtained fro
 
 #plot time series, coloring particular sections 
 #pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig1.pdf',
-#    height = 3, width = 4)
-#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(5,5,4,1.1)) #for pretty exporting
-par(mfrow = c(1, 1))
+#    height = 4, width = 5)
+par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.5,3,2,1.1), oma = c(1, 1, 1, 1)) #for pretty exporting
 plot(drift$lon[startTs:endTs], drift$lat[startTs:endTs], type = "l", 
      xlab = "Longitude", ylab = "Latitude", main = "Replication of Figure 1 (Right)")
 points(drift$lon[gStartTs:gEndTs], drift$lat[gStartTs:gEndTs], type = "l", col = "green")
@@ -90,7 +89,7 @@ gCF <- mean(-4*pi*(drift$f[gStartTs:gEndTs]))
 bFit <- fitModel(bZ, bCF, delta = DELTA, fracNeg = 0.4, fracPos = 0, quantSet = .5)
 #pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig4.pdf',
 #    height = 4, width = 8.5)
-#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4,5,4,1.1)) #for pretty exporting
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(5,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1,2))
 plot(DELTA*blue$omega, blue$sZ, type = "l", col = "blue", xlab = expression(paste(omega, Delta)),
      ylab = "dB", xlim = c(-pi, pi + 1), ylim = c(-20, 60))
@@ -144,9 +143,9 @@ N <- 1000
 fbmSamp <- simFBM(B = 10, alpha  = 0.9, H = 0.1, N = N)
 
 #plot figure
-pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig3.pdf',
-    height = 6, width = 6)
[…]

### 2016-06-03 Code/functions.R

@@ -159,7 +159,6 @@ llSimp <- function(theta, delta, curr, firstIndex, lastIndex, medIndex,
     tempIndex <- tempIndex[-which(tempIndex == medIndex)]
     llVal <- sum(curr$sZ[tempIndex]/sBar[tempIndex]) + sum(log(sBar[tempIndex]))
   }
-  print(llVal)
   return(llVal)
 }
 

### 2016-06-04 Paper/paper.tex
submitted version

@@
  Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
-We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
+We conclude the report with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.
 
@@
 	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
-	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
+	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
 	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.
@@
 Although intuitively appealing, this estimator has known problems that stem from  sampling a time series at discrete intervals. 
-Specifically, \textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.
+Specifically, \textbf{\it{aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.
 \textbf{\it{Leakage}} describes how as a result of convolution, power from one part of the spectral density function can be transferred to another part of the density.
@@
 \end{align}
-\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$,  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
+\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$,  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \eqref{eq:diffEqDeriv},
  \begin{equation}
@@
 
-To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated  from the data.
+To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial frequency, $f_{0}$,  in Equation \eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated  from the data.
 If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.
@@
 For time series modeling, we are interested in the autocovariance of this process at different time lags. 
-To derive the autocovariance as given in Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
+We derive the autocovariance as given in Equation 4 of \citet{Sykulski2016} by following \citet{Arato1999} and considering  the bivariate rather than complex-valued OU process.
 This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ with time lag $\tau >0$ is defined as 
@@
 We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$.
-Thus, Equation \ref{eq: expec} simplifies to
+Thus, Equation \eqref{eq: expec} simplifies to
 \begin{equation}
@@
 where the second equality can be obtained by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
-Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined by \citet{Sykulski2016}:
+Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as used in \citet{Sykulski2016}:
 \begin{align}
@@
 \label{eq:ouPSD}
-S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
+S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(o)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
 \end{align}
@@
 \begin{align*}
-l_{t}(\boldsymbol{\theta}) = - \frac{1}{2} \log \left|\boldsymbol{\Sigma}_{z}(\boldsymbol{\theta}) \right| - \frac{1}{2} Z^{T} \boldsymbol{\Sigma}_{z}^{-1}(\boldsymbol{\theta})\boldsymbol{Z},
+l_{t}(\boldsymbol{\theta}) = - \frac{1}{2} \log \left|\boldsymbol{\Sigma}_{z}(\boldsymbol{\theta}) \right| - \frac{1}{2} \boldsymbol{Z}^{T} \boldsymbol{\Sigma}_{z}^{-1}(\boldsymbol{\theta})\boldsymbol{Z},
 \end{align*}
@@
 The authors note that how the overall model is fit can be adjusted to reflect background knowledge.
-They refer to doing this as applying a semi-parametric approach.  
-In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
+They refer to making these adjustments as applying a semi-parametric approach.  
+In their analysis, they apply this approach to account for information about the the physical system and to avoid artifacts from interpolation. 
 Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative frequencies. 
-To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.
+To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation (which is known based on the latitude of the drifter).
 Similarly, sampling effects due to interpolation make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model.
@@
 \end{equation}
- now be a functions of time. 
+ now be functions of time. 
 Similarly, the time-varying version of the Mat\'{e}rn power spectral density is 
@@
 The simulation is set to be `very similar' to the SPEM 5.2 set-up of \citet{Danioux2008}. 
-In the  simulation, the turbulence is designed to be at a steady-state and wind forcing is set to be  constant. 
+In the  simulation, the turbulence is designed to be at steady-state and wind forcing is set to be  constant. 
 The simulated data is useful as a verification tool, since  it contains no eddies and its only  notable asymmetry between the two directions of rotation is due to the inertial oscillations.
@@
   The fitted density appears to be a smoothed version of the observed density,  indicating a reasonable model fit.
-   Typically the inertial frequency matches the Coriolis frequency, which would be expected given that shifts from the Coriolis frequency are only rare events.
+   Typically the inertial frequency matches the Coriolis frequency, which would be expected given that shifts from the Coriolis frequency are not particularly common.
 
@@
 This provide an efficient way to detect eddies for large numbers and/or long time series. 
-So, the authors have  satisfied their scientific, and, in so doing, introduced new statistical ideas.
+So, the authors have  satisfied their scientific goal, and, in so doing, introduced new statistical ideas.
 
@@
   This would subsequently invalidate conclusions about observed changes in specific parameters, since they could be masked by changes in other parameters. 
- For example, in Figure \ref{fig: fig10}, we see a high correlation (0.74) between the OU dampening and amplitude parameters, suggesting caution should be taken in interpreting their physical meaning.
+ For example, in Figure \ref{fig: fig10}, we see a high correlation (0.74) between the OU dampening and amplitude parameters, suggesting caution should be taken in interpreting the physical meaning of these parameters.
  For the primary variable of interest, inertial frequency, all correlations with other parameters are low, so the main conclusions of the paper are not affected by this issue. 
@@
 \bibliography{prelim}
+
+\clearpage
 

### 2016-06-04 Code/figures.R

@@ -286,9 +286,9 @@ nWin <- 1000 #rolling window length
 xVal <- (500:(N - 500))/MPD #Convert x-axis to days not number of observations
 
 ###Figure 6###
-pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig6.pdf',
-    height = 4, width = 8.5)
-par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4,5,4,1.1)) #for pretty exporting
+#pdf('/users/hdirector/Documents/prelim/prelim/Paper/ReplicatedFigures/fig6.pdf',
+#    height = 4, width = 8.5)
+#par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(4,5,4,1.1)) #for pretty exporting
 par(mfrow = c(1, 2))
 #plot langrangian time series
 plot(driftUlys$lon, driftUlys$lat, type = "l", col = "blue", xlim = c(-120, -80),
@@ -318,8 +318,8 @@ points(DELTA*samp$omega, 10*log10(sBar), col = "red", lty = 3, type = "l", lwd =
 points(DELTA*samp$omega[sampFit$firstIndex:sampFit$lastIndex], 10*log10(sBar)[sampFit$firstIndex:sampFit$lastIndex], 
        col = "green", type = "l", lwd = 3)
 abline(v = CF, lwd = 3) #inertial frequency
-legend("topright", legend = c("data", "fit", "extended fit"),
-       lty = c("solid", "solid", "dotted"), col = c("blue", "green", "red"),
+legend("topright", legend = c("data", "fit", "extended fit", "inertial freq."),
+       lty = c("solid", "solid", "dotted", "solid"), col = c("blue", "green", "red", "black"),
        cex = .7, bg = "white", horiz = FALSE)
 dev.off()
 
